### 2020-09-25 paper/paper.tex
working on relations and paper

\documentclass{article}

\usepackage{arxiv}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{lipsum}
\usepackage{amsthm}
\usepackage{mathtools}
\usepackage{stmaryrd}
\usepackage{amsfonts}

\theoremstyle{plain}% default
\newtheorem{thm}{Theorem}[section]
\newtheorem{lem}[thm]{Lemma}
\newtheorem{prop}[thm]{Proposition}
\newtheorem*{cor}{Corollary}
\theoremstyle{definition}
\newtheorem{defn}{Definition}[section]
\newtheorem{conj}{Conjecture}[section]
\newtheorem{exmp}{Example}[section]
\newtheorem{exrc}[exmp]{Exercise}
\theoremstyle{remark}
\newtheorem*{comm}{Comment}
\newtheorem*{note}{Note}
\newtheorem{caso}{Case}

\def\C{\mathbb{C}}
\def\N{\mathbb{N}}
\def\Q{\mathbb{Q}}
\def\R{\mathbb{R}}
\def\Z{\mathbb{Z}}
\def\K{\mathbb{K}}

\DeclarePairedDelimiter\sqmap{\llbracket}{\rrbracket}

% matrix commands
\newcommand{\vvec}[2]{
	\begin{bmatrix}
		#1 \\ #2
	\end{bmatrix}
}
\newcommand{\vvvec}[3]{
	\begin{bmatrix}
		#1 \\ #2 \\ #3
	\end{bmatrix}
}

\title{Go Implementation of Up-To Techniques for Equivalence of Weighted Languages}

%\thanks{Use footnote for providing further
%information about author (webpage, alternative
%address)---\emph{not} for acknowledging funding agencies.}

\author{
  Alessandro Cheli \\
  Undergraduate Student \\
  Department of Computer Science \\
  Università di Pisa\\
  Pisa, PI 56127 \\
  \texttt{a.cheli6@studenti.unipi.it} \\
  %% examples of more authors
  %% \And
  %% Elias D.~Striatum \\
  %% Department of Electrical Engineering\\
  %% Mount-Sheikh University\\
  %% Santa Narimana, Levand \\
  %% \texttt{stariate@ee.mount-sheikh.edu} \\
  %% \AND
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
}

\begin{document}
\maketitle

\begin{abstract}
Weighted automata generalize non-deterministic automata by adding
a quantity expressing the weight (or probability) of the execution of each transition.
In this work we propose an implementation of two algorithms for computing language 
equivalence in finite state linear weighted automata (LWAs). The first, a
linear partition refinement algorithm, calculates the linear weighted bisimulation
for any given LWA, the second algorithm checks the equivalence 
of two vectors (states) for a given weighted automata, and is used to verify the results
of the first algorithm. We then provide and compare runtime results. 
\end{abstract}

% keywords can be removed
\keywords{First keyword \and Second keyword \and More}

\section{Introduction}
\label{sec:intro}

In \cite{DBLP:journals/corr/Bonchi0K17}, up-to techniques are defined for
weighted systems over arbitrary semirings, while in \cite{BONCHI201277}, up-to techniques
are studied under a more abstract coalgebraic perspective. For the sake of simplicity of 
the implementation, we focus only on weighted automata over the field of real numbers 
$\R$. Real numbers are implemented with double precision floating point numbers,  
known as the \texttt{float64} type in the Go programming language.

\section{Preliminaries and Notation}
\label{sec:notation}

\begin{defn}
  A \textit{weighted automaton} over a field $\K$ and an alphabet $A$ is a triple 
  $(X,o,t)$ such that $X$ is a finite set of states, $t = (t_a : X \to K)_{a \in A}$
  is a set of transition functions indexed over the symbols of the alphabet and $o : X 
  \to \K$ is the output function. The transition functions $t_a$ are represented as
  $\K^{X \times X}$ matrices. $o \in \K^X$ is represented as a row vector.
  $t_a(v)$ denotes the vector obtained by multiplying the matrix $t_a$ by the vector $v 
  \in \K^X$. $o(v)$ denotes the scalar $s \in \K^X$ obtained by multiplying the row
  vector $o$ by the column vector $v \in \K^X$.
  $A^*$ is the set of all words over $A$. $\epsilon$ is the empty word and $aw$ is the
  concatenation of a letter $a$ to the word $w \in A^*$.
  A weighted language is a function $\psi: A^* \to \K$.
  A function mapping each state vector into its 
  accepted language, $\sqmap{\cdot}: \K^X \to K^{A^*}$ is defined as follows for every weighted automaton:

  \begin{equation*}
    \begin{aligned}
      \forall v \in \K^X, a \in A, w \in A^* \quad \quad
      \sqmap{v}(\epsilon) = o(v) \quad \quad
      \sqmap{v}(aw) = \sqmap{t_a(v)}(w)  
    \end{aligned}
  \end{equation*}

  Two vectors $v_1, v_2 \in \K^X$ are called language equivalent $v_1 \sim v_2 \iff \sqmap
  {v_1} = \sqmap{v_2}$. One can extend the notion of language equivalence to states rather
  than for vectors by assigning to each state $x \in X$ the corresponding unit vector 
  $e_x \in \K^X$. When given an initial state $i$ for a weighted automaton, the language 
  of the automaton can be defined as $\sqmap{i}$.
\end{defn}

\begin{defn}
  A binary relation $R \subseteq X \times Y$ between two sets $X, Y$ is a subset of the 
  cartesian product of the sets. A relation is called \textit{homogeneous} or an \textit
  {endorelation} if it is a binary relation over $X$ and itself: $R \subseteq X \times
   X$. 
  In such case, it is simply called a binary relation over $X$.
  An \textit{equivalence relation} is a binary relation that is reflexive, symmetric and
  transitive. 
\end{defn}

An equivalence relation which is compatible with all the operations of
the algebraic structure on which it is defined on, is called a 
\textit{congruence relation}. Compatibility with the algebraic structure operations
means that algebraic operations applied on equivalent elements will still
yield equivalent elements.

\begin{defn}
  The congruence closure $c(R)$ of a relation R is the smallest congruence relation 
  $R'$ such that $R \subseteq R'$ 
\end{defn}

\begin{defn}
  \textbf{} \\

\end{defn}

\section{Implementation}
\label{sec:impl}

The algorithms and data structures for this paper are implemented in the Go programming 
language. This implementation makes use of the \textit{Gonum} library for numerical 
computations. We only import the Gonum libraries for matrices and linear algebra 
and visual plotting of samples and functions.

See Section \ref{sec:headings}.

\subsection{Headings: second level}

\subsubsection{Headings: third level}

\paragraph{Paragraph}

\section{Examples of citations, figures, tables, references}
\label{sec:others}

\cite{kour2014real,kour2014fast} and see \cite{hadash2018estimate}.

The documentation for \verb+natbib+ may be found at
\begin{center}
  \url{http://mirrors.ctan.org/macros/latex/contrib/natbib/natnotes.pdf}
\end{center}
Of note is the command \verb+\citet+, which produces citations
appropriate for use in inline text.  For example,
\begin{verbatim}
   \citet{hasselmo} investigated\dots
\end{verbatim}
produces
\begin{quote}
  Hasselmo, et al.\ (1995) investigated\dots
\end{quote}

\begin{center}
  \url{https://www.ctan.org/pkg/booktabs}
\end{center}

\subsection{Figures}
See Figure \ref{fig:fig1}. Here is how you add footnotes. \footnote{Sample of the first footnote.}

\begin{figure}
  \centering
  \fbox{\rule[-.5cm]{4cm}{4cm} \rule[-.5cm]{4cm}{0cm}}
  \caption{Sample figure caption.}
  \label{fig:fig1}
\end{figure}

\subsection{Tables}
See awesome Table~\ref{tab:table}.

\begin{table}
 \caption{Sample table title}
  \centering
  \begin{tabular}{lll}
    \toprule
    \multicolumn{2}{c}{Part}                   \\
    \cmidrule(r){1-2}
    Name     & Description     & Size ($\mu$m) \\
    \midrule
    Dendrite & Input terminal  & $\sim$100     \\
    Axon     & Output terminal & $\sim$10      \\
    Soma     & Cell body       & up to $10^6$  \\
    \bottomrule
  \end{tabular}
  \label{tab:table}
\end{table}

\subsection{Lists}
\begin{itemize}
\item Lorem ipsum dolor sit amet
\item consectetur adipiscing elit. 
\item Aliquam dignissim blandit est, in dictum tortor gravida eget. In ac rutrum magna.
\end{itemize}

\bibliographystyle{unsrt}  
\bibliography{references}  %%% Remove comment to use the external .bib file (using bibtex).
%%% and comment out the ``thebibliography'' section.

% 
% %%% Comment out this section when you \bibliography{references} is enabled.
% \begin{thebibliography}{1}
% 
% \bibitem{kour2014real}
% George Kour and Raid Saabne.
% \newblock Real-time segmentation of on-line handwritten arabic script.
% \newblock In {\em Frontiers in Handwriting Recognition (ICFHR), 2014 14th
%   International Conference on}, pages 417--422. IEEE, 2014.
% 
% \bibitem{kour2014fast}
% George Kour and Raid Saabne.
% \newblock Fast classification of handwritten on-line arabic characters.
% \newblock In {\em Soft Computing and Pattern Recognition (SoCPaR), 2014 6th
%   International Conference of}, pages 312--318. IEEE, 2014.
% 
% \bibitem{hadash2018estimate}
% Guy Hadash, Einat Kermany, Boaz Carmeli, Ofer Lavi, George Kour, and Alon
%   Jacovi.
% \newblock Estimate and replace: A novel approach to integrating deep neural
%   networks with existing applications.
% \newblock {\em arXiv preprint arXiv:1804.09028}, 2018.
% 
% \end{thebibliography}
% 
% 
\end{document}

### 2020-09-25 automata/relation.go

@@ -12,6 +12,7 @@ type Relation struct {
 	u []*mat.VecDense // generating set for the congruence closure
 }
 
+// creates a new empty relation
 func NewRelation() Relation {
 	return Relation{}
 }
@@ -28,13 +29,13 @@ func (r Relation) Has(p *Pair) bool {
 
 // adds a pair of vectors to the relation
 func (r *Relation) Add(p *Pair) {
-	// if the set already contains the pair, return
+	// if the set already contains the pair (v, v'), return
 	if r.Has(p) {
 		return
 	}
-	// add the pair to the set
+	// add the pair (v,v') to the set
 	r.s = append(r.s, p)
-	// subtract the elements of the pair
+	// sub = v - v'
 	sub := mat.VecDenseCopyOf(p.Left)
 	sub.SubVec(p.Left, p.Right)
 
@@ -50,10 +51,13 @@ func (r *Relation) Add(p *Pair) {
 	}
 }
 
+// check if a pair of vectors is in a relation's congruence closure.
 func (r Relation) PairIsInCongruenceClosure(p *Pair) bool {
+	// sub = v - v'
 	sub := mat.VecDenseCopyOf(p.Left)
 	sub.SubVec(p.Left, p.Right)
 
+	// (v, v') ∈ c(R) iff v - v' ∈ U_R
 	for _, v := range r.u {
 		if mat.Equal(v, sub) {
 			return true

### 2020-09-27 paper/paper.tex
congruence closure algorithm implemented

@@
 In this work we propose an implementation of two algorithms for computing language 
-equivalence in finite state linear weighted automata (LWAs). The first, a
-linear partition refinement algorithm, calculates the linear weighted bisimulation
-for any given LWA, the second algorithm checks the equivalence 
-of two vectors (states) for a given weighted automata, and is used to verify the results
-of the first algorithm. We then provide and compare runtime results. 
+equivalence in finite state weighted automata (WAs). The first, a
+linear partition refinement algorithm, calculates the largest linear weighted bisimulation
+for any given LWA (Linear Weighted Automata) through an iterative method, 
+the second algorithm checks the language equivalence 
+of two vectors (states) for a given weighted automata by using an additional
+data structure representing a congruence relationship between states.
+We then compare the two algorithms results on randomly generated automata samples,
+to verify their correctness and provide some runtime statistics.
+
 \end{abstract}
@@
 weighted systems over arbitrary semirings, while in \cite{BONCHI201277}, up-to techniques
-are studied under a more abstract coalgebraic perspective. For the sake of simplicity of 
-the implementation, we focus only on weighted automata over the field of real numbers 
-$\R$. Real numbers are implemented with double precision floating point numbers,  
-known as the \texttt{float64} type in the Go programming language.
+are defined for Linear Weighted Automata (LWAs), under a more abstract coalgebraic perspective.
+For ease of implementation, we focus only on weighted automata over the field of real numbers 
+$\R$.
 
@@
   \to \K$ is the output function. The transition functions $t_a$ are represented as
-  $\K^{X \times X}$ matrices. $o \in \K^X$ is represented as a row vector.
-  $t_a(v)$ denotes the vector obtained by multiplying the matrix $t_a$ by the vector $v 
-  \in \K^X$. $o(v)$ denotes the scalar $s \in \K^X$ obtained by multiplying the row
-  vector $o$ by the column vector $v \in \K^X$.
+  $\K^{X \times X}$ matrices. $o \in \K^{1 \times X}$ is represented as a row vector.
+  $t_a(v)$ denotes the vector obtained by multiplying the matrix $t_a$ by the column 
+  vector $v  \in \K^{X \times 1}$. $o(v)$ denotes the scalar $s \in \K$ obtained by 
+  multiplying the row vector $o$ by the column vector $v \in \K^{X \times 1}$.
   $A^*$ is the set of all words over $A$. $\epsilon$ is the empty word and $aw$ is the
@@
 
-  Two vectors $v_1, v_2 \in \K^X$ are called language equivalent $v_1 \sim v_2 \iff \sqmap
+  Two vectors $v_1, v_2 \in \K^{X\times 1}$ are called language equivalent, denoted with $v_1 \sim v_2 \iff \sqmap
   {v_1} = \sqmap{v_2}$. One can extend the notion of language equivalence to states rather
@@
 and visual plotting of samples and functions.
+
+Real numbers are implemented with double precision floating point numbers,  
+known as the \texttt{float64} type in the Go programming language.
 

### 2020-09-27 automata/automata.go

@@ -16,6 +16,17 @@ func errRead(msg string) error { return errors.New("could not read automaton:" +
 
 type Automaton struct {
 	A []string              // The input alphabet
-	T map[string]*mat.Dense // Transition matrices are maps from input symbols to dense real valued matrices
+	T map[string]*mat.Dense // Transition matrices are maps from input strings to dense real valued matrices
 	O *mat.VecDense         // Output vector uses a dense real valued vector
+}
+
+// applies the output function to a given state vector (o * v)
+func (a Automaton) GetOutput(v *mat.VecDense) float64 {
+	return mat.Dot(a.O, v)
+}
+
+func (a Automaton) ApplyTransition(s string, v *mat.VecDense) *mat.VecDense {
+	res := mat.VecDenseCopyOf(v)
+	res.MulVec(a.T[s], v)
+	return res
 }

### 2020-09-27 automata/hkc.go

@@ -1,0 +1,72 @@
+package automata
+
+import (
+	"fmt"
+
+	"gonum.org/v1/gonum/mat"
+)
+
+// checks the language equivalence of two state vectors for a given weighted automaton
+func (a Automaton) HKC(v1, v2 *mat.VecDense) (bool, error) {
+	rel := NewRelation()
+	todo := NewPairStack()
+
+	p, err := NewPair(v1, v2)
+	if err != nil {
+		return false, err
+	}
+
+	fmt.Println("p = ", p)
+
+	// insert (v1, v2) into the todo list
+	todo.Push(*p)
+
+	i := 0
+	for !todo.Empty() {
+		fmt.Println("Step", i)
+		fmt.Println("todo =", todo)
+
+		// extract (v1', v2') from todo
+		q, err := todo.Pop()
+		if err != nil {
+			return false, err
+		}
+		fmt.Println("Extracted pair is", q)
+
+		fmt.Println("todo =", todo)
+
+		if rel.PairIsInCongruenceClosure(q) {
+			fmt.Println("(v1', v2') \\in c(R)")
+			continue
+		}
+
+		o1 := a.GetOutput(q.Left)
+		o2 := a.GetOutput(q.Right)
+		fmt.Println("o(v1) =", o1)
+		fmt.Println("o(v2) =", o2)
+		if o1 != o2 {
+			fmt.Println("o(v1) =/= o(v2)")
+			return false, nil
+		}
+
+		for _, sym := range a.A {
+
+			w1 := a.ApplyTransition(sym, q.Left)
+			w2 := a.ApplyTransition(sym, q.Right)
+			wp, err := NewPair(w1, w2)
+			if err != nil {
+				return false, err
+			}
+
+			todo.Push(*wp)
+		}
+
+		// insert (v1', v2') into R
+		rel.Add(q)
+
+		fmt.Println("R = ", rel)
+		i++
+	}
+
+	return true, nil
+}

### 2020-09-27 automata/io.go

@@ -124,6 +124,8 @@ func ReadAutomaton(r io.Reader, prompt bool) (*Automaton, error) {
 	}
 	a.O = mat.NewVecDense(len(outv), outv)
 
+	fmt.Println(a.O.T().Dims())
+
 	// Read a transition matrix for each symbol in the alphabet
 	a.T = make(map[string]*mat.Dense)
 	for _, symb := range a.A {

### 2020-09-27 automata/pair.go

@@ -2,6 +2,7 @@ package automata
 
 import (
 	"errors"
+	"fmt"
 
 	"gonum.org/v1/gonum/mat"
 )
@@ -33,4 +34,11 @@ func (p Pair) Eqs(p1 *Pair) bool {
 	eqr := mat.Equal(p.Left, p1.Right) && mat.Equal(p.Right, p1.Left)
 
 	return eql || eqr
+}
+
+func (p Pair) String() string {
+	fl := mat.Formatted(p.Left, mat.FormatMATLAB())
+	fr := mat.Formatted(p.Right, mat.FormatMATLAB())
+
+	return fmt.Sprintf("(%.5g, %.5g)", fl, fr)
 }

### 2020-09-27 automata/pairstack.go

@@ -1,0 +1,47 @@
+package automata
+
+import (
+	"errors"
+	"fmt"
+)
+
+// stack data structure for real valued vector pairs, used in the HKC algorithm
+type PairStack struct {
+	stack []Pair
+}
+
+// create a new empty pair stack
+func NewPairStack() PairStack {
+	return PairStack{stack: []Pair{}}
+}
+
+// returns true if the stack is empty, false otherwise
+func (ps PairStack) Empty() bool {
+	return len(ps.stack) == 0
+}
+
+// returns the size
+func (ps PairStack) Size() int {
+	return len(ps.stack)
+}
+
+// push a pair into the stack
+func (ps *PairStack) Push(p Pair) {
+	ps.stack = append(ps.stack, p)
+}
+
+func (ps PairStack) String() string {
+	return fmt.Sprintf("%v", ps.stack)
+}
+
+// pop a pair from the stack
+func (ps *PairStack) Pop() (*Pair, error) {
+	siz := ps.Size()
+	if siz == 0 {
+		return nil, errors.New("stack is empty")
+	}
+	el := ps.stack[siz-1]
+	ps.stack = ps.stack[:siz-1]
+
+	return &el, nil
+}

### 2020-09-27 automata/relation.go

@@ -1,6 +1,9 @@
 package automata
 
 import (
+	"bytes"
+	"fmt"
+
 	"gonum.org/v1/gonum/mat"
 )
 
@@ -65,4 +68,12 @@ func (r Relation) PairIsInCongruenceClosure(p *Pair) bool {
 	}
 
 	return false
+}
+
+func (r Relation) String() string {
+	b := bytes.Buffer{}
+	for _, uel := range r.u {
+		b.WriteString(fmt.Sprintf("%.5g,", mat.Formatted(uel, mat.FormatMATLAB())))
+	}
+	return fmt.Sprintf("(s = %v, u = [%s])", r.s, b.String())
 }

### 2020-09-27 main.go

@@ -15,22 +15,30 @@ func check(err error) {
 }
 
 func main() {
-	v1 := mat.NewVecDense(3, []float64{1.0, 2.0, 3.0})
-	v2 := mat.NewVecDense(3, []float64{3.0, 2.0, 1.0})
-	p1, err := aut.NewPair(v1, v2)
-	check(err)
+	/* 		v1 := mat.NewVecDense(3, []float64{1.0, 2.0, 3.0})
+	   	v2 := mat.NewVecDense(3, []float64{3.0, 2.0, 1.0})
+	   	p1, err := aut.NewPair(v1, v2)
+	   	check(err)
 
-	v3 := mat.NewVecDense(3, []float64{4.0, 3.0, 5.0})
-	v4 := mat.NewVecDense(3, []float64{6.0, 3.0, 3.0})
-	p2, err := aut.NewPair(v3, v4)
-	check(err)
+	   	v3 := mat.NewVecDense(3, []float64{4.0, 3.0, 5.0})
+	   	v4 := mat.NewVecDense(3, []float64{6.0, 3.0, 3.0})
+	   	p2, err := aut.NewPair(v3, v4)
+	   	check(err)
 
 	r := aut.NewRelation()
 	r.Add(p1)
 
 	fmt.Println(r.PairIsInCongruenceClosure(p2))
+	*/
 
 	a, err := aut.ReadAutomaton(os.Stdin, true)
 	check(err)
 	fmt.Println(a.String())
+
+	v1 := mat.NewVecDense(3, []float64{0, 1, 0})
+	v2 := mat.NewVecDense(3, []float64{0, 0, 1})
+
+	res, err := a.HKC(v1, v2)
+	fmt.Println(res, err)
+
 }

### 2020-09-28 lin/lin.go
intersection of linear subspaces

@@ -1,0 +1,2 @@
+// package lin contains miscellaneous linear algebra methods
+package lin

### 2020-09-28 lin/nullspace.go

@@ -1,0 +1,48 @@
+package lin
+
+import (
+	"log"
+	"math"
+
+	"gonum.org/v1/gonum/mat"
+)
+
+const tol = 10e-13
+
+// computes a basis for the nullspace of a matrix a through svd decomposition.
+// also returns the maximum residual
+func Nullspace(a mat.Matrix) ([]*mat.VecDense, float64) {
+	// compute svd decomposition
+	var svd mat.SVD
+	if ok := svd.Factorize(a, mat.SVDFullV); !ok {
+		log.Fatal("failed to factorize A")
+	}
+	vt := mat.NewDense(1, 1, nil)
+	vt.Reset()
+	svd.VTo(vt)
+
+	ker := []*mat.VecDense{}
+
+	// residual
+	res := 0.0
+
+	// the (right) null space of A is the columns of vt corresponding to
+	// singular values equal to zero.
+	for j, σ := range svd.Values(nil) {
+		if σ <= tol {
+			v := mat.NewVecDense(1, nil)
+			v.Reset()
+			v.MulVec(a, vt.ColView(j))
+			// current residual
+			currRes := mat.Norm(v, math.Inf(1))
+			if currRes > res {
+				res = currRes
+			}
+
+			ker = append(ker, vt.ColView(j).(*mat.VecDense))
+		}
+	}
+
+	return ker, res
+
+}

### 2020-09-28 main.go

@@ -4,7 +4,7 @@ import (
 	"fmt"
 	"os"
 
-	aut "github.com/0x0f0f0f/linear-weighted-automata-bisimulation/automata"
+	aut "github.com/0x0f0f0f/lwa-techniques/automata"
 	"gonum.org/v1/gonum/mat"
 )
 
@@ -41,4 +41,16 @@ func main() {
 	res, err := a.HKC(v1, v2)
 	fmt.Println(res, err)
 
+	u1 := mat.NewVecDense(5, []float64{1.0, 2.0, 3.0, -1, 2})
+	u2 := mat.NewVecDense(5, []float64{2.0, 4.0, 7.0, 2, -1})
+	w1 := mat.NewVecDense(5, []float64{1, 2, 0, -2, -1})
+	w2 := mat.NewVecDense(5, []float64{0, 1, 1, -1, -1})
+	w3 := mat.NewVecDense(5, []float64{0, 1, -3, -6, 1})
+
+	base := aut.Intersect([]*mat.VecDense{u1, u2}, []*mat.VecDense{w1, w2, w3})
+
+	// print the intersection
+	for _, el := range base {
+		fmt.Println(mat.Formatted(el, mat.Squeeze()))
+	}
 }

### 2020-09-28 paper/paper.tex
intersection of linear subspaces
working on intersection and paper

@@
 \usepackage{amsthm}
+\usepackage{amsmath}
 \usepackage{mathtools}
+\usepackage[table]{xcolor}
 \usepackage{stmaryrd}
 \usepackage{amsfonts}
+
+% Code with syntax highlighting
+\usepackage{fancyvrb}
+\usepackage{listings}
+\usepackage{textcomp}
+\usepackage{placeins}
 
@@
 \DeclarePairedDelimiter\sqmap{\llbracket}{\rrbracket}
+\DeclarePairedDelimiter\spanset{\langle}{\rangle}
+
+\newcommand{\diag}[1]{\text{diag}\left(#1\right)}
+\newcommand{\mspan}[1]{\text{span}\left(#1\right)}
+\newcommand{\dom}{\text{dom}}
+\newcommand{\mker}[1]{\text{ker}\left(#1\right)}
+\newcommand{\mrank}[1]{\text{rank}\left(#1\right)}
 
@@
 \newcommand{\vvec}[2]{
-	\begin{bmatrix}
+	\begin{pmatrix}
 		#1 \\ #2
-	\end{bmatrix}
+	\end{pmatrix}
 }
 \newcommand{\vvvec}[3]{
-	\begin{bmatrix}
+	\begin{pmatrix}
 		#1 \\ #2 \\ #3
-	\end{bmatrix}
+	\end{pmatrix}
+}
+\newcommand{\vvvvec}[4]{
+	\begin{pmatrix}
+		#1 \\ #2 \\ #3 \\ #4
+	\end{pmatrix}
+}
+
+% Code Listings
+\definecolor{vgreen}{RGB}{104,180,104}
+\definecolor{vblue}{RGB}{49,49,255}
+\definecolor{vorange}{RGB}{255,143,102}
+\lstset{
+	basicstyle=\ttfamily,
+	columns=fullflexible,
+	keepspaces=true,
 }
@@
 linear partition refinement algorithm, calculates the largest linear weighted bisimulation
-for any given LWA (Linear Weighted Automata) through an iterative method, 
+for any given LWA (Linear Weighted Automaton) through an iterative method, 
 the second algorithm checks the language equivalence 
@@
 data structure representing a congruence relationship between states.
-We then compare the two algorithms results on randomly generated automata samples,
-to verify their correctness and provide some runtime statistics.
+We then compare the two algorithms results to verify their correctness
+on randomly generated automata samples, and provide some runtime statistics.
 
@@
   $(X,o,t)$ such that $X$ is a finite set of states, $t = (t_a : X \to K)_{a \in A}$
-  is a set of transition functions indexed over the symbols of the alphabet and $o : X 
+  is a set of transition functions indexed over the symbols of the alphabet $A$ and $o : X 
   \to \K$ is the output function. The transition functions $t_a$ are represented as
@@
   vector $v  \in \K^{X \times 1}$. $o(v)$ denotes the scalar $s \in \K$ obtained by 
-  multiplying the row vector $o$ by the column vector $v \in \K^{X \times 1}$.
+  dot product of the row vector $o$ with $v \in \K^{X \times 1}$.
   $A^*$ is the set of all words over $A$. $\epsilon$ is the empty word and $aw$ is the
-  concatenation of a letter $a$ to the word $w \in A^*$.
+  concatenation of a symbol $a$ to the word $w \in A^*$.
   A weighted language is a function $\psi: A^* \to \K$.
@@
 
-  Two vectors $v_1, v_2 \in \K^{X\times 1}$ are called language equivalent, denoted with $v_1 \sim v_2 \iff \sqmap
+  Two vectors $v_1, v_2 \in \K^{X\times 1}$ are called language equivalent, denoted with 
+  $v_1 \sim v_2 $ if and only if $ \sqmap
   {v_1} = \sqmap{v_2}$. One can extend the notion of language equivalence to states rather
@@
 
+\section{Algorithms}
+
+\subsection{Forward Partition Refinement Algorithm for the Largest Weighted Bisimulation}
+\label{sec:algo2}
+
+The algorithm is defined by the iterative method:
+\begin{eqnarray}
+  R_0 = \mker{o}, & \quad & R_{i+1} = R_i \cap \bigcap_{a \in A} t_a(R_i)^{-1} \label{fwd} 
+\end{eqnarray}
+The algorithm stops when $R_{j+1} = R_j$. Proof is available in section 4.1 of \cite{BONCHI201277}
+
+\subsection{Backwards Partition Refinement Algorithm for the Largest Weighted Bisimulation}
+\label{sec:algo2}
+
+The algorithm is defined by the iterative method:
+\begin{eqnarray}
+  R_0 = \mker{o}^0, & \quad & R_{i+1} = R_i + \sum_{a \in A} t_a(R_i)^t \label{back} 
+\end{eqnarray}
+The algorithm stops when $R_{j+1} = R_j$. Proof is available in section 4.2 of \cite{BONCHI201277}
+
 \section{Implementation}
@@
 
-\subsection{Headings: second level}
+\subsection{Input file format}
+
+\subsection{Calculating the intersection of linear subspaces}
+
+To compute algorithm (\ref{fwd}), we need an efficient way to calculate a basis 
+of the intersection of an arbitrary number of linear subspaces of $\R^n$, given their 
+spanning sets.
+
+\begin{defn}
+  \textbf{Algorithm for a basis of the intersection of subspaces} \\
+  Defined in file \texttt{lin/intersect.go}.
+  Let's consider two linear subspaces of 
+  $\R^n$ defined with their spanning sets of column vectors: 
+  $U = \spanset{u_1, u_2, \hdots, u_p}$ 
+  and $W = \spanset{w_1, w_2, \hdots, w_k}$.
+  We create the block matrix $A \in \R^{n \times (p+k)}$ defined as follows:
+
+  \begin{equation*}
+    \begin{aligned}
+      a_{ij} = \begin{cases}
+        (u_j)_i & \text{for  } j = 1, \hdots, p \quad i = 1, \hdots, n \\
+        (w_{j-p})_i & \text{for  } j = p+1, \hdots, p+k \quad i = 1, \hdots, n \\
+      \end{cases}
+    \end{aligned}
+  \end{equation*}
+
+  The matrix will have the form
+  \begin{equation*}
+    A = \begin{bmatrix}
+      \vvvec{}{u_1}{\,}, \hdots \vvvec{}{u_p}{\,}, 
+      \vvvec{}{w_1}{\,}, \hdots \vvvec{}{w_k}{\,} 
+    \end{bmatrix}
+  \end{equation*}
+
+  It follows that $\mker{A}$ is a basis for $U \cap W$. This can be done with an 
+  arbitrary finite number of subspaces of $\R^n$, as long as their spanning set is 
+  \textit{finite}.
+\end{defn}
+
+\begin{defn}
+  \textbf{Algorithm for computing the nullspace} \\
+  The algorithm implementation can be found in file \texttt{lin/nullspace.go}; it is
+  adapted from \cite{scipy/ranknullspace}. 
+  Let's consider the singular value decomposition of a matrix $A \in \R^{m \times n}$:
+
+  \begin{equation*}
+    \begin{aligned}
+      A = U \Sigma V^T & \quad & \Sigma = \diag{\sigma_1, \sigma_2, \hdots, \sigma_{\min(m, n)}  } 
+       & \quad &  U \in \R^{m \times m} & \quad & V \in \R^{n \times n}
+    \end{aligned}
+  \end{equation*}
+
+  Where $V$ and $U$ are orthogonal and the singular values are ordered: $\sigma_1 \geq \sigma_2 \geq \hdots \geq \sigma_{\min(m,n)} \geq 0$.
+  It follows that $\mrank{A}$ is equal to the number of nonzero singular values and
+  a basis of the (right) null space of $A$ is the spanning set of the columns of V
+  corresponding to singular values equal to zero. 
+\end{defn}
+
+\begin{exmp}
+  The actual Go implementation is quite verbose, we can show a shorter Python 
+  implementation of the algorithm to compute the nullspace, using the 
+  \textit{SciPy} library \cite{scipy/ranknullspace}:
+
+  \begin{lstlisting}[language=python]
+def nullspace(A, atol=1e-13, rtol=0):
+    A = np.atleast_2d(A)
+    u, s, vh = svd(A)
+    tol = max(atol, rtol * s[0])
+    nnz = (s >= tol).sum()
+    ns = vh[nnz:].conj().T
+    return ns
+    \end{lstlisting}
+\end{exmp}
 
@@
 
-\subsection{Lists}
-\begin{itemize}
-\item Lorem ipsum dolor sit amet
-\item consectetur adipiscing elit. 
-\item Aliquam dignissim blandit est, in dictum tortor gravida eget. In ac rutrum magna.
-\end{itemize}
-
 \bibliographystyle{unsrt}

### 2020-09-28 automata/subspace.go

@@ -1,12 +1,7 @@
-// this file contains method for calculating the basis of the intersection and sum
+// this file contains method for calculating a basis of the intersection and sum
 // of linear subspaces of R^n
 
 package automata
-
-import (
-	"github.com/0x0f0f0f/lwa-techniques/lin"
-	"gonum.org/v1/gonum/mat"
-)
 
 //  the Zassenhaus algorithm is a method to calculate a basis for the intersection
 // and sum of two subspaces of a vector space.
@@ -16,31 +11,3 @@ import (
 // create the block matrix of size ((m+k) * (2m))
 //
 // }
-
-func Intersect(uspan, wspan []*mat.VecDense) []*mat.VecDense {
-	// dimension of vector spaces
-	m := len(uspan)
-	k := len(wspan)
-
-	if m == 0 || k == 0 {
-		return []*mat.VecDense{}
-	}
-
-	// dimensions of vectors in U and W
-	n, _ := uspan[0].Dims()
-
-	// create block matrix a of the column vectors in U and W
-	a := mat.NewDense(n, m+k, nil)
-	var j int
-	for j = 0; j < m; j++ {
-		a.SetCol(j, uspan[j].RawVector().Data)
-	}
-	for j := 0; j < k; j++ {
-		a.SetCol(m+j, wspan[j].RawVector().Data)
-	}
-
-	// compute the nullspace
-	ker, _ := lin.Nullspace(a)
-
-	return ker
-}

### 2020-09-28 lin/intersect.go

@@ -1,0 +1,32 @@
+package lin
+
+import "gonum.org/v1/gonum/mat"
+
+// compute the intersection of two vector spaces
+func Intersect(uspan, wspan []*mat.VecDense) []*mat.VecDense {
+	// dimension of vector spaces
+	m := len(uspan)
+	k := len(wspan)
+
+	if m == 0 || k == 0 {
+		return []*mat.VecDense{}
+	}
+
+	// dimensions of vectors in U and W
+	n, _ := uspan[0].Dims()
+
+	// create block matrix a of the column vectors in U and W
+	a := mat.NewDense(n, m+k, nil)
+	var j int
+	for j = 0; j < m; j++ {
+		a.SetCol(j, uspan[j].RawVector().Data)
+	}
+	for j := 0; j < k; j++ {
+		a.SetCol(m+j, wspan[j].RawVector().Data)
+	}
+
+	// compute the nullspace
+	ker, _ := Nullspace(a)
+
+	return ker
+}

### 2020-09-28 lin/intersect_test.go

@@ -1,0 +1,31 @@
+// this file contains unit tests for intersection of linear subspaces
+
+package lin
+
+import (
+	"testing"
+
+	"github.com/stretchr/testify/assert"
+	"gonum.org/v1/gonum/mat"
+)
+
+func TestIntersect(t *testing.T) {
+
+	u1 := mat.NewVecDense(5, []float64{1.0, 2.0, 3.0, -1, 2})
+	u2 := mat.NewVecDense(5, []float64{2.0, 4.0, 7.0, 2, -1})
+	w1 := mat.NewVecDense(5, []float64{1, 2, 0, -2, -1})
+	w2 := mat.NewVecDense(5, []float64{0, 1, 1, -1, -1})
+	w3 := mat.NewVecDense(5, []float64{0, 1, -3, -6, 1})
+
+	base := Intersect([]*mat.VecDense{u1, u2}, []*mat.VecDense{w1, w2, w3})
+
+	result := []float64{-4.472135954999579277036048097215825691819190979003906250000000e-01,
+		4.472135954999578721924535784637555480003356933593750000000000e-01,
+		-4.472135954999584273039658910420257598161697387695312500000000e-01,
+		-4.472135954999575946366974221746204420924186706542968750000000e-01,
+		4.472135954999579277036048097215825691819190979003906250000000e-01}
+
+	for i, v := range result {
+		assert.Equal(t, v, base[0].AtVec(i))
+	}
+}

### 2020-09-28 lin/nullspace_test.go

@@ -1,0 +1,28 @@
+// this file contains unit tests for nullspace calculation
+
+package lin
+
+import (
+	"testing"
+
+	"github.com/stretchr/testify/assert"
+	"gonum.org/v1/gonum/mat"
+)
+
+func TestNullspace(t *testing.T) {
+	a := mat.NewDense(3, 3, []float64{1.0, 2.0, 3.0, 4.0, 5.0, 6.0, 7.0, 8.0, 9.0})
+	ker, res := Nullspace(a)
+
+	//fmt.Printf("%.60e\n", res)
+
+	expres := 8.8817841970012523233890533447265625e-16
+	assert.Equal(t, expres, res)
+
+	expker := []float64{
+		-4.082482904638624066073759877326665446162223815917968750000000e-01,
+		8.164965809277261454823815256531815975904464721679687500000000e-01,
+		-4.082482904638635723415518441470339894294738769531250000000000e-01}
+	for i, v := range expker {
+		assert.Equal(t, v, ker[0].AtVec(i))
+	}
+}

### 2020-09-28 main.go

@@ -41,16 +41,4 @@ func main() {
 	res, err := a.HKC(v1, v2)
 	fmt.Println(res, err)
 
-	u1 := mat.NewVecDense(5, []float64{1.0, 2.0, 3.0, -1, 2})
-	u2 := mat.NewVecDense(5, []float64{2.0, 4.0, 7.0, 2, -1})
-	w1 := mat.NewVecDense(5, []float64{1, 2, 0, -2, -1})
-	w2 := mat.NewVecDense(5, []float64{0, 1, 1, -1, -1})
-	w3 := mat.NewVecDense(5, []float64{0, 1, -3, -6, 1})
-
-	base := aut.Intersect([]*mat.VecDense{u1, u2}, []*mat.VecDense{w1, w2, w3})
-
-	// print the intersection
-	for _, el := range base {
-		fmt.Println(mat.Formatted(el, mat.Squeeze()))
-	}
 }

### 2020-09-29 automata/subspace.go
fixed bug in svd nullspace computation

@@ -1,13 +1,0 @@
-// this file contains method for calculating a basis of the intersection and sum
-// of linear subspaces of R^n
-
-package automata
-
-//  the Zassenhaus algorithm is a method to calculate a basis for the intersection
-// and sum of two subspaces of a vector space.
-// func Zassenhaus(uspan, wspan []*mat.VecDense) []*mat.VecDense {
-// m := len(uspan)
-// k := len(wspan)
-// create the block matrix of size ((m+k) * (2m))
-//
-// }

### 2020-09-29 lin/complement.go

@@ -1,0 +1,28 @@
+package lin
+
+import (
+	"fmt"
+
+	"gonum.org/v1/gonum/mat"
+)
+
+// compute the orthogonal complement of a vector subspace of R^n
+func Complement(n int, spanset []*mat.VecDense) []*mat.VecDense {
+	dim := len(spanset)
+	// the orthogonal complement of {} is R^n
+	if dim == 0 {
+		return UnitaryVecs(n, n)
+	}
+
+	a := mat.NewDense(len(spanset), n, nil)
+
+	for i, vec := range spanset {
+		a.SetRow(i, vec.RawVector().Data)
+	}
+
+	fmt.Printf("%.5g\n", mat.Formatted(a, mat.Squeeze()))
+
+	basis, _ := NullspaceSVD(a)
+
+	return basis
+}

### 2020-09-29 lin/complement_test.go

@@ -1,0 +1,21 @@
+package lin
+
+import (
+	"fmt"
+	"testing"
+
+	"gonum.org/v1/gonum/mat"
+)
+
+func TestComplement(t *testing.T) {
+	w1 := mat.NewVecDense(3, []float64{1, 7, 2})
+	w2 := mat.NewVecDense(3, []float64{-2, 3, 1})
+	//assert.Equal()
+
+	wperp := Complement(3, []*mat.VecDense{w1, w2})
+
+	for _, vec := range wperp {
+		fmt.Printf("%.20g\n", mat.Formatted(vec, mat.Squeeze()))
+	}
+
+}

### 2020-09-29 lin/gauss.go

@@ -1,0 +1,144 @@
+package lin
+
+import (
+	"errors"
+	"fmt"
+	"math"
+
+	"gonum.org/v1/gonum/mat"
+)
+
+type testCase struct {
+	a [][]float64
+	b []float64
+	x []float64
+}
+
+// result from above test case turns out to be correct to this tolerance.
+const ε = 1e-14
+
+// gaussian elimination with scaled partial pivoting
+func GaussPartialo(a0 [][]float64, b0 []float64) ([]float64, error) {
+	m := len(b0)
+	a := make([][]float64, m)
+	for i, ai := range a0 {
+		row := make([]float64, m+1)
+		copy(row, ai)
+		row[m] = b0[i]
+		a[i] = row
+	}
+	for k := range a {
+		iMax := 0
+		max := -1.
+		for i := k; i < m; i++ {
+			row := a[i]
+			// compute scale factor s = max abs in row
+			s := -1.
+			for j := k; j < m; j++ {
+				x := math.Abs(row[j])
+				if x > s {
+					s = x
+				}
+			}
+			// scale the abs used to pick the pivot.
+			if abs := math.Abs(row[k]) / s; abs > max {
+				iMax = i
+				max = abs
+			}
+		}
+		if a[iMax][k] == 0 {
+			return nil, errors.New("singular")
+		}
+		a[k], a[iMax] = a[iMax], a[k]
+		for i := k + 1; i < m; i++ {
+			for j := k + 1; j <= m; j++ {
+				a[i][j] -= a[k][j] * (a[i][k] / a[k][k])
+			}
+			a[i][k] = 0
+		}
+	}
+	x := make([]float64, m)
+	for i := m - 1; i >= 0; i-- {
+		x[i] = a[i][m]
+		for j := i + 1; j < m; j++ {
+			x[i] -= a[i][j] * x[j]
+		}
+		x[i] /= a[i][i]
+	}
+	return x, nil
+}
+
+func GaussPartial(a0 mat.Matrix) (*mat.Dense, *mat.VecDense, error) {
+	a := mat.DenseCopyOf(a0)
+	a.Copy(a0)
+
+	m, n := a.Dims()
+
+	fmt.Printf("==========\n%.20g\n========\n", mat.Formatted(a, mat.Squeeze()))
+
+	// permutation matrix
+	p := mat.NewDense(m, m, nil)
+
[…]

### 2020-09-29 lin/intersect.go

@@ -1,32 +1,38 @@
 package lin
 
-import "gonum.org/v1/gonum/mat"
+import (
+	"gonum.org/v1/gonum/mat"
+)
 
-// compute the intersection of two vector spaces
-func Intersect(uspan, wspan []*mat.VecDense) []*mat.VecDense {
+// compute a basis for the intersection of many vector spaces
+func Intersect(spansets ...[]*mat.VecDense) []*mat.VecDense {
 	// dimension of vector spaces
-	m := len(uspan)
-	k := len(wspan)
-
-	if m == 0 || k == 0 {
-		return []*mat.VecDense{}
+	lengths := make([]int, len(spansets))
+	m := 0
+	for i, span := range spansets {
+		lengths[i] = len(span)
+		if lengths[i] == 0 {
+			return []*mat.VecDense{}
+		}
+		m += lengths[i]
 	}
 
-	// dimensions of vectors in U and W
-	n, _ := uspan[0].Dims()
+	// dimensions of vectors in subspaces
+	n, _ := spansets[0][0].Dims()
 
 	// create block matrix a of the column vectors in U and W
-	a := mat.NewDense(n, m+k, nil)
-	var j int
-	for j = 0; j < m; j++ {
-		a.SetCol(j, uspan[j].RawVector().Data)
-	}
-	for j := 0; j < k; j++ {
-		a.SetCol(m+j, wspan[j].RawVector().Data)
+	a := mat.NewDense(n, m, nil)
+
+	j := 0
+	for _, span := range spansets {
+		for _, vec := range span {
+			a.SetCol(j, vec.RawVector().Data)
+			j++
+		}
 	}
 
 	// compute the nullspace
-	ker, _ := Nullspace(a)
+	ker, _ := NullspaceSVD(a)
 
 	return ker
 }

### 2020-09-29 lin/nullspace.go

@@ -1,6 +1,7 @@
 package lin
 
 import (
+	"fmt"
 	"log"
 	"math"
 
@@ -9,9 +10,10 @@ import (
 
 const tol = 10e-13
 
-// computes a basis for the nullspace of a matrix a through svd decomposition.
-// also returns the maximum residual
-func Nullspace(a mat.Matrix) ([]*mat.VecDense, float64) {
+// computes an orthonormal basis for the nullspace of a matrix a
+// through svd decomposition. also returns the maximum residual
+func NullspaceSVD(a mat.Matrix) ([]*mat.VecDense, float64) {
+
 	// compute svd decomposition
 	var svd mat.SVD
 	if ok := svd.Factorize(a, mat.SVDFullV); !ok {
@@ -26,23 +28,34 @@ func Nullspace(a mat.Matrix) ([]*mat.VecDense, float64) {
 	// residual
 	res := 0.0
 
+	PrintMat(vt)
+
+	fmt.Println(svd.Values(nil))
+
 	// the (right) null space of A is the columns of vt corresponding to
 	// singular values equal to zero.
-	for j, σ := range svd.Values(nil) {
+	j := 0
+	for _, σ := range svd.Values(nil) {
 		if σ <= tol {
-			v := mat.NewVecDense(1, nil)
-			v.Reset()
-			v.MulVec(a, vt.ColView(j))
-			// current residual
-			currRes := mat.Norm(v, math.Inf(1))
-			if currRes > res {
-				res = currRes
-			}
-
-			ker = append(ker, vt.ColView(j).(*mat.VecDense))
+			break
 		}
+		j++
+	}
+
+	fmt.Println(j)
+
+	for j := j; j < vt.RawMatrix().Cols; j++ {
+		v := mat.NewVecDense(1, nil)
+		v.Reset()
+		v.MulVec(a, vt.ColView(j))
+		// current residual
+		currRes := mat.Norm(v, math.Inf(1))
+		if currRes > res {
+			res = currRes
+		}
+
+		ker = append(ker, vt.ColView(j).(*mat.VecDense))
 	}
 
 	return ker, res
-
 }

### 2020-09-29 lin/nullspace_test.go

@@ -9,9 +9,9 @@ import (
 	"gonum.org/v1/gonum/mat"
 )
 
-func TestNullspace(t *testing.T) {
+func TestNullspaceSVD(t *testing.T) {
 	a := mat.NewDense(3, 3, []float64{1.0, 2.0, 3.0, 4.0, 5.0, 6.0, 7.0, 8.0, 9.0})
-	ker, res := Nullspace(a)
+	ker, res := NullspaceSVD(a)
 
 	//fmt.Printf("%.60e\n", res)
 
@@ -22,6 +22,23 @@ func TestNullspace(t *testing.T) {
 		-4.082482904638624066073759877326665446162223815917968750000000e-01,
 		8.164965809277261454823815256531815975904464721679687500000000e-01,
 		-4.082482904638635723415518441470339894294738769531250000000000e-01}
+	for i, v := range expker {
+		assert.Equal(t, v, ker[0].AtVec(i))
+	}
+}
+
+func TestNullspaceSVD2(t *testing.T) {
+	a := mat.NewDense(2, 3, []float64{1, 7, 2, -2, 3, 1})
+	ker, _ := NullspaceSVD(a)
+
+	//fmt.Printf("%.60e\n", res)
+
+	expker := []float64{5.634361698190108e-02,
+		-2.8171808490950556e-01,
+		9.578414886923188e-01}
+
+	assert.Len(t, ker, 1)
+
 	for i, v := range expker {
 		assert.Equal(t, v, ker[0].AtVec(i))
 	}

### 2020-09-29 lin/union.go

@@ -1,0 +1,13 @@
+package lin
+
+import "gonum.org/v1/gonum/mat"
+
+// compute a basis for the union of many vector spaces
+func Union(spansets ...[]*mat.VecDense) []*mat.VecDense {
+	basis := []*mat.VecDense{}
+	for _, span := range spansets {
+		basis = append(basis, span...)
+	}
+
+	return basis
+}

### 2020-09-29 main.go

@@ -2,9 +2,8 @@ package main
 
 import (
 	"fmt"
-	"os"
 
-	aut "github.com/0x0f0f0f/lwa-techniques/automata"
+	"github.com/0x0f0f0f/lwa-techniques/lin"
 	"gonum.org/v1/gonum/mat"
 )
 
@@ -31,14 +30,22 @@ func main() {
 	fmt.Println(r.PairIsInCongruenceClosure(p2))
 	*/
 
-	a, err := aut.ReadAutomaton(os.Stdin, true)
-	check(err)
-	fmt.Println(a.String())
+	// a, err := aut.ReadAutomaton(os.Stdin, true)
+	// check(err)
+	// fmt.Println(a.String())
 
-	v1 := mat.NewVecDense(3, []float64{0, 1, 0})
-	v2 := mat.NewVecDense(3, []float64{0, 0, 1})
+	// v1 := mat.NewVecDense(3, []float64{0, 1, 0})
+	// v2 := mat.NewVecDense(3, []float64{0, 0, 1})
 
-	res, err := a.HKC(v1, v2)
-	fmt.Println(res, err)
+	// res, err := a.HKC(v1, v2)
+	// fmt.Println(res, err)
+
+	w1 := mat.NewVecDense(3, []float64{1, 7, 2})
+	w2 := mat.NewVecDense(3, []float64{-2, 3, 1})
+	//assert.Equal()
+
+	wperp := lin.Complement(3, []*mat.VecDense{w1, w2})
+
+	fmt.Println(len(wperp))
 
 }

### 2020-09-30 paper/paper.tex
fixed bug in svd nullspace computation
annihilator, go code cleanup

@@
 \usepackage{amsfonts}
+\usepackage{lstlistings-golang}
 
@@
 
+% TODO linear weighted automata
 \begin{defn}
@@
 \end{eqnarray}
+Where $\mker{o}^0$ is an annihilator.
 The algorithm stops when $R_{j+1} = R_j$. Proof is available in section 4.2 of \cite{BONCHI201277}
@@
 and visual plotting of samples and functions.
-
 Real numbers are implemented with double precision floating point numbers,  
@@
 \begin{defn}
-  \textbf{Algorithm for computing the nullspace} \\
+  \textbf{Algorithm for computing the null space of a vector subspace} \\
   The algorithm implementation can be found in file \texttt{lin/nullspace.go}; it is
@@
 \begin{exmp}
-  The actual Go implementation is quite verbose, we can show a shorter Python 
+  First, we show a shorter Python 
   implementation of the algorithm to compute the nullspace, using the 
@@
 
-  \begin{lstlisting}[language=python]
-def nullspace(A, atol=1e-13, rtol=0):
-    A = np.atleast_2d(A)
-    u, s, vh = svd(A)
-    tol = max(atol, rtol * s[0])
-    nnz = (s >= tol).sum()
-    ns = vh[nnz:].conj().T
-    return ns
-    \end{lstlisting}
+  \lstinputlisting[language=python]{nullspace.py}
+
+    The Go implementation is quite longer:
+
+    \lstinputlisting[language=go]{../lin/nullspace.go}
+
 \end{exmp}
+
+\subsection{Implementing the backwards algorithm}
+
+In \cite{BONCHI201277}, for the first step of the iterative backwards 
+algorithm to compute the largest linear weighted bisimulation, we 
+need to compute $\mker{o}^0$. If $V$ is a vector space and $W$ is a
+subspace of $W$, the annihilator of $W$, respectively $W^0$ is 
+a subspace of the space $V^*$ of linear functionals on $V$.
+$W^0$ are the functionals that annihilate on $W$. Since 
+we are working on subspaces of $\R^n$, we can directly compute 
+the orthogonal complement in our implementation instead of the
+annihilator.
+
+\begin{prop}
+  Within finite dimensional spaces, the annihilator $W^0$ is isomorphic
+  to the orthogonal complement.  
+\end{prop}
+
+\begin{proof}
+  Let $V$ be an inner product space over the field $\K$ with an inner product
+  $\langle \cdot , \cdot \rangle : V \times V \to \K$ defined. 
+  With the inner product defined on $V$, every linear functional can be 
+  represented with a vector: if $\xi : V \to \K$ is a functional, $c$ 
+  is an unique vector such that $\forall x \in V \: . \: \langle c, x \rangle = \xi(x)$.
+  The map associating every $\xi \in V^*$ with the vector $c \in V$ that corresponds 
+  to $\xi$ with respect to the inner product is a linear isomorphism. Through this 
+  isomorphism, the annihilator $W^0$ becomes $W^\perp$.
+\end{proof}
+
+To compute the orthogonal complement of a vector subspace $W$, we
+compute $W^\perp = \mker{A^T}$, where $A$ is the matrix with 
+column vectors in the spanning set of $W$ as its columns. Precisely, $W$ is 
+represented as the 
+column space of $A$. Proof is available in \cite{ila}.
 

### 2020-09-30 lin/complement.go

@@ -1,18 +1,0 @@
-package lin
-
-import (
-	"gonum.org/v1/gonum/mat"
-)
-
-// compute the orthogonal complement of a vector subspace of R^n
-func Complement(spanset mat.Matrix) mat.Matrix {
-	n, m := spanset.Dims()
-	// the orthogonal complement of {} is R^n
-	if m == 0 {
-		return EyeDense(n)
-	}
-
-	basis, _ := Nullspace(spanset.T())
-
-	return basis
-}

### 2020-09-30 lin/complement_test.go

@@ -1,30 +1,0 @@
-package lin
-
-import (
-	"testing"
-
-	"github.com/stretchr/testify/assert"
-	"gonum.org/v1/gonum/mat"
-)
-
-func TestComplement(t *testing.T) {
-	wspan := mat.NewDense(3, 3, []float64{
-		1, 2, 3,
-		4, 5, 6,
-		7, 8, 9,
-	})
-	//assert.Equal()
-
-	expected := mat.NewDense(3, 1, []float64{
-		-0.408248290463863294785795687857898883521556854248046875,
-		0.81649658092772614548238152565318159759044647216796875,
-		-0.4082482904638629062077370690531097352504730224609375,
-	})
-
-	wperp := Complement(wspan)
-
-	assert.True(t, mat.Equal(expected, wperp))
-
-	//PrintMat(wperp)
-
-}

### 2020-09-30 lin/insubspace.go

@@ -1,50 +1,0 @@
-package lin
-
-import (
-	"gonum.org/v1/gonum/mat"
-)
-
-type MyDense struct {
-	m *mat.Dense
-}
-
-func (m MyDense) T() mat.Matrix {
-	return m.m.T()
-}
-
-func (m MyDense) At(i, j int) float64 {
-	return m.m.At(i, j)
-}
-
-func (m MyDense) Dims() (int, int) {
-	return m.m.Dims()
-}
-
-// returns true if vector b is contained in the subspace spanned by
-// the columns of u
-
-func (m MyDense) MulVecTo(dst *mat.VecDense, trans bool, x mat.Vector) {
-	if trans {
-		dst.MulVec(m.m.T(), x)
-		return
-	}
-	dst.MulVec(m.m, x)
-
-}
-
-func InSubspace(u *mat.Dense, b *mat.VecDense) bool {
-
-	var svd mat.SVD
-	svd.Factorize(u, mat.SVDNone)
-
-	ranku := svd.Rank(tol)
-
-	r, c := u.Dims()
-	aug := mat.NewDense(r, c+1, nil)
-
-	aug.Augment(u, b)
-	svd.Factorize(aug, mat.SVDNone)
-	rankub := svd.Rank(tol)
-
-	return ranku == rankub
-}

### 2020-09-30 lin/insubspace_test.go

@@ -1,21 +1,0 @@
-package lin
-
-import (
-	"testing"
-
-	"github.com/stretchr/testify/assert"
-	"gonum.org/v1/gonum/mat"
-)
-
-func TestInSubspace(t *testing.T) {
-	m := mat.NewDense(3, 1, []float64{1, 1, 0})
-
-	v := mat.NewVecDense(3, []float64{3, 3, 0})
-
-	nv := mat.NewVecDense(3, []float64{3, 3, 1})
-
-	assert.True(t, InSubspace(m, v))
-
-	assert.False(t, InSubspace(m, nv))
-
-}

### 2020-09-30 lin/intersect_test.go

@@ -1,40 +1,0 @@
-// this file contains unit tests for intersection of linear subspaces
-
-package lin
-
-import (
-	"testing"
-
-	"github.com/stretchr/testify/assert"
-	"gonum.org/v1/gonum/mat"
-)
-
-func TestIntersect(t *testing.T) {
-
-	u := mat.NewDense(5, 2, []float64{
-		1, 2,
-		2, 4,
-		3, 7,
-		-1, 2,
-		2, -1,
-	})
-	w := mat.NewDense(5, 3, []float64{
-		1, 0, 0,
-		2, 1, 1,
-		0, 1, -3,
-		-2, -1, -6,
-		-1, -1, 1,
-	})
-
-	base := Intersect(u, w)
-
-	expected := mat.NewDense(5, 1, []float64{
-		-0.447213595499957927703604809721582569181919097900390625,
-		0.447213595499957872192453578463755548000335693359375,
-		-0.44721359549995842730396589104202575981616973876953125,
-		-0.447213595499957594636697422174620442092418670654296875,
-		0.447213595499957927703604809721582569181919097900390625,
-	})
-
-	assert.True(t, mat.Equal(expected, base))
-}

### 2020-09-30 lin/subspace.go

@@ -1,0 +1,76 @@
+package lin
+
+import (
+	"gonum.org/v1/gonum/mat"
+)
+
+// returns true iff the vector b is included in U's column space
+// to do so we check if the matrix U and [U|b] have the same rank
+func InSubspace(u *mat.Dense, b *mat.VecDense) bool {
+	var svd mat.SVD
+	svd.Factorize(u, mat.SVDNone)
+
+	ranku := svd.Rank(tol)
+
+	// we use the same backing data for the matrix U and [U|b]
+	// we then check the rank with SVD factorization
+	r, c := u.Dims()
+	u = u.Grow(0, 1).(*mat.Dense)
+	u.SetCol(c, b.RawVector().Data)
+	svd.Factorize(u, mat.SVDNone)
+	rankub := svd.Rank(tol)
+	u = u.Slice(0, r, 0, c).(*mat.Dense)
+
+	return ranku == rankub
+}
+
+// compute a basis for the intersection of many vector spaces
+func Intersect(spansets ...mat.Matrix) mat.Matrix {
+	a := mat.DenseCopyOf(spansets[0])
+	for i, span := range spansets {
+		if i == 0 {
+			continue
+		}
+
+		ar, ac := a.Dims()
+		_, sc := span.Dims()
+		m := mat.NewDense(ar, ac+sc, nil)
+		m.Augment(a, span)
+		a = m
+	}
+	// compute the nullspace
+	ker, _ := Nullspace(a)
+	return ker
+}
+
+// compute a basis for the union of many vector spaces
+func Union(spansets ...*mat.Dense) *mat.Dense {
+	n, j := spansets[0].Dims()
+
+	basis := mat.DenseCopyOf(spansets[0])
+	for k := 1; k < len(spansets); k++ {
+		r, c := spansets[k].Dims()
+		if r != n {
+			panic("mat dimension mismatch")
+		}
+		newbasis := mat.NewDense(n, j+c, nil)
+		newbasis.Augment(basis, spansets[k])
+		j = j + c
+		basis = newbasis
+	}
+
+	return basis
+}
+
+// compute the orthogonal complement of a vector subspace of R^n
+func Complement(spanset mat.Matrix) mat.Matrix {
+	n, m := spanset.Dims()
+	// the orthogonal complement of {} is R^n
+	if m == 0 {
+		return EyeDense(n)
+	}
+
+	basis, _ := Nullspace(spanset.T())
+
+	return basis
+}

### 2020-09-30 lin/subspace_test.go

@@ -1,0 +1,102 @@
+package lin
+
+import (
+	"testing"
+
+	"github.com/stretchr/testify/assert"
+	"gonum.org/v1/gonum/mat"
+)
+
+func TestInSubspace(t *testing.T) {
+	m := mat.NewDense(3, 1, []float64{1, 1, 0})
+
+	om := mat.DenseCopyOf(m)
+
+	v := mat.NewVecDense(3, []float64{3, 3, 0})
+
+	nv := mat.NewVecDense(3, []float64{3, 3, 1})
+
+	assert.True(t, InSubspace(m, v))
+	assert.False(t, InSubspace(m, nv))
+
+	// matrix is not modified
+	assert.True(t, mat.Equal(m, om))
+}
+
+func TestIntersect(t *testing.T) {
+
+	u := mat.NewDense(5, 2, []float64{
+		1, 2,
+		2, 4,
+		3, 7,
+		-1, 2,
+		2, -1,
+	})
+	w := mat.NewDense(5, 3, []float64{
+		1, 0, 0,
+		2, 1, 1,
+		0, 1, -3,
+		-2, -1, -6,
+		-1, -1, 1,
+	})
+
+	base := Intersect(u, w)
+
+	expected := mat.NewDense(5, 1, []float64{
+		-0.447213595499957927703604809721582569181919097900390625,
+		0.447213595499957872192453578463755548000335693359375,
+		-0.44721359549995842730396589104202575981616973876953125,
+		-0.447213595499957594636697422174620442092418670654296875,
+		0.447213595499957927703604809721582569181919097900390625,
+	})
+
+	assert.True(t, mat.Equal(expected, base))
+}
+
+func TestComplement(t *testing.T) {
+	wspan := mat.NewDense(3, 3, []float64{
+		1, 2, 3,
+		4, 5, 6,
+		7, 8, 9,
+	})
+	//assert.Equal()
+
+	expected := mat.NewDense(3, 1, []float64{
+		-0.408248290463863294785795687857898883521556854248046875,
+		0.81649658092772614548238152565318159759044647216796875,
+		-0.4082482904638629062077370690531097352504730224609375,
+	})
+
+	wperp := Complement(wspan)
+
+	assert.True(t, mat.Equal(expected, wperp))
+
+	//PrintMat(wperp)
+}
+
[…]

### 2020-09-30 lin/union.go

@@ -1,14 +1,0 @@
-package lin
-
-import "gonum.org/v1/gonum/mat"
-
-// compute a basis for the union of many vector spaces
-// TODO matrix form
-func Union(spansets ...[]*mat.VecDense) []*mat.VecDense {
-	basis := []*mat.VecDense{}
-	for _, span := range spansets {
-		basis = append(basis, span...)
-	}
-
-	return basis
-}

### 2020-09-30 paper/nullspace.py

@@ -1,0 +1,7 @@
+def nullspace(A, atol=1e-13, rtol=0):
+    A = np.atleast_2d(A)
+    u, s, vh = svd(A)
+    tol = max(atol, rtol * s[0])
+    nnz = (s >= tol).sum()
+    ns = vh[nnz:].conj().T
+    return ns

### 2020-09-30 paper/notused.tex
preliminary definitions

--- 2020-09-27 paper/paper.tex
@@
-\documentclass{article}
+\subsection{Calculating the intersection of linear subspaces}
 
-\usepackage{arxiv}
-
-\usepackage[utf8]{inputenc} % allow utf-8 input
-\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
-\usepackage{hyperref}       % hyperlinks
-\usepackage{url}            % simple URL typesetting
-\usepackage{booktabs}       % professional-quality tables
-\usepackage{amsfonts}       % blackboard math symbols
-\usepackage{nicefrac}       % compact symbols for 1/2, etc.
-\usepackage{microtype}      % microtypography
-\usepackage{lipsum}
-\usepackage{amsthm}
-\usepackage{mathtools}
-\usepackage{stmaryrd}
-\usepackage{amsfonts}
-
-\theoremstyle{plain}% default
-\newtheorem{thm}{Theorem}[section]
-\newtheorem{lem}[thm]{Lemma}
-\newtheorem{prop}[thm]{Proposition}
-\newtheorem*{cor}{Corollary}
-\theoremstyle{definition}
-\newtheorem{defn}{Definition}[section]
-\newtheorem{conj}{Conjecture}[section]
-\newtheorem{exmp}{Example}[section]
-\newtheorem{exrc}[exmp]{Exercise}
-\theoremstyle{remark}
-\newtheorem*{comm}{Comment}
-\newtheorem*{note}{Note}
-\newtheorem{caso}{Case}
-
-\def\C{\mathbb{C}}
-\def\N{\mathbb{N}}
-\def\Q{\mathbb{Q}}
-\def\R{\mathbb{R}}
-\def\Z{\mathbb{Z}}
-\def\K{\mathbb{K}}
-
-\DeclarePairedDelimiter\sqmap{\llbracket}{\rrbracket}
-
-% matrix commands
-\newcommand{\vvec}[2]{
-	\begin{bmatrix}
-		#1 \\ #2
-	\end{bmatrix}
-}
-\newcommand{\vvvec}[3]{
-	\begin{bmatrix}
-		#1 \\ #2 \\ #3
-	\end{bmatrix}
-}
-
-\title{Go Implementation of Up-To Techniques for Equivalence of Weighted Languages}
-
-%\thanks{Use footnote for providing further
-%information about author (webpage, alternative
-%address)---\emph{not} for acknowledging funding agencies.}
-
-\author{
-  Alessandro Cheli \\
-  Undergraduate Student \\
-  Department of Computer Science \\
-  Università di Pisa\\
-  Pisa, PI 56127 \\
-  \texttt{a.cheli6@studenti.unipi.it} \\
-  %% examples of more authors
-  %% \And
-  %% Elias D.~Striatum \\
-  %% Department of Electrical Engineering\\
-  %% Mount-Sheikh University\\
-  %% Santa Narimana, Levand \\
-  %% \texttt{stariate@ee.mount-sheikh.edu} \\
-  %% \AND
-  %% Coauthor \\
-  %% Affiliation \\
-  %% Address \\
-  %% \texttt{email} \\
-  %% \And
-  %% Coauthor \\
-  %% Affiliation \\
-  %% Address \\
-  %% \texttt{email} \\
-  %% \And
-  %% Coauthor \\
-  %% Affiliation \\
-  %% Address \\
-  %% \texttt{email} \\
-}
-
-\begin{document}
-\maketitle
-
-\begin{abstract}
-Weighted automata generalize non-deterministic automata by adding
-a quantity expressing the weight (or probability) of the execution of each transition.
-In this work we propose an implementation of two algorithms for computing language 
-equivalence in finite state weighted automata (WAs). The first, a
-linear partition refinement algorithm, calculates the largest linear weighted bisimulation
-for any given LWA (Linear Weighted Automata) through an iterative method, 
-the second algorithm checks the language equivalence 
-of two vectors (states) for a given weighted automata by using an additional
-data structure representing a congruence relationship between states.
-We then compare the two algorithms results on randomly generated automata samples,
-to verify their correctness and provide some runtime statistics.
-
-\end{abstract}
-
-% keywords can be removed
-\keywords{First keyword \and Second keyword \and More}
-
-\section{Introduction}
-\label{sec:intro}
-
-In \cite{DBLP:journals/corr/Bonchi0K17}, up-to techniques are defined for
-weighted systems over arbitrary semirings, while in \cite{BONCHI201277}, up-to techniques
-are defined for Linear Weighted Automata (LWAs), under a more abstract coalgebraic perspective.
-For ease of implementation, we focus only on weighted automata over the field of real numbers 
-$\R$.
-
-\section{Preliminaries and Notation}
-\label{sec:notation}
+To compute algorithm (\ref{fwd}), we need an efficient way to calculate a basis 
+of the intersection of an arbitrary number of linear subspaces of $\R^n$, given their 
+spanning sets.
 
 \begin{defn}
-  A \textit{weighted automaton} over a field $\K$ and an alphabet $A$ is a triple 
-  $(X,o,t)$ such that $X$ is a finite set of states, $t = (t_a : X \to K)_{a \in A}$
-  is a set of transition functions indexed over the symbols of the alphabet and $o : X 
-  \to \K$ is the output function. The transition functions $t_a$ are represented as
-  $\K^{X \times X}$ matrices. $o \in \K^{1 \times X}$ is represented as a row vector.
-  $t_a(v)$ denotes the vector obtained by multiplying the matrix $t_a$ by the column 
-  vector $v  \in \K^{X \times 1}$. $o(v)$ denotes the scalar $s \in \K$ obtained by 
-  multiplying the row vector $o$ by the column vector $v \in \K^{X \times 1}$.
-  $A^*$ is the set of all words over $A$. $\epsilon$ is the empty word and $aw$ is the
-  concatenation of a letter $a$ to the word $w \in A^*$.
-  A weighted language is a function $\psi: A^* \to \K$.
-  A function mapping each state vector into its 
-  accepted language, $\sqmap{\cdot}: \K^X \to K^{A^*}$ is defined as follows for every weighted automaton:
+  \textbf{Algorithm for a basis of the intersection of subspaces} \\
+  Defined in file \texttt{lin/intersect.go}.
+  Let's consider two linear subspaces of 
+  $\R^n$ defined with their spanning sets of column vectors: 
+  $U = \spanset{u_1, u_2, \hdots, u_p}$ 
+  and $W = \spanset{w_1, w_2, \hdots, w_k}$.
+  We create the block matrix $A \in \R^{n \times (p+k)}$ defined as follows:
 
@@
     \begin{aligned}
-      \forall v \in \K^X, a \in A, w \in A^* \quad \quad
-      \sqmap{v}(\epsilon) = o(v) \quad \quad
-      \sqmap{v}(aw) = \sqmap{t_a(v)}(w)  
+      a_{ij} = \begin{cases}
+        (u_j)_i & \text{for  } j = 1, \hdots, p \quad i = 1, \hdots, n \\
+        (w_{j-p})_i & \text{for  } j = p+1, \hdots, p+k \quad i = 1, \hdots, n \\
+      \end{cases}
     \end{aligned}
@@
 
-  Two vectors $v_1, v_2 \in \K^{X\times 1}$ are called language equivalent, denoted with $v_1 \sim v_2 \iff \sqmap
-  {v_1} = \sqmap{v_2}$. One can extend the notion of language equivalence to states rather
-  than for vectors by assigning to each state $x \in X$ the corresponding unit vector 
-  $e_x \in \K^X$. When given an initial state $i$ for a weighted automaton, the language 
-  of the automaton can be defined as $\sqmap{i}$.
+  The matrix will have the form
+  \begin{equation*}
+    A = \begin{bmatrix}
+      \vvvec{}{u_1}{\,}, \hdots \vvvec{}{u_p}{\,}, 
+      \vvvec{}{w_1}{\,}, \hdots \vvvec{}{w_k}{\,} 
+    \end{bmatrix}
+  \end{equation*}
+
+  It follows that $\mker{A}$ is a basis for $U \cap W$. This can be done with an 
+  arbitrary finite number of subspaces of $\R^n$, as long as their spanning set is 
+  \textit{finite}.
 \end{defn}
-
-\begin{defn}
-  A binary relation $R \subseteq X \times Y$ between two sets $X, Y$ is a subset of the 
-  cartesian product of the sets. A relation is called \textit{homogeneous} or an \textit
-  {endorelation} if it is a binary relation over $X$ and itself: $R \subseteq X \times
-   X$. 
-  In such case, it is simply called a binary relation over $X$.
-  An \textit{equivalence relation} is a binary relation that is reflexive, symmetric and
-  transitive. 
-\end{defn}
-
-An equivalence relation which is compatible with all the operations of
-the algebraic structure on which it is defined on, is called a 
-\textit{congruence relation}. Compatibility with the algebraic structure operations
-means that algebraic operations applied on equivalent elements will still
-yield equivalent elements.
-
-\begin{defn}
-  The congruence closure $c(R)$ of a relation R is the smallest congruence relation 
-  $R'$ such that $R \subseteq R'$ 
-\end{defn}
-
-\begin{defn}
-  \textbf{} \\
-
-\end{defn}
-
-\section{Implementation}
-\label{sec:impl}
-
-The algorithms and data structures for this paper are implemented in the Go programming 
-language. This implementation makes use of the \textit{Gonum} library for numerical 
-computations. We only import the Gonum libraries for matrices and linear algebra 
-and visual plotting of samples and functions.
-
-Real numbers are implemented with double precision floating point numbers,  
-known as the \texttt{float64} type in the Go programming language.
-
-See Section \ref{sec:headings}.
-
-\subsection{Headings: second level}
-
-\subsubsection{Headings: third level}
-
-\paragraph{Paragraph}
-
-\section{Examples of citations, figures, tables, references}
-\label{sec:others}
-
-\cite{kour2014real,kour2014fast} and see \cite{hadash2018estimate}.
-
-The documentation for \verb+natbib+ may be found at
-\begin{center}
-  \url{http://mirrors.ctan.org/macros/latex/contrib/natbib/natnotes.pdf}
-\end{center}
-Of note is the command \verb+\citet+, which produces citations
-appropriate for use in inline text.  For example,
-\begin{verbatim}
-   \citet{hasselmo} investigated\dots
-\end{verbatim}
-produces
-\begin{quote}
-  Hasselmo, et al.\ (1995) investigated\dots
-\end{quote}
-
-\begin{center}
-  \url{https://www.ctan.org/pkg/booktabs}
-\end{center}
-
-\subsection{Figures}
-See Figure \ref{fig:fig1}. Here is how you add footnotes. \footnote{Sample of the first footnote.}
-
-\begin{figure}
-  \centering
-  \fbox{\rule[-.5cm]{4cm}{4cm} \rule[-.5cm]{4cm}{0cm}}
-  \caption{Sample figure caption.}
-  \label{fig:fig1}
-\end{figure}
-
-\subsection{Tables}
-See awesome Table~\ref{tab:table}.
-
-\begin{table}
- \caption{Sample table title}
-  \centering
-  \begin{tabular}{lll}
-    \toprule
-    \multicolumn{2}{c}{Part}                   \\
-    \cmidrule(r){1-2}
-    Name     & Description     & Size ($\mu$m) \\
-    \midrule
-    Dendrite & Input terminal  & $\sim$100     \\
-    Axon     & Output terminal & $\sim$10      \\
-    Soma     & Cell body       & up to $10^6$  \\
-    \bottomrule
-  \end{tabular}
-  \label{tab:table}
-\end{table}
-
-\subsection{Lists}
-\begin{itemize}
-\item Lorem ipsum dolor sit amet
-\item consectetur adipiscing elit. 
-\item Aliquam dignissim blandit est, in dictum tortor gravida eget. In ac rutrum magna.
-\end{itemize}
-
-\bibliographystyle{unsrt}  
-\bibliography{references}  %%% Remove comment to use the external .bib file (using bibtex).
-%%% and comment out the ``thebibliography'' section.
-
-% 
-% %%% Comment out this section when you \bibliography{references} is enabled.
-% \begin{thebibliography}{1}
-% 
-% \bibitem{kour2014real}
-% George Kour and Raid Saabne.
-% \newblock Real-time segmentation of on-line handwritten arabic script.
-% \newblock In {\em Frontiers in Handwriting Recognition (ICFHR), 2014 14th
-%   International Conference on}, pages 417--422. IEEE, 2014.
-% 
-% \bibitem{kour2014fast}
-% George Kour and Raid Saabne.
-% \newblock Fast classification of handwritten on-line arabic characters.
-% \newblock In {\em Soft Computing and Pattern Recognition (SoCPaR), 2014 6th
-%   International Conference of}, pages 312--318. IEEE, 2014.
-% 
-% \bibitem{hadash2018estimate}
-% Guy Hadash, Einat Kermany, Boaz Carmeli, Ofer Lavi, George Kour, and Alon
-%   Jacovi.
-% \newblock Estimate and replace: A novel approach to integrating deep neural
-%   networks with existing applications.
-% \newblock {\em arXiv preprint arXiv:1804.09028}, 2018.
-% 
-% \end{thebibliography}
-% 
-% 
-\end{document}

### 2020-09-30 paper/paper.tex
preliminary definitions

@@
 \newtheorem*{note}{Note}
-\newtheorem{caso}{Case}
+\newtheorem{case}{Case}
 
@@
 \newcommand{\mrank}[1]{\text{rank}\left(#1\right)}
+\newcommand{\llwb}{\approx_{\mathcal{L}}}
+\newcommand{\langlwa}[2]{\sqmap{#1}^\mathcal{L}_{#2}}
 
@@
   %% \texttt{stariate@ee.mount-sheikh.edu} \\
-  %% \AND
-  %% Coauthor \\
-  %% Affiliation \\
-  %% Address \\
-  %% \texttt{email} \\
-  %% \And
-  %% Coauthor \\
-  %% Affiliation \\
-  %% Address \\
-  %% \texttt{email} \\
-  %% \And
-  %% Coauthor \\
-  %% Affiliation \\
-  %% Address \\
-  %% \texttt{email} \\
 }
@@
 are defined for Linear Weighted Automata (LWAs), under a more abstract coalgebraic perspective.
-For ease of implementation, we focus only on weighted automata over the field of real numbers 
-$\R$.
 
@@
 
+\begin{note}
+  Given two vector spaces $V_1, V_2$ we write $V_1 + V_2$ to denote $\mspan{V_1 \cup V_2}$
+\end{note}
+
 \begin{defn}
   A \textit{weighted automaton} over a field $\K$ and an alphabet $A$ is a triple 
-  $(X,o,t)$ such that $X$ is a finite set of states, $t = (t_a : X \to K)_{a \in A}$
+  $(X,o,t)$ such that $X$ is a finite set of states, 
+  $t = \left\lbrace t_a : X \to \K\right\rbrace_{a \in A}$
   is a set of transition functions indexed over the symbols of the alphabet $A$ and $o : X 
-  \to \K$ is the output function. The transition functions $t_a$ are represented as
-  $\K^{X \times X}$ matrices. $o \in \K^{1 \times X}$ is represented as a row vector.
-  $t_a(v)$ denotes the vector obtained by multiplying the matrix $t_a$ by the column 
-  vector $v  \in \K^{X \times 1}$. $o(v)$ denotes the scalar $s \in \K$ obtained by 
-  dot product of the row vector $o$ with $v \in \K^{X \times 1}$.
+  \to \K$ is the output function. 
   $A^*$ is the set of all words over $A$. $\epsilon$ is the empty word and $aw$ is the
@@
   A function mapping each state vector into its 
-  accepted language, $\sqmap{\cdot}: \K^X \to K^{A^*}$ is defined as follows for every weighted automaton:
+  accepted language, $\sqmap{\cdot}: \K^X \to \K^{A^*}$ is defined as follows for every weighted automaton:
 
@@
   \end{equation*}
+\end{defn}
 
-  Two vectors $v_1, v_2 \in \K^{X\times 1}$ are called language equivalent, denoted with 
-  $v_1 \sim v_2 $ if and only if $ \sqmap
+Two vectors $v_1, v_2 \in \K^{X\times 1}$ are weighted called language equivalent, denoted with 
+$v_1 \sim_l v_2 $ if and only if $ \sqmap
 {v_1} = \sqmap{v_2}$. One can extend the notion of language equivalence to states rather
@@
 of the automaton can be defined as $\sqmap{i}$.
-\end{defn}
 
@@
 
+\begin{defn}
+  The congruence closure $c(R)$ of a relation R is the smallest congruence relation 
+  $R'$ such that $R \subseteq R'$ 
+\end{defn}
+
 An equivalence relation which is compatible with all the operations of
@@
 
+%===================================================================================
+
+We omit the coalgebraic definition for \textit{linear weighted automata} seen in 
+\cite{BONCHI201277} and give a more intuitive definition. 
+In this implementation, we focus only on weighted automata defined over 
+the field of real numbers $\R$.
+
 \begin{defn}
-  The congruence closure $c(R)$ of a relation R is the smallest congruence relation 
-  $R'$ such that $R \subseteq R'$ 
+  A \textit{linear weighted automaton} (in short, LWA) over the field $\K$ and an alphabet $A$
+  is a triple  $L = (V, o, \left\lbrace t_a \right\rbrace_{a \in A})$ 
+  where $V$ is a vector space representing the state space, 
+  $o: V \to \K$ is a linear map associating to each state its output weight,
+  and $t = \left\lbrace t_a = V \times V \right\rbrace_{a \in A}$ is
+  the set of transition functions, represented with liner maps 
+  that for each input $a \in A$ associate the next state, in this case a vector
+  in $V$.
+  As in \cite{boreale2009weighted}, we have that $\dim{(L)} = \dim{(V)}$.
 \end{defn}
 
-% TODO linear weighted automata
+Given a weighted automaton, one can build an isomorphic linear weighted automaton
+by considering the free vector space generated by the set of states $X$ in the WA,
+and by linearizing $o$ and $t$. If $X$ is finite, we can use the same matrices for 
+$t$ and $o$ in both the WA and the corresponding LWA.
+We are only considering a finite number of states and therefore finite dimensional
+vector spaces. Let $n$ be the (finite) number of states in an WA.
+We have that in the corresponding LWA, the transition functions $t_a$ are still
+ represented as
+$\K^{n \times n}$ matrices. $o \in \K^{1 \times n}$ is represented as a row vector.
+$t_a(v)$ denotes the vector obtained by multiplying the matrix $t_a$ by the column 
+vector $v  \in \K^{n \times 1}$. $o(v)$ denotes the scalar $s \in \K$ obtained by 
+dot product of the row vector $o$ with $v \in \K^{n \times 1}$.
+
 \begin{defn}
-  \textbf{} \\
+  The language recognized by a vector $v \in V$ of an LWA $(V,o,t)$ is defined
+  for all words $w \in A*$ as $\langlwa{v}{V}(w) = o(v_n)$ where $v_n$ is the 
+  vector reached from $v$ through the composition of the transition functions
+  corresponding to the words in $w$.
+
+  \begin{equation*}
+    \begin{aligned}
+      \langlwa{v}{V}(w) = \begin{cases}
+        o(v) & \text{if } w = \epsilon \\ 
+        \langlwa{t_a(v)}{V}(w') & \text{if } w = aw' 
+      \end{cases}
+    \end{aligned}
+  \end{equation*}
 
@@
 
+We define $\llwb$ as the behavioral equivalence for a given LWA $(V, o, t)$ as
+
+\begin{equation}
+  \forall v_1, v_2 \in V, \; v_1 \llwb v_2 \iff \langlwa{v_1}{V} = \langlwa{v_2}{V}
+\end{equation}
+
+\begin{lem}
+  $\llwb$ coincides with $\sim_l$:
+
+  Let $(X, o, t)$ be a WA and $(\K^X, o^\sharp, t^\sharp)$ the corresponding linear 
+  weighted automaton. Then $\forall x \in X, \;\; \sqmap{x} = \langlwa{x}{\K^X}$
+\end{lem}
+
+\begin{proof}
+  Proved in section 3.2 of \cite{BONCHI201277}
+\end{proof}
+
+%=========================================================================
+
+\section{The Problem}
+In this work, we
+
 \section{Algorithms}
 
-\subsection{Forward Partition Refinement Algorithm for the Largest Weighted Bisimulation}
-\label{sec:algo2}
-
-The algorithm is defined by the iterative method:
-\begin{eqnarray}
-  R_0 = \mker{o}, & \quad & R_{i+1} = R_i \cap \bigcap_{a \in A} t_a(R_i)^{-1} \label{fwd} 
-\end{eqnarray}
-The algorithm stops when $R_{j+1} = R_j$. Proof is available in section 4.1 of \cite{BONCHI201277}
+We now recall the backwards algorithm for computing $\llwb$ defined in \cite{BONCHI201277}.
 
@@
 
+%\subsection{Forward Partition Refinement Algorithm for the Largest Weighted Bisimulation}
+%\label{sec:algo2}
+%
+%The algorithm is defined by the iterative method:
+%\begin{eqnarray}
+%  R_0 = \mker{o}, & \quad & R_{i+1} = R_i \cap \bigcap_{a \in A} t_a(R_i)^{-1} \label{fwd} 
+%\end{eqnarray}
+%The algorithm stops when $R_{j+1} = R_j$. Proof is available in section 4.1 of \cite{BONCHI201277}
+
 \section{Implementation}
@@
 The algorithms and data structures for this paper are implemented in the Go programming 
-language. This implementation makes use of the \textit{Gonum} library for numerical 
+language. This implementation makes use of the Gonum library for numerical 
 computations. We only import the Gonum libraries for matrices and linear algebra 
@@
 See Section \ref{sec:headings}.
-
-\subsection{Input file format}
-
-\subsection{Calculating the intersection of linear subspaces}
-
-To compute algorithm (\ref{fwd}), we need an efficient way to calculate a basis 
-of the intersection of an arbitrary number of linear subspaces of $\R^n$, given their 
-spanning sets.
-
-\begin{defn}
-  \textbf{Algorithm for a basis of the intersection of subspaces} \\
-  Defined in file \texttt{lin/intersect.go}.
-  Let's consider two linear subspaces of 
-  $\R^n$ defined with their spanning sets of column vectors: 
-  $U = \spanset{u_1, u_2, \hdots, u_p}$ 
-  and $W = \spanset{w_1, w_2, \hdots, w_k}$.
-  We create the block matrix $A \in \R^{n \times (p+k)}$ defined as follows:
-
-  \begin{equation*}
-    \begin{aligned}
-      a_{ij} = \begin{cases}
-        (u_j)_i & \text{for  } j = 1, \hdots, p \quad i = 1, \hdots, n \\
-        (w_{j-p})_i & \text{for  } j = p+1, \hdots, p+k \quad i = 1, \hdots, n \\
-      \end{cases}
-    \end{aligned}
-  \end{equation*}
-
-  The matrix will have the form
-  \begin{equation*}
-    A = \begin{bmatrix}
-      \vvvec{}{u_1}{\,}, \hdots \vvvec{}{u_p}{\,}, 
-      \vvvec{}{w_1}{\,}, \hdots \vvvec{}{w_k}{\,} 
-    \end{bmatrix}
-  \end{equation*}
-
-  It follows that $\mker{A}$ is a basis for $U \cap W$. This can be done with an 
-  arbitrary finite number of subspaces of $\R^n$, as long as their spanning set is 
-  \textit{finite}.
-\end{defn}
 
@@
 \begin{prop}
-  Within finite dimensional spaces, the annihilator $W^0$ is isomorphic
-  to the orthogonal complement.  
+  If $V$ is a finite dimensional vector space defined with an inner product
+  $\langle \cdot , \cdot \rangle$ and $W$ is a subspace of $V$
+  then the image of the annhilitaor $W^0$ through the linear 
+  isomorphism $\varphi: V^* \to V$ induced by the inner product, 
+  is the orthogonal of $W$ with respect to the said inner product.
 \end{prop}
@@
 \begin{proof}
-  Let $V$ be an inner product space over the field $\K$ with an inner product
-  $\langle \cdot , \cdot \rangle : V \times V \to \K$ defined. 
-  With the inner product defined on $V$, every linear functional can be 
-  represented with a vector: if $\xi : V \to \K$ is a functional, $c$ 
-  is an unique vector such that $\forall x \in V \: . \: \langle c, x \rangle = \xi(x)$.
-  The map associating every $\xi \in V^*$ with the vector $c \in V$ that corresponds 
-  to $\xi$ with respect to the inner product is a linear isomorphism. Through this 
-  isomorphism, the annihilator $W^0$ becomes $W^\perp$.
+  Let $V$ be an inner product space over the field $\K$ with an inner product defined as
+  $\langle \cdot , \cdot \rangle : V \times V \to \K$. 
+  Every linear functional can be 
+  represented with a vector. Let $\xi : V \to \K$ be a functional, 
+  $\xi \in  W^0$. Because $\xi(w)=0 \quad \forall  w \in W$, 
+  if $v$ represents $\xi$ we have that $(v, w)=\xi(w)=0$ for all $w \in W$. 
+  We obtain that $\varphi(W^0) \subseteq W^{\perp}$.
+  If $v \in W^\perp$  
+  then the functional $x \mapsto (v, x)$ cancels over $W$ 
+  (by the definition of orthogonality).
 \end{proof}

### 2020-10-01 paper/paper.tex
badly working partition refinement algorithm

@@
 of two vectors (states) for a given weighted automata by using an additional
-data structure representing a congruence relationship between states.
-We then compare the two algorithms results to verify their correctness
-on randomly generated automata samples, and provide some runtime statistics.
-
+data structure representing a congruence relation between states.
+We then compare results of the two algorithms to verify their correctness
+and performance on randomly generated samples. 
+We finally provide comparison of runtime statistics and suggest 
+which of the two algorithm is a the best choice for some usage cases.
 \end{abstract}
@@
 \begin{eqnarray}
-  R_0 = \mker{o}^0, & \quad & R_{i+1} = R_i + \sum_{a \in A} t_a(R_i)^t \label{back} 
+  R_0 = \mker{o}^0, & \quad & R_{i+1} = R_i + \sum_{a \in A} t_a^t(R_i) \label{back} 
 \end{eqnarray}

### 2020-10-01 automata/automata.go

@@ -15,9 +15,10 @@ import (
 func errRead(msg string) error { return errors.New("could not read automaton:" + msg) }
 
 type Automaton struct {
-	A []string              // The input alphabet
-	T map[string]*mat.Dense // Transition matrices are maps from input strings to dense real valued matrices
-	O *mat.VecDense         // Output vector uses a dense real valued vector
+	A   []string              // The input alphabet
+	T   map[string]*mat.Dense // Transition matrices are maps from input strings to dense real valued matrices
+	O   *mat.VecDense         // Output vector uses a dense real valued vector
+	Dim int                   // Number of states/dimension of vector space V in LWA
 }
 
 // applies the output function to a given state vector (o * v)
@@ -28,5 +29,17 @@ func (a Automaton) GetOutput(v *mat.VecDense) float64 {
 func (a Automaton) ApplyTransition(s string, v *mat.VecDense) *mat.VecDense {
 	res := mat.VecDenseCopyOf(v)
 	res.MulVec(a.T[s], v)
+	return res
+}
+
+func (a Automaton) ApplyTransposeTransition(s string, v *mat.VecDense) *mat.VecDense {
+	res := mat.VecDenseCopyOf(v)
+	res.MulVec(a.T[s].T(), v)
+	return res
+}
+
+func (a Automaton) ApplyTransposeTransitionBasis(s string, b *mat.Dense) *mat.Dense {
+	res := mat.DenseCopyOf(b)
+	res.Mul(a.T[s].T(), b)
 	return res
 }

### 2020-10-01 automata/backwards_test.go

@@ -1,0 +1,44 @@
+package automata
+
+import (
+	"fmt"
+	"testing"
+
+	"github.com/0x0f0f0f/lwa-techniques/lin"
+	"gonum.org/v1/gonum/mat"
+)
+
+func TestBackwardsPartitionRefinement(t *testing.T) {
+	a := Automaton{
+		A: []string{"a", "b"},
+		T: map[string]*mat.Dense{
+			"a": mat.NewDense(3, 3, []float64{
+				1, 0.33333333333333333333333333333333333333, 1,
+				0, 0, 0,
+				0, 0, 0,
+			}),
+			"b": mat.NewDense(3, 3, []float64{
+				1, 0, 0,
+				0, 0, 3,
+				0, 0.33333333333333333333333333333333333333, 0,
+			}),
+		},
+		Dim: 3,
+		O:   mat.NewVecDense(3, []float64{2, 1, 1}),
+	}
+
+	b := a.BackwardsPartitionRefinement()
+	lin.PrintMat(b)
+
+	v1 := mat.NewVecDense(3, []float64{1, 0, 0})
+	v2 := mat.NewVecDense(3, []float64{0, 1.5, 0.5})
+
+	res := a.BPREquivalence(v1, v2, b)
+	fmt.Println(res)
+
+	v1 = mat.NewVecDense(3, []float64{1, 0, 1})
+	v2 = mat.NewVecDense(3, []float64{2, 4, 0.5})
+
+	res = a.BPREquivalence(v1, v2, b)
+	fmt.Println(res)
+}

### 2020-10-01 automata/io.go

@@ -110,6 +110,7 @@ func ReadAutomaton(r io.Reader, prompt bool) (*Automaton, error) {
 	if err != nil {
 		return nil, err
 	}
+	a.Dim = numStates
 
 	// Read the output vector
 	if prompt {
@@ -153,12 +154,12 @@ func (a Automaton) FancyString() string {
 
 	// Print the output vector
 	fo := mat.Formatted(a.O, mat.Prefix("      "), mat.Squeeze())
-	buf.WriteString(fmt.Sprintf("o = %v\n\n", fo))
+	buf.WriteString(fmt.Sprintf("o = %.5g\n\n", fo))
 
 	// Print the matrices
 	for sym, m := range a.T {
 		fo := mat.Formatted(m, mat.Prefix("      "), mat.Squeeze())
-		buf.WriteString(fmt.Sprintf("T_%s = %v\n\n", sym, fo))
+		buf.WriteString(fmt.Sprintf("T_%s = %.5g\n\n", sym, fo))
 	}
 
 	return buf.String()
@@ -174,12 +175,12 @@ func (a Automaton) String() string {
 
 	// Print the output vector
 	fo := mat.Formatted(a.O, mat.Squeeze(), mat.FormatMATLAB())
-	buf.WriteString(fmt.Sprintf("o = %v\n\n", fo))
+	buf.WriteString(fmt.Sprintf("o = %.5g\n\n", fo))
 
 	// Print the matrices
 	for sym, m := range a.T {
 		fo := mat.Formatted(m, mat.Prefix("      "), mat.Squeeze())
-		buf.WriteString(fmt.Sprintf("T_%s = %v\n\n", sym, fo))
+		buf.WriteString(fmt.Sprintf("T_%s = %.5g\n\n", sym, fo))
 	}
 
 	return buf.String()

### 2020-10-01 lin/independent.go

@@ -1,0 +1,76 @@
+package lin
+
+import (
+	"math"
+
+	"gonum.org/v1/gonum/mat"
+)
+
+// find which column vectors in m are linearly independent and
+// return them
+func Independent(a *mat.Dense) *mat.Dense {
+	var qr mat.QR
+	qr.Factorize(a)
+	m, n := a.Dims()
+	r := mat.NewDense(m, m, nil)
+	r.Reset()
+	qr.RTo(r)
+
+	rank := 0
+	cols := []int{}
+	for k := 0; k < n; k++ {
+		if math.Abs(r.At(k, k)) > tol {
+			rank++
+			cols = append(cols, k)
+		}
+	}
+
+	res := mat.NewDense(m, rank, nil)
+
+	j := 0
+	for _, k := range cols {
+		for i := 0; i < m; i++ {
+			res.Set(i, j, a.At(i, k))
+		}
+		j++
+	}
+
+	return res
+}
+
+// find which column vectors in m are linearly independent and
+// return them
+func IndependentGauss(a *mat.Dense) *mat.Dense {
+	m, n := a.Dims()
+	ab := a.Grow(0, 1)
+
+	red, _, err := GaussPartial(ab)
+	if err != nil {
+		panic(err)
+	}
+
+	red = red.Slice(0, m, 0, n).(*mat.Dense)
+
+	rank := 0
+	cols := []int{}
+	for k := 0; k < n; k++ {
+		if k < m {
+			if math.Abs(red.At(k, k)) > tol {
+				rank++
+				cols = append(cols, k)
+			}
+		}
+	}
+
+	res := mat.NewDense(m, rank, nil)
+
+	j := 0
+	for _, k := range cols {
+		for i := 0; i < m; i++ {
+			res.Set(i, j, a.At(i, k))
+		}
+		j++
+	}
+
+	return res
+}

### 2020-10-01 lin/util.go

@@ -26,7 +26,10 @@ func EyeDense(n int) *mat.Dense {
 	return a
 }
 
-func PrintMat(a mat.Matrix) {
-	fmt.Printf("%.90g\n", mat.Formatted(a, mat.Squeeze()))
+func StringMat(a mat.Matrix) string {
+	return fmt.Sprintf("%.5g", mat.Formatted(a, mat.Squeeze()))
+}
 
+func PrintMat(a mat.Matrix) {
+	fmt.Println(StringMat(a))
 }

### 2020-10-01 main.go

@@ -2,8 +2,9 @@ package main
 
 import (
 	"fmt"
+	"os"
 
-	"github.com/0x0f0f0f/lwa-techniques/lin"
+	"github.com/0x0f0f0f/lwa-techniques/automata"
 	"gonum.org/v1/gonum/mat"
 )
 
@@ -30,22 +31,14 @@ func main() {
 	fmt.Println(r.PairIsInCongruenceClosure(p2))
 	*/
 
-	// a, err := aut.ReadAutomaton(os.Stdin, true)
-	// check(err)
-	// fmt.Println(a.String())
+	a, err := automata.ReadAutomaton(os.Stdin, true)
+	check(err)
+	fmt.Println(a.String())
 
-	// v1 := mat.NewVecDense(3, []float64{0, 1, 0})
-	// v2 := mat.NewVecDense(3, []float64{0, 0, 1})
+	v1 := mat.NewVecDense(3, []float64{0, 1, 0})
+	v2 := mat.NewVecDense(3, []float64{0, 0, 1})
 
-	// res, err := a.HKC(v1, v2)
-	// fmt.Println(res, err)
-
-	w1 := mat.NewVecDense(3, []float64{1, 7, 2})
-	w2 := mat.NewVecDense(3, []float64{-2, 3, 1})
-	//assert.Equal()
-
-	wperp := lin.Complement(3, []*mat.VecDense{w1, w2})
-
-	fmt.Println(len(wperp))
+	res, err := a.HKC(v1, v2)
+	fmt.Println(res, err)
 
 }

### 2020-10-02 paper/notused.tex
working on experimental data

@@
 \end{defn}
+
+\begin{defn}
+  \textbf{Algorithm for computing the null space of a vector subspace} \\
+  The algorithm implementation can be found in file \texttt{lin/nullspace.go}; it is
+  adapted from \cite{scipy/ranknullspace}. 
+  Let's consider the singular value decomposition of a matrix $A \in \R^{m \times n}$:
+
+  \begin{equation*}
+    \begin{aligned}
+      A = U \Sigma V^T & \quad & \Sigma = \diag{\sigma_1, \sigma_2, \hdots, \sigma_{\min(m, n)}  } 
+       & \quad &  U \in \R^{m \times m} & \quad & V \in \R^{n \times n}
+    \end{aligned}
+  \end{equation*}
+
+  Where $V$ and $U$ are orthogonal and the singular values are ordered: $\sigma_1 \geq \sigma_2 \geq \hdots \geq \sigma_{\min(m,n)} \geq 0$.
+  It follows that $\mrank{A}$ is equal to the number of nonzero singular values and
+  a basis of the (right) null space of $A$ is the spanning set of the columns of V
+  corresponding to singular values equal to zero. 
+\end{defn}
+
+\begin{exmp}
+  First, we show a shorter Python 
+  implementation of the algorithm to compute the nullspace, using the 
+  \textit{SciPy} library \cite{scipy/ranknullspace}:
+
+  \lstinputlisting[language=python]{nullspace.py}
+
+    The Go implementation is quite longer:
+
+    \lstinputlisting[language=go]{../lin/nullspace.go}
+
+\end{exmp}

### 2020-10-02 paper/paper.tex
working on experimental data

@@
 \begin{proof}
-  Proved in section 3.2 of \cite{BONCHI201277}
+  Proved in section 3.3 of \cite{BONCHI201277}
 \end{proof}
+
+Language equivalence can be expressed in terms of linear weighted bisimulations (LWBs for short).
+Differently from weighted bisimulations, LWBs can be seen both as relations and as subspaces.
+The subspace representation of LWBs is used in the backwards partition refinement algorithm 
+implemented in \cite{BONCHI201277} and in this work.
 
@@
 \section{The Problem}
-In this work, we
+We want to check if two vectors are language equivalent $v_1 \sim_l v_2$.
 
@@
 Where $\mker{o}^0$ is an annihilator.
-The algorithm stops when $R_{j+1} = R_j$. Proof is available in section 4.2 of \cite{BONCHI201277}
+The algorithm stops when $R_{j+1} = R_j$. An index $j \leq \dim(V)$ is 
+guaranteed to exist, such that the algorithms terminates at step $j$.
+It follows that $\llwb \equiv R_j^0$.
+Proof is available in section 4.2 of \cite{BONCHI201277}
 
@@
 
-See Section \ref{sec:headings}.
+\begin{defn}
+  \textbf{Applications of SVD} \\
 
-\begin{defn}
-  \textbf{Algorithm for computing the null space of a vector subspace} \\
-  The algorithm implementation can be found in file \texttt{lin/nullspace.go}; it is
-  adapted from \cite{scipy/ranknullspace}. 
   Let's consider the singular value decomposition of a matrix $A \in \R^{m \times n}$:
@@
     \begin{aligned}
-      A = U \Sigma V^T & \quad & \Sigma = \diag{\sigma_1, \sigma_2, \hdots, \sigma_{\min(m, n)}  } 
+      A = U \Sigma V^T & \quad & \Sigma = \diag{\sigma_1, \sigma_2, \hdots, \sigma_r  } 
        & \quad &  U \in \R^{m \times m} & \quad & V \in \R^{n \times n}
@@
 
-  Where $V$ and $U$ are orthogonal and the singular values are ordered: $\sigma_1 \geq \sigma_2 \geq \hdots \geq \sigma_{\min(m,n)} \geq 0$.
-  It follows that $\mrank{A}$ is equal to the number of nonzero singular values and
-  a basis of the (right) null space of $A$ is the spanning set of the columns of V
-  corresponding to singular values equal to zero. 
+  Where $V$ and $U$ are orthogonal and the singular values are ordered: $\sigma_1 \geq \sigma_2 \geq \hdots \geq \sigma_r \geq 0$.
+  It follows that $\mrank{A}$ is equal to the number of nonzero singular values, and
+  as explained in \cite{svd}:
+
+  \begin{enumerate}
+    \item  $\mrank{A} = \mrank{\Sigma} = r$
+    \item The column space of $A$ is spanned by the first $r$ columns of $U$.
+    \item The null space of $A$ is spanned by the last $n − r$ columns of $V$.
+    \item The row space of $A$ is spanned by the first $r$ columns of $V$.
+    \item The null space of $A^T$ is spanned by the last $m − r$ columns of $U$.
+  \end{enumerate}
+
+  Of our interest, are only the computation of the null space and columns space.
+  The implementation, applying SVD, 
+  can be found in files \texttt{lin/colspace.go} and \texttt{lin/nullspace.go}.
 \end{defn}
 
-\begin{exmp}
-  First, we show a shorter Python 
-  implementation of the algorithm to compute the nullspace, using the 
-  \textit{SciPy} library \cite{scipy/ranknullspace}:
+\subsection{Implementing the backwards partition refinement algorithm}
 
-  \lstinputlisting[language=python]{nullspace.py}
-
-    The Go implementation is quite longer:
-
-    \lstinputlisting[language=go]{../lin/nullspace.go}
-
-\end{exmp}
-
-\subsection{Implementing the backwards algorithm}
-
-In \cite{BONCHI201277}, for the first step of the iterative backwards 
-algorithm to compute the largest linear weighted bisimulation, we 
-need to compute $\mker{o}^0$. If $V$ is a vector space and $W$ is a
+To compute $\llwb$ at the last step of the algorithm,
+we need to compute $R_j^0$.
+If $V$ is a vector space and $W$ is a
 subspace of $W$, the annihilator of $W$, respectively $W^0$ is 
@@
 
-\subsubsection{Headings: third level}
-
-\paragraph{Paragraph}
-
-\section{Examples of citations, figures, tables, references}
-\label{sec:others}
-
-\cite{kour2014real,kour2014fast} and see \cite{hadash2018estimate}.
-
-The documentation for \verb+natbib+ may be found at
-\begin{center}
-  \url{http://mirrors.ctan.org/macros/latex/contrib/natbib/natnotes.pdf}
-\end{center}
-Of note is the command \verb+\citet+, which produces citations
-appropriate for use in inline text.  For example,
-\begin{verbatim}
-   \citet{hasselmo} investigated\dots
-\end{verbatim}
-produces
-\begin{quote}
-  Hasselmo, et al.\ (1995) investigated\dots
-\end{quote}
-
-\begin{center}
-  \url{https://www.ctan.org/pkg/booktabs}
-\end{center}
-
-\subsection{Figures}
-See Figure \ref{fig:fig1}. Here is how you add footnotes. \footnote{Sample of the first footnote.}
+%=============================================================
 
@@
 \end{figure}
-
-\subsection{Tables}
-See awesome Table~\ref{tab:table}.
 

### 2020-10-02 automata/automata.go

@@ -4,7 +4,9 @@ package automata
 
 import (
 	"errors"
+	"math/rand"
 
+	"github.com/0x0f0f0f/lwa-techniques/lin"
 	"gonum.org/v1/gonum/mat"
 )
 
@@ -15,10 +17,37 @@ import (
 func errRead(msg string) error { return errors.New("could not read automaton:" + msg) }
 
 type Automaton struct {
-	A   []string              // The input alphabet
-	T   map[string]*mat.Dense // Transition matrices are maps from input strings to dense real valued matrices
-	O   *mat.VecDense         // Output vector uses a dense real valued vector
-	Dim int                   // Number of states/dimension of vector space V in LWA
+	A        []string              // The input alphabet
+	T        map[string]*mat.Dense // Transition matrices are maps from input strings to dense real valued matrices
+	O        *mat.VecDense         // Output vector uses a dense real valued vector
+	Dim      int                   // Number of states/dimension of vector space V in LWA
+	LLWB     *mat.Dense            // Col(LLWB) is a basis of the largest linear weighted bisimulation, a binary linear relation. vRw iff (v-w) in Ker(R)
+	LLWBperp *mat.Dense
+}
+
+func RandAutomaton(syms, states int) Automaton {
+	// create the alphabet
+	A := make([]string, syms)
+	if syms <= 57 {
+		for i := 0; i < syms; i++ {
+			A[i] = string(rune(i + 65))
+		}
+	}
+
+	T := map[string]*mat.Dense{}
+	for _, sym := range A {
+		T[sym] = lin.RandIntDense(states, 2)
+		lin.PokeHoles(T[sym], rand.Intn((states*states)/2))
+	}
+
+	aut := Automaton{
+		A:   A,
+		T:   T,
+		O:   lin.RandVec(states),
+		Dim: states,
+	}
+
+	return aut
 }
 
 // applies the output function to a given state vector (o * v)

### 2020-10-02 automata/backwards.go

@@ -4,50 +4,64 @@
 package automata
 
 import (
-	"fmt"
+	"log"
 
 	"github.com/0x0f0f0f/lwa-techniques/lin"
 	"gonum.org/v1/gonum/mat"
 )
 
-func (a Automaton) BackwardsPartitionRefinement() *mat.Dense {
+// compute and store a basis for the largest linear weighted bisimulation of
+// the linear weighted automaton
+func (a *Automaton) BackwardsPartitionRefinement() {
 	// i = 0
-
-	fmt.Println(a.FancyString())
-
 	lastBasis := mat.NewDense(a.Dim, 1, a.O.RawVector().Data)
 	currBasis := lastBasis
 	// index of the column of last basis that has already been computed
-	lastIndex := 0
+	// lastIndex := 0
 
-	for i := 1; i < a.Dim; i++ {
-		fmt.Printf("==============================\nstep %d\n", i)
-		// Σ_{a \in A} T
+	for i := 1; i <= a.Dim; i++ {
+		// fmt.Printf("==============================\nstep %d\n", i)
+		// Σ_{a \in A} T_a^t(R_i)
 		for _, sym := range a.A {
-			_, n := lastBasis.Dims()
-			toCompute := lastBasis.Slice(0, a.Dim, lastIndex, n).(*mat.Dense)
-			newBasis := a.ApplyTransposeTransitionBasis(sym, toCompute)
-			fmt.Printf("T_(%s)^t of \n%s \n = \n", sym, lin.StringMat(toCompute))
-			lin.PrintMat(newBasis)
+			//_, n := lastBasis.Dims()
+			// if lastIndex == n {
+			// break
+			// }
+			// toCompute := lastBasis.Slice(0, a.Dim, lastIndex, n).(*mat.Dense)
+			newBasis := a.ApplyTransposeTransitionBasis(sym, lastBasis)
+			// fmt.Printf("T_(%s)^t of \n%s \n = \n", sym, lin.StringMat(lastBasis))
+			// lin.PrintMat(newBasis)
 			currBasis = lin.Union(currBasis, newBasis)
-			fmt.Println("curr basis = ")
-			lin.PrintMat(currBasis)
-			currBasis = lin.IndependentGauss(currBasis)
-			fmt.Printf("independent columns of b = \n%s\n", lin.StringMat(currBasis))
+			// fmt.Println("curr basis = ")
[…]

### 2020-10-02 automata/hkc.go

@@ -1,8 +1,6 @@
 package automata
 
 import (
-	"fmt"
-
 	"gonum.org/v1/gonum/mat"
 )
 
@@ -16,36 +14,35 @@ func (a Automaton) HKC(v1, v2 *mat.VecDense) (bool, error) {
 		return false, err
 	}
 
-	fmt.Println("p = ", p)
+	//fmt.Println("p = ", p)
 
 	// insert (v1, v2) into the todo list
 	todo.Push(*p)
 
 	i := 0
 	for !todo.Empty() {
-		fmt.Println("Step", i)
-		fmt.Println("todo =", todo)
+		//fmt.Println("Step", i)
+		//fmt.Println("todo =", todo)
 
 		// extract (v1', v2') from todo
 		q, err := todo.Pop()
 		if err != nil {
 			return false, err
 		}
-		fmt.Println("Extracted pair is", q)
-
-		fmt.Println("todo =", todo)
+		//fmt.Println("Extracted pair is", q)
+		//fmt.Println("todo =", todo)
 
 		if rel.PairIsInCongruenceClosure(q) {
-			fmt.Println("(v1', v2') \\in c(R)")
+			//fmt.Println("(v1', v2') \\in c(R)")
 			continue
 		}
 
 		o1 := a.GetOutput(q.Left)
 		o2 := a.GetOutput(q.Right)
-		fmt.Println("o(v1) =", o1)
-		fmt.Println("o(v2) =", o2)
+		//fmt.Println("o(v1) =", o1)
+		//fmt.Println("o(v2) =", o2)
 		if o1 != o2 {
-			fmt.Println("o(v1) =/= o(v2)")
+			//fmt.Println("o(v1) =/= o(v2)")
 			return false, nil
 		}
 
@@ -64,7 +61,7 @@ func (a Automaton) HKC(v1, v2 *mat.VecDense) (bool, error) {
 		// insert (v1', v2') into R
 		rel.Add(q)
 
-		fmt.Println("R = ", rel)
+		//fmt.Println("R = ", rel)
 		i++
 	}
 

### 2020-10-02 lin/util.go

@@ -1,7 +1,9 @@
 package lin
 
 import (
+	"errors"
 	"fmt"
+	"math/rand"
 
 	"gonum.org/v1/gonum/mat"
 )
@@ -26,10 +28,90 @@ func EyeDense(n int) *mat.Dense {
 	return a
 }
 
+// generate a normally distributed random n*n matrix
+func RandDense(n int) *mat.Dense {
+	data := make([]float64, n*n)
+	for i := range data {
+		data[i] = rand.NormFloat64()
+	}
+	return mat.NewDense(n, n, data)
+}
+
+// generate a normally distributed random integer n*n matrix
+func RandIntDense(n, max int) *mat.Dense {
+	data := make([]float64, n*n)
+	for i := range data {
+		data[i] = float64(rand.Intn(max))
+	}
+	return mat.NewDense(n, n, data)
+}
+
+// set z randomly chosen values in the matrix to 0
+func PokeHoles(a *mat.Dense, z int) {
+	m, n := a.Dims()
+	for k := 0; k < z; k++ {
+		i := rand.Intn(m)
+		j := rand.Intn(n)
+
+		a.Set(i, j, 0)
+	}
+}
+
+// performs a linear combination of the columns of v and the coefficients
+// in the elements of b
+func LinearCombination(a *mat.Dense, b *mat.VecDense) *mat.VecDense {
+	m, n := a.Dims()
+	bn, _ := b.Dims()
+
+	if n != bn {
+		panic(errors.New("mat-vector dimension mismatch"))
+	}
+
+	res := mat.NewVecDense(m, nil)
+
+	for i := 0; i < n; i++ {
+		res.AddScaledVec(res, b.AtVec(i), a.ColView(i))
+	}
+
+	return res
+}
+
+// generate a normally distributed random n*1 vector
+func RandVec(n int) *mat.VecDense {
+	data := make([]float64, n)
+	for i := range data {
+		data[i] = rand.NormFloat64()
+	}
+	return mat.NewVecDense(n, data)
+}
+
+// generate a zero matrix of size m*n
+func ZeroDense(m, n int) *mat.Dense {
+	return mat.NewDense(m, n, nil)
+}
+
+// generate a zero vector of length n
+func ZeroVec(n int) *mat.VecDense {
+	return mat.NewVecDense(n, nil)
+}
+
+// string representation of a matrix
[…]

### 2020-10-02 main.go

@@ -2,9 +2,11 @@ package main
 
 import (
 	"fmt"
-	"os"
+	"math/rand"
+	"time"
 
 	"github.com/0x0f0f0f/lwa-techniques/automata"
+	"github.com/0x0f0f0f/lwa-techniques/lin"
 	"gonum.org/v1/gonum/mat"
 )
 
@@ -14,31 +16,47 @@ func check(err error) {
 	}
 }
 
+func randTest(dim, numSyms, numSamples int) bool {
+	az := automata.RandAutomaton(numSyms, dim)
+
+	az.BackwardsPartitionRefinement()
+
+	if mat.Equal(az.LLWB, mat.NewDense(dim, 1, nil)) {
+		return false
+	}
+
+	fmt.Println(az)
+	lin.PrintMat(az.LLWB)
+	_, dimLLWB := az.LLWB.Dims()
+
+	samples := make([]*mat.VecDense, numSamples)
+	for i := range samples {
+		samples[i] = lin.LinearCombination(az.LLWB, lin.RandVec(dimLLWB))
+	}
+
+	for i := range samples {
+		j := rand.Intn(numSamples)
+		for j == i {
+			j = rand.Intn(numSamples)
+		}
+		resBPR := az.BPREquivalence(samples[i], samples[j])
+		resHKC, _ := az.HKC(samples[i], samples[j])
+
+		if resBPR == resHKC {
+			fmt.Println(resBPR, resHKC)
+			lin.PrintMat(samples[i])
+			lin.PrintMat(samples[j])
+		}
+	}
+	return false
+}
+
 func main() {
-	/* 		v1 := mat.NewVecDense(3, []float64{1.0, 2.0, 3.0})
-	   	v2 := mat.NewVecDense(3, []float64{3.0, 2.0, 1.0})
-	   	p1, err := aut.NewPair(v1, v2)
-	   	check(err)
+	//defer profile.Start(profile.MemProfile).Stop()
 
-	   	v3 := mat.NewVecDense(3, []float64{4.0, 3.0, 5.0})
-	   	v4 := mat.NewVecDense(3, []float64{6.0, 3.0, 3.0})
-	   	p2, err := aut.NewPair(v3, v4)
-	   	check(err)
+	rand.Seed(time.Now().UnixNano())
 
-	r := aut.NewRelation()
-	r.Add(p1)
-
-	fmt.Println(r.PairIsInCongruenceClosure(p2))
[…]

### 2020-10-05 paper/paper.tex
working on tex, added random tests

@@
 \usepackage{amsmath}
+\usepackage{eucal}
 \usepackage{mathtools}
@@
 \usepackage{lstlistings-golang}
+\usepackage{caption}
 
@@
 \usepackage{fancyvrb}
-\usepackage{listings}
+\usepackage{listingsutf8}
 \usepackage{textcomp}
@@
 \lstset{
+  numbers=left,
+  frame=tb,
+  commentstyle=\color{mygreen},
 	basicstyle=\ttfamily,
 	columns=fullflexible,
-	keepspaces=true,
+  keepspaces=false,
+  showstringspaces=false,
+  extendedchars=true,
+  mathescape=true,
+  breaklines, breakatwhitespace=true
 }
 
-\title{Go Implementation of Up-To Techniques for Equivalence of Weighted Languages}
+\title{DRAFT: Go Implementation of Up-To Techniques for Equivalence of Weighted Languages}
 
@@
 
+\definecolor{mygreen}{rgb}{0,0.6,0}
+
 \begin{document}
@@
 equivalence in finite state weighted automata (WAs). The first, a
-linear partition refinement algorithm, calculates the largest linear weighted bisimulation
+linear partition refinement algorithm, computes the largest linear weighted bisimulation
 for any given LWA (Linear Weighted Automaton) through an iterative method, 
@@
 of two vectors (states) for a given weighted automata by using an additional
-data structure representing a congruence relation between states.
+data structure representing a congruence relation.
 We then compare results of the two algorithms to verify their correctness
@@
 We finally provide comparison of runtime statistics and suggest 
-which of the two algorithm is a the best choice for some usage cases.
+which of the two algorithm is the best choice for different usage cases.
 \end{abstract}
@@
 
-\section{Introduction}
-\label{sec:intro}
+\input{sections/introduction.tex}
 
-In \cite{DBLP:journals/corr/Bonchi0K17}, up-to techniques are defined for
-weighted systems over arbitrary semirings, while in \cite{BONCHI201277}, up-to techniques
-are defined for Linear Weighted Automata (LWAs), under a more abstract coalgebraic perspective.
-
-\section{Preliminaries and Notation}
-\label{sec:notation}
-
-\begin{note}
-  Given two vector spaces $V_1, V_2$ we write $V_1 + V_2$ to denote $\mspan{V_1 \cup V_2}$
-\end{note}
-
-\begin{defn}
-  A \textit{weighted automaton} over a field $\K$ and an alphabet $A$ is a triple 
-  $(X,o,t)$ such that $X$ is a finite set of states, 
-  $t = \left\lbrace t_a : X \to \K\right\rbrace_{a \in A}$
-  is a set of transition functions indexed over the symbols of the alphabet $A$ and $o : X 
-  \to \K$ is the output function. 
-  $A^*$ is the set of all words over $A$. $\epsilon$ is the empty word and $aw$ is the
-  concatenation of a symbol $a$ to the word $w \in A^*$.
-  A weighted language is a function $\psi: A^* \to \K$.
-  A function mapping each state vector into its 
-  accepted language, $\sqmap{\cdot}: \K^X \to \K^{A^*}$ is defined as follows for every weighted automaton:
-
-  \begin{equation*}
-    \begin{aligned}
-      \forall v \in \K^X, a \in A, w \in A^* \quad \quad
-      \sqmap{v}(\epsilon) = o(v) \quad \quad
-      \sqmap{v}(aw) = \sqmap{t_a(v)}(w)  
-    \end{aligned}
-  \end{equation*}
-\end{defn}
-
-Two vectors $v_1, v_2 \in \K^{X\times 1}$ are weighted called language equivalent, denoted with 
-$v_1 \sim_l v_2 $ if and only if $ \sqmap
-{v_1} = \sqmap{v_2}$. One can extend the notion of language equivalence to states rather
-than for vectors by assigning to each state $x \in X$ the corresponding unit vector 
-$e_x \in \K^X$. When given an initial state $i$ for a weighted automaton, the language 
-of the automaton can be defined as $\sqmap{i}$.
-
-\begin{defn}
-  A binary relation $R \subseteq X \times Y$ between two sets $X, Y$ is a subset of the 
-  cartesian product of the sets. A relation is called \textit{homogeneous} or an \textit
-  {endorelation} if it is a binary relation over $X$ and itself: $R \subseteq X \times
-   X$. 
-  In such case, it is simply called a binary relation over $X$.
-  An \textit{equivalence relation} is a binary relation that is reflexive, symmetric and
-  transitive. 
-\end{defn}
-
-\begin{defn}
-  The congruence closure $c(R)$ of a relation R is the smallest congruence relation 
-  $R'$ such that $R \subseteq R'$ 
-\end{defn}
-
-An equivalence relation which is compatible with all the operations of
-the algebraic structure on which it is defined on, is called a 
-\textit{congruence relation}. Compatibility with the algebraic structure operations
-means that algebraic operations applied on equivalent elements will still
-yield equivalent elements.
-
-%===================================================================================
-
-We omit the coalgebraic definition for \textit{linear weighted automata} seen in 
-\cite{BONCHI201277} and give a more intuitive definition. 
-In this implementation, we focus only on weighted automata defined over 
-the field of real numbers $\R$.
-
-\begin{defn}
-  A \textit{linear weighted automaton} (in short, LWA) over the field $\K$ and an alphabet $A$
-  is a triple  $L = (V, o, \left\lbrace t_a \right\rbrace_{a \in A})$ 
-  where $V$ is a vector space representing the state space, 
-  $o: V \to \K$ is a linear map associating to each state its output weight,
-  and $t = \left\lbrace t_a = V \times V \right\rbrace_{a \in A}$ is
-  the set of transition functions, represented with liner maps 
-  that for each input $a \in A$ associate the next state, in this case a vector
-  in $V$.
-  As in \cite{boreale2009weighted}, we have that $\dim{(L)} = \dim{(V)}$.
-\end{defn}
-
-Given a weighted automaton, one can build an isomorphic linear weighted automaton
-by considering the free vector space generated by the set of states $X$ in the WA,
-and by linearizing $o$ and $t$. If $X$ is finite, we can use the same matrices for 
-$t$ and $o$ in both the WA and the corresponding LWA.
-We are only considering a finite number of states and therefore finite dimensional
-vector spaces. Let $n$ be the (finite) number of states in an WA.
-We have that in the corresponding LWA, the transition functions $t_a$ are still
- represented as
-$\K^{n \times n}$ matrices. $o \in \K^{1 \times n}$ is represented as a row vector.
-$t_a(v)$ denotes the vector obtained by multiplying the matrix $t_a$ by the column 
-vector $v  \in \K^{n \times 1}$. $o(v)$ denotes the scalar $s \in \K$ obtained by 
-dot product of the row vector $o$ with $v \in \K^{n \times 1}$.
-
-\begin{defn}
-  The language recognized by a vector $v \in V$ of an LWA $(V,o,t)$ is defined
-  for all words $w \in A*$ as $\langlwa{v}{V}(w) = o(v_n)$ where $v_n$ is the 
-  vector reached from $v$ through the composition of the transition functions
-  corresponding to the words in $w$.
-
-  \begin{equation*}
-    \begin{aligned}
-      \langlwa{v}{V}(w) = \begin{cases}
-        o(v) & \text{if } w = \epsilon \\ 
-        \langlwa{t_a(v)}{V}(w') & \text{if } w = aw' 
-      \end{cases}
-    \end{aligned}
-  \end{equation*}
-
-\end{defn}
-
-We define $\llwb$ as the behavioral equivalence for a given LWA $(V, o, t)$ as
-
-\begin{equation}
-  \forall v_1, v_2 \in V, \; v_1 \llwb v_2 \iff \langlwa{v_1}{V} = \langlwa{v_2}{V}
-\end{equation}
-
-\begin{lem}
-  $\llwb$ coincides with $\sim_l$:
-
-  Let $(X, o, t)$ be a WA and $(\K^X, o^\sharp, t^\sharp)$ the corresponding linear 
-  weighted automaton. Then $\forall x \in X, \;\; \sqmap{x} = \langlwa{x}{\K^X}$
-\end{lem}
-
-\begin{proof}
-  Proved in section 3.3 of \cite{BONCHI201277}
-\end{proof}
-
-Language equivalence can be expressed in terms of linear weighted bisimulations (LWBs for short).
-Differently from weighted bisimulations, LWBs can be seen both as relations and as subspaces.
-The subspace representation of LWBs is used in the backwards partition refinement algorithm 
-implemented in \cite{BONCHI201277} and in this work.
+\input{sections/notation.tex}
 
@@
 
-\section{The Problem}
-We want to check if two vectors are language equivalent $v_1 \sim_l v_2$.
-
-\section{Algorithms}
-
-We now recall the backwards algorithm for computing $\llwb$ defined in \cite{BONCHI201277}.
-
-\subsection{Backwards Partition Refinement Algorithm for the Largest Weighted Bisimulation}
-\label{sec:algo2}
-
-The algorithm is defined by the iterative method:
-\begin{eqnarray}
-  R_0 = \mker{o}^0, & \quad & R_{i+1} = R_i + \sum_{a \in A} t_a^t(R_i) \label{back} 
-\end{eqnarray}
-Where $\mker{o}^0$ is an annihilator.
-The algorithm stops when $R_{j+1} = R_j$. An index $j \leq \dim(V)$ is 
-guaranteed to exist, such that the algorithms terminates at step $j$.
-It follows that $\llwb \equiv R_j^0$.
-Proof is available in section 4.2 of \cite{BONCHI201277}
-
-%\subsection{Forward Partition Refinement Algorithm for the Largest Weighted Bisimulation}
-%\label{sec:algo2}
-%
-%The algorithm is defined by the iterative method:
-%\begin{eqnarray}
-%  R_0 = \mker{o}, & \quad & R_{i+1} = R_i \cap \bigcap_{a \in A} t_a(R_i)^{-1} \label{fwd} 
-%\end{eqnarray}
-%The algorithm stops when $R_{j+1} = R_j$. Proof is available in section 4.1 of \cite{BONCHI201277}
+\input{sections/algorithms.tex}
 
@@
 
-\begin{figure}
-  \centering
-  \fbox{\rule[-.5cm]{4cm}{4cm} \rule[-.5cm]{4cm}{0cm}}
-  \caption{Sample figure caption.}
-  \label{fig:fig1}
-\end{figure}
+%\begin{figure}
+%  \centering
+%  \fbox{\rule[-.5cm]{4cm}{4cm} \rule[-.5cm]{4cm}{0cm}}
+%  \caption{Sample figure caption.}
+%  \label{fig:fig1}
+%\end{figure}
 
-\begin{table}
- \caption{Sample table title}
-  \centering
-  \begin{tabular}{lll}
-    \toprule
-    \multicolumn{2}{c}{Part}                   \\
-    \cmidrule(r){1-2}
-    Name     & Description     & Size ($\mu$m) \\
-    \midrule
-    Dendrite & Input terminal  & $\sim$100     \\
-    Axon     & Output terminal & $\sim$10      \\
-    Soma     & Cell body       & up to $10^6$  \\
-    \bottomrule
-  \end{tabular}
-  \label{tab:table}
-\end{table}
+%\begin{table}
+% \caption{Sample table title}
+%  \centering
+%  \begin{tabular}{lll}
+%    \toprule
+%    \multicolumn{2}{c}{Part}                   \\
+%    \cmidrule(r){1-2}
+%    Name     & Description     & Size ($\mu$m) \\
+%    \midrule
+%    Dendrite & Input terminal  & $\sim$100     \\
+%    Axon     & Output terminal & $\sim$10      \\
+%    Soma     & Cell body       & up to $10^6$  \\
+%    \bottomrule
+%  \end{tabular}
+%  \label{tab:table}
+%\end{table}
 
@@
 % 
+\appendix
+
+%\input{sections/appendix.tex}
 \end{document}

### 2020-10-05 paper/sections/algorithms.tex
working on tex, added random tests

\section{Algorithms}
The first algorithm we implement to compute language equivalence, called \texttt{HKC},
is adapted from \cite{DBLP:journals/corr/Bonchi0K17}. It was first introduced by 
Bonchi and Pous in \cite{bonchi2013checking}.
The algorithm, extending the Hopcroft and Karp procedure 
\cite{hopcroft1971linear} with \textit{congruence closure}, is 
proven to be sound and complete \cite{DBLP:journals/corr/Bonchi0K17}.
Originally, the algorithm was defined for 
WAs over semirings, but we recall that in this work we are only 
considering fields, in particular 
the field of real numbers ($\K = \R$).
The problem of checking language equivalence 
has been proven undecidable for an arbitrary semiring, so termination 
may not always be guaranteed. However, it has been shown to be decidable
for a broad range of semirings, for example, all the complete and
distributive lattices.
\texttt{HKC} computes $v_1 \sim_l v_2$ for a given weighted automaton
$W = (X, t, o)$ and two vectors $v_1, v_2 \in \K^X$. 
by computing a congruence closure,
and it does so without linearizing the state space.

We compare \texttt{HKC} with an algorithm called 
\textit{Backwards Partition Refinement}, that we will call BPR for short.

% TODO
(TODO chiedi a filippo: nelle conclusioni di \cite{BONCHI201277} dice che è diverso 
dall'algoritmo visto nel paper di boreale \cite{boreale2009weighted}. è il caso 
anche per i field e i numeri reali???? o sono lo equivalenti?).

Adapted from \cite{BONCHI201277}, 
\texttt{BPR} is a fixed-point iterative method that, given an LWA
$L = (V, t, o)$,
computes a basis of the subspace of $V$
representing the complementary relation of $\llwb$ (we later show it to be the 
orthogonal complement in case $V$ is an inner product space). 
Another version of the algorithm is defined in the same work,
called \textit{Forward Partition Refinement}, which directly computes
a basis for $\llwb$
but is shown to be way less efficient than the backwards version.

\begin{note}
    Recall from section \ref{sec:notation} that $\llwb$ is a linear relation, 
    therefore $v_1 \llwb v_2 \iff (v_1 - v_2) \in \mker{\llwb}$
\end{note}

The \texttt{BPR} algorithm starts from a relation $R_0$, that is the complement 
of the relation identifying vectors with equal weights.
It then incrementally computes the space of all states that are reachable from 
$R_0$ in a \textit{backwards} direction. Intuitively, "going backwards" means 
working with the transpose transitions functions $t_a^t$.

In the next section we compare execution results of our implementation of the algorithms
\texttt{BPR} and \texttt{HKC} to verify the correctness of the implementation,
and to analyze when one of the two algorithms is the convenient choice.
Lemma \ref{lem:coincide}, introduced above, is key to our work. By stating that 
$\llwb$ coincides with $\sim_l$, we can confidently say that the two algorithms 
compute an answer for same the decision problem:

\begin{center}
    Are two vectors $v_1$ and $v_2$ language-equivalent for a given weighted automata? 
\end{center}

%TODO
TODO costo computazionale HKC.
\texttt{BPR} has a cost of $O(n^4)$ operations 
to initially compute the largest linear weighted bisimulation,
which can be eventually reduced to $O(n^3)$.
In our implementation, by computing a basis of the orthogonal complement of $\llwb$,
the cost of checking if two vectors are language equivalent is reduced to the
cost of matrix multiplication ($O(n^2)$). \texttt{BPR} is efficient when we
have to decide if a large number of vectors in a WA are language equivalent.

%=========================================================================

\subsection{HKC Algorithm}
We give a pseudocode definition of the \texttt{HKC} procedure:

\captionof{figure}{The \texttt{HKC}($v_1, v_2$) procedure}
\lstinputlisting[escapeinside={(*}{*)}]{hkc.txt}
\label{fig:hkc}

\subsection{Backwards Partition Refinement Algorithm for the Largest Weighted Bisimulation}
\label{sec:algo2}

We now recall the backwards algorithm for computing $\llwb$ defined in \cite{BONCHI201277}.
The algorithm is defined by the iterative method:
\begin{eqnarray}
  R_0 = \mker{o}^0, & \quad & R_{i+1} = R_i + \sum_{a \in A} t_a^t(R_i) \label{back} 
\end{eqnarray}
Where $\mker{o}^0$ is an annihilator.
The algorithm stops when $R_{j+1} = R_j$. An index $j \leq \dim(V)$ is 
guaranteed to exist, such that the algorithms terminates at step $j$.
It follows that $\llwb = R_j^0$.
Proof is available in section 4.2 of \cite{BONCHI201277}

### 2020-10-05 paper/sections/introduction.tex
working on tex, added random tests

\section{Introduction}
\label{sec:intro}
TODO da fare

Checking whether two nondeterministic finite automata (NFA) accept the same
language is important in many application domains such as compiler construction and model checking. 
Unfortunately, solving this problem is costly: it is
PSPACE-complete (TODO cita) 
% TODO cita

In \cite{DBLP:journals/corr/Bonchi0K17}, up-to techniques are defined for
weighted systems over arbitrary semirings, while in \cite{BONCHI201277}, up-to techniques
are defined for Linear Weighted Automata (LWAs), under a more abstract coalgebraic perspective.

A well detailed example of the comparison of algorithms to compute 
language equivalence, precisely between \texttt{HKC} and an alternative
algorithm called the \textit{antichain algorithm} (\cite{de2006antichains}),
was published in 2017 \cite{fu2017equivalence}.

### 2020-10-05 paper/sections/notation.tex
working on tex, added random tests

\section{Preliminaries and Notation}
\label{sec:notation}

\begin{note}
  Given two vector spaces $V_1, V_2$ we write $V_1 + V_2$ to 
  denote $\mspan{V_1 \cup V_2}$
\end{note}

\begin{defn}
  A \textit{weighted automaton} over a field $\K$ and an alphabet $A$ is a triple 
  $(X,o,t)$ such that $X$ is a finite set of states, 
  $t = \left\lbrace t_a : X \to \K^X\right\rbrace_{a \in A}$
  is a set of transition functions indexed over the symbols of the alphabet $A$ and 
  $o : X  \to \K$ is the output function. 
  The transition functions will be represented as $X \times X$ matrices.
  $A^*$ is the set of all words over $A$, more precisely the free monoid
  with string concatenation as the monoid operation and the empty word $\epsilon$ 
  as the identity element. We denote with $aw$ the
  concatenation of a symbol $a$ to the word $w \in A^*$.
  A weighted language is a function $\psi: A^* \to \K$.
  A function mapping each state vector into its 
  accepted language, $\sqmap{\cdot}: \K^X \to \K^{A^*}$ is defined as follows for 
  every weighted automaton:

  \begin{equation*}
    \begin{aligned}
      \forall v \in \K^X, a \in A, w \in A^* \quad \quad
      \sqmap{v}(\epsilon) = o(v) \quad \quad
      \sqmap{v}(aw) = \sqmap{t_a(v)}(w)  
    \end{aligned}
  \end{equation*}
\end{defn}

Two vectors $v_1, v_2 \in \K^{X\times 1}$ are called weighted language equivalent, 
denoted with  $v_1 \sim_l v_2 $ if and only if 
$ \sqmap {v_1} = \sqmap{v_2}$. One can extend the notion of language 
equivalence to states rather than for vectors by assigning 
to each state $x \in X$ the corresponding  unit vector 
$e_x \in \K^X$. When given an initial state $i$ for a weighted automaton, 
the language  of the automaton can be defined as $\sqmap{i}$.

\begin{defn}
  A binary relation $R \subseteq X \times Y$ between two sets $X, Y$ is a 
  subset of the 
  cartesian product of the sets. A relation is called \textit{homogeneous} 
  or an \textit
  {endorelation} if it is a binary relation over $X$ and itself: 
  $R \subseteq X \times
   X$. 
  In such case, it is simply called a binary relation over $X$.
  An \textit{equivalence relation} is a binary relation that is reflexive, 
  symmetric and
  transitive. 
\end{defn}

\begin{defn}
  The congruence closure $c(R)$ of a relation R is the smallest congruence relation 
  $R'$ such that $R \subseteq R'$ 
\end{defn}

An equivalence relation which is compatible with all the operations of
the algebraic structure on which it is defined on, is called a 
\textit{congruence relation}. Compatibility with the algebraic structure operations
means that algebraic operations applied on equivalent elements will still
yield equivalent elements.

%===================================================================================

We omit the coalgebraic definition for \textit{linear weighted automata} seen in 
\cite{BONCHI201277} and give a more intuitive definition, which fits our  
implementation when $\K = \R$.
In this implementation, we focus only on weighted automata defined over 
the field of real numbers $\R$.

\begin{defn}
  A \textit{linear weighted automaton} (in short, LWA) over the field $\K$ and 
  an alphabet $A$
  is a triple  $L = (V, o, \left\lbrace t_a \right\rbrace_{a \in A})$ 
  where $V$ is a vector space representing the state space, 
  $o: V \to \K$ is a linear map associating to each state its output weight,
  and $t = \left\lbrace t_a = V \times V \right\rbrace_{a \in A}$ is
  the set of transition functions, represented with liner maps 
  that for each input $a \in A$ associate the next state, in this case a vector
  in $V$.
  As in \cite{boreale2009weighted}, we have that $\dim{(L)} = \dim{(V)}$.
\end{defn}

Given a weighted automaton, one can build a corresponding linear weighted automaton
by considering the free vector space generated by the set of states $X$ in the WA,
and by linearizing $o$ and $t$. If $X$ is finite, as in our implementations
of the algorithms, we can use the same matrices for 
$t$ and $o$ in both the WA and the corresponding LWA.
We are only considering a finite number of states and therefore finite dimensional
vector spaces. Let $n$ be the number of states in an WA.
We have that in the corresponding LWA, the transition functions $t_a$ are still
 represented as
$\K^{n \times n}$ matrices. $o \in \K^{1 \times n}$ is represented as a row vector.
$t_a(v)$ denotes the vector obtained by multiplying the matrix $t_a$ by the column 
vector $v  \in \K^{n \times 1}$. $o(v)$ denotes the scalar $s \in \K$ obtained by 
dot product of the row vector $o$ with $v \in \K^{n \times 1}$.

\begin{defn}
  The language recognized by a vector $v \in V$ of an LWA $(V,o,t)$ is defined
  for all words $w \in A*$ as $\langlwa{v}{V}(w) = o(v_n)$ where $v_n$ is the 
  vector reached from $v$ through the composition of the transition functions
  corresponding to the words in $w$.

  \begin{equation*}
    \begin{aligned}
      \langlwa{v}{V}(w) = \begin{cases}
        o(v) & \text{if } w = \epsilon \\ 
        \langlwa{t_a(v)}{V}(w') & \text{if } w = aw' 
      \end{cases}
    \end{aligned}
  \end{equation*}

\end{defn}

We define $\llwb$ as the behavioral equivalence for a given LWA $(V, o, t)$ as

\begin{equation}
  \forall v_1, v_2 \in V, \; v_1 \llwb v_2 \iff \langlwa{v_1}{V} = \langlwa{v_2}{V}
\end{equation}

Proof is available in section 3.3 of \cite{BONCHI201277}

Language equivalence can be now expressed in terms of linear weighted 
bisimulations (LWBs for short).
Differently from weighted bisimulations, LWBs can be seen both as relations 
and as subspaces.
The subspace representation of LWBs is used in the backwards partition 
refinement algorithm 
implemented in \cite{BONCHI201277} and in this work.

\begin{defn}
  \textit{Linear Relations:}\\
  Let $U$ be a subspace of $V$. The binary relation $R_U$ over $V$ is defined by
  \begin{equation*}
    \begin{aligned}
      v_1 \; R_U \; v_2 \quad \iff \quad v_1 - v_2 \in U
    \end{aligned}
  \end{equation*}
  The relation $R$ is linear if there exists a subspace $U$ such that $R \equiv R_U$.
  A linear relation is a total equivalence relation on $V$.
\end{defn}

\begin{defn}
  \textit{Kernel of a Relation and Linear Extension} \\
  Let $R$ be a binary relation over V. 
  The \textit{kernel} of $R$, is the set 
  $\mker{R} = \left\lbrace v_1 - v_2 \mid v_1 \; R \; v_2 \right\rbrace$.
  The \textit{linear extension} of $R$, written as $R^\ell$, is defined by 
  \begin{equation*}
    \begin{aligned}
      v_1 \; R^\ell \; v_2  \quad \iff \quad (v_1 - v_2) \in \mspan{\mker{R}}
    \end{aligned}
  \end{equation*}
\end{defn}

\begin{lem}
  Let $U$ be a subspace of $V$, then $\mker{R_U} = U$
\end{lem}

\begin{defn}
  \textit{Linear Weighted Bisimulation:} \\
  Let $(V, o, t)$ be a linear weighted automaton. A linear relation 
  $R \subseteq V \times V$ is a \textit{linear weighted bisimulation} if 
  $\forall (v_1, v_2) \in R$ it holds that: 
  \begin{center}
    \begin{enumerate}
      \item $o(v_1) = o(v_2)$
      \item $\forall a \in A, \; t_a(v_1) \; R \; t_a(v_2)$
    \end{enumerate}
  \end{center}
\end{defn}

\begin{lem}
  Let $(V, o, t)$ be a linear weighted automaton. A linear relation 
  $R$ over $V$ is a linear weighted bisimulation if and only if
\end{lem}
  \begin{center}
    \begin{enumerate}
      \item $R \subseteq \mker{o}$
      \item $R$ is $t_a$-invariant $\forall a \in A$
    \end{enumerate}
  \end{center}

Theorem 3 in section 3.3 of \cite{BONCHI201277}, states that 
$\mker{\langlwa{-}{V}}$ is the largest linear weighted bisimulation on $V$.
As a corollary, we obtain that $\llwb$ is the largest linear weighted bisimulation.

We now introduce a lemma that will be fundamental in the next sections of this work.

\begin{lem}
  \label{lem:coincide}
  $\llwb$ coincides with $\sim_l$:

  Let $(X, o, t)$ be a WA and $(\K^X, o^\sharp, t^\sharp)$ the corresponding linear 
  weighted automaton. Then $\forall x \in X, \;\; \sqmap{x} = \langlwa{x}{\K^X}$
\end{lem}

### 2020-10-05 automata/automata.go

@@ -1,56 +1,30 @@
-// This file contains functions for reading automatas transition matrices and output vectors
+// This file contains weighted automata data structure definition
+// and basic transition functions.
 
 package automata
 
 import (
-	"errors"
 	"math"
-	"math/rand"
 
-	"github.com/0x0f0f0f/lwa-techniques/lin"
 	"gonum.org/v1/gonum/mat"
 )
 
-// ======================================================================
-
-// Error helpers
-
-func errRead(msg string) error { return errors.New("could not read automaton:" + msg) }
-
 type Automaton struct {
-	A   []string              // The input alphabet
-	T   map[string]*mat.Dense // Transition matrices are maps from input strings to dense real valued matrices
-	O   *mat.VecDense         // Output vector uses a dense real valued vector
-	Dim int                   // Number of states/dimension of vector space V in LWA
-	//LLWB     *mat.Dense            // Col(LLWB) is a basis of the largest linear weighted bisimulation, a binary linear relation. vRw iff (v-w) in Ker(R)
+	// The input alphabet
+	A []string
+	// Transition matrices are maps from input strings
+	// to dense real valued matrices
+	T map[string]*mat.Dense
+	// Output vector uses a dense real valued vector
+	O *mat.VecDense
+	// Number of states/dimension of vector space V in LWA
+	Dim int
+	// Col(LLWB) is a basis of the largest linear weighted bisimulation,
+	// a binary linear relation. vRw iff (v-w) in Ker(R)
+	// with LLWBperp, we denote a basis of the orthogonal
+	// complement of the basis LLWB
 	LLWBperp *mat.Dense
-	Tol      float64 // if abs(something) < tol, something is interpreted as 0
-}
-
-func RandAutomaton(syms, states, maxweight int, tol float64) Automaton {
-	// create the alphabet
-	A := make([]string, syms)
-	if syms <= 57 {
-		for i := 0; i < syms; i++ {
-			A[i] = string(rune(i + 65))
-		}
-	}
-
-	T := map[string]*mat.Dense{}
-	for _, sym := range A {
-		T[sym] = lin.RandIntDense(states, maxweight)
-		lin.PokeHoles(T[sym], rand.Intn((states*states)/2))
-	}
-
-	aut := Automaton{
-		A:   A,
-		T:   T,
-		O:   lin.RandVec(states),
-		Dim: states,
[…]

### 2020-10-05 automata/backwards.go

@@ -20,37 +20,19 @@ func (a *Automaton) BackwardsPartitionRefinement() {
 	// lastIndex := 0
 
 	for i := 1; i <= a.Dim; i++ {
-		// fmt.Printf("==============================\nstep %d\n", i)
 		// Σ_{a \in A} T_a^t(R_i)
 		for _, sym := range a.A {
-			//_, n := lastBasis.Dims()
-			// if lastIndex == n {
-			// break
-			// }
-			// toCompute := lastBasis.Slice(0, a.Dim, lastIndex, n).(*mat.Dense)
 			newBasis := a.ApplyTransposeTransitionBasis(sym, lastBasis)
-			// fmt.Printf("T_(%s)^t of \n%s \n = \n", sym, lin.StringMat(lastBasis))
-			// lin.PrintMat(newBasis)
 			currBasis = lin.Union(currBasis, newBasis)
-			// fmt.Println("curr basis = ")
-			// lin.PrintMat(currBasis)
-
 			currBasis = lin.OrthonormalColumnSpaceBasis(currBasis, a.Tol).(*mat.Dense)
-			// fmt.Printf("orthonormal basis of col space of b = \n%s\n", lin.StringMat(currBasis))
 		}
 
-		//if i > 1 && mat.Equal(lastBasis, currBasis) {
-		//	break
-		//}
-		// _, lastIndex = lastBasis.Dims()
-		// _, currSize := currBasis.Dims()
-		// fmt.Printf("B_%d has size %d \n", i-1, lastIndex)
-		// fmt.Printf("B_%d has size %d \n", i, currSize)
 		lastBasis = currBasis
 	}
 
 	a.LLWBperp = currBasis
-	//a.LLWB = lin.Complement(currBasis).(*mat.Dense)
+	// we could compute the orthogonal complement to find a basis of LLWB:
+	// a.LLWB = lin.Complement(currBasis).(*mat.Dense)
 }
 
 func (a Automaton) BPREquivalence(v1, v2 *mat.VecDense) bool {

### 2020-10-05 automata/hkc.go

@@ -1,14 +1,12 @@
 package automata
 
 import (
-	"errors"
-
 	"gonum.org/v1/gonum/mat"
 )
 
 // checks the language equivalence of two state vectors for a given weighted automaton
 func (a Automaton) HKC(v1, v2 *mat.VecDense, maxiter int) (bool, error) {
-	rel := NewRelation(0)
+	rel := NewRelation(0, a.Dim)
 	todo := NewPairStack()
 
 	p, err := NewPair(v1, v2)
@@ -16,52 +14,37 @@ func (a Automaton) HKC(v1, v2 *mat.VecDense, maxiter int) (bool, error) {
 		return false, err
 	}
 
-	// fmt.Println("p = ", p)
-
 	// insert (v1, v2) into the todo list
-	todo.Push(*p)
+	todo = PairStackPush(todo, p)
 
 	i := 0
-	for !todo.Empty() {
-		if i >= maxiter {
-			return false, errors.New("maximum iteration exceeded")
-		}
-		// fmt.Println("Step", i)
-
-		// fmt.Println("todo =", todo)
-
+	for !todo.IsEmpty() {
 		// extract (v1', v2') from todo
-		q, err := todo.Pop()
+		q, err := PairStackPop(todo)
 		if err != nil {
 			return false, err
 		}
-		//fmt.Println("Extracted pair is", q)
-		//fmt.Println("todo =", todo)
 
 		if rel.PairIsInCongruenceClosure(q) {
-			// fmt.Println("(v1', v2') \\in c(R)")
 			continue
 		}
 
-		o1 := a.GetOutput(q.Left)
-		o2 := a.GetOutput(q.Right)
-		// fmt.Println("o(v1) =", o1)
-		// fmt.Println("o(v2) =", o2)
+		o1 := a.GetOutput(PairLeft(q))
+		o2 := a.GetOutput(PairRight(q))
 		if o1 != o2 {
-			// fmt.Println("o(v1) =/= o(v2)")
 			return false, nil
 		}
 
 		for _, sym := range a.A {
 
-			w1 := a.ApplyTransition(sym, q.Left)
-			w2 := a.ApplyTransition(sym, q.Right)
+			w1 := a.ApplyTransition(sym, PairLeft(q))
+			w2 := a.ApplyTransition(sym, PairRight(q))
 			wp, err := NewPair(w1, w2)
 			if err != nil {
 				return false, err
 			}
 
-			todo.Push(*wp)
[…]

### 2020-10-05 automata/io.go

@@ -3,6 +3,7 @@ package automata
 import (
 	"bufio"
 	"bytes"
+	"errors"
 	"fmt"
 	"io"
 	"os"
@@ -12,6 +13,8 @@ import (
 
 	"gonum.org/v1/gonum/mat"
 )
+
+func errRead(msg string) error { return errors.New("could not read automaton:" + msg) }
 
 // Helper function to sort and deduplicate in-place a slice of strings
 func dedupStr(in []string) []string {

### 2020-10-05 automata/pair.go

@@ -2,44 +2,64 @@ package automata
 
 import (
 	"errors"
-	"fmt"
 
 	"github.com/0x0f0f0f/lwa-techniques/lin"
 	"gonum.org/v1/gonum/mat"
 )
 
-// represents an unordered pair of vectors
-type Pair struct {
-	Left  *mat.VecDense
-	Right *mat.VecDense
-	Dim   int
-}
-
 // creates a new pair of vectors
-func NewPair(l, r *mat.VecDense) (*Pair, error) {
+func NewPair(l, r *mat.VecDense) (*mat.Dense, error) {
 	ld, _ := l.Dims()
 	rd, _ := r.Dims()
 	if ld != rd {
 		return nil, errors.New("dimensions of vector do not match")
 	}
-	return &Pair{Left: l, Right: r, Dim: ld}, nil
+	m := mat.NewDense(ld, 2, nil)
+
+	for i := 0; i < ld; i++ {
+		m.Set(i, 0, l.AtVec(i))
+		m.Set(i, 1, r.AtVec(i))
+	}
+	return m, nil
+}
+
+// return left element of a vector pair
+func PairLeft(p *mat.Dense) *mat.VecDense {
+	return p.ColView(0).(*mat.VecDense)
+}
+
+// return right element of a vector pair
+func PairRight(p *mat.Dense) *mat.VecDense {
+	return p.ColView(1).(*mat.VecDense)
+}
+
+// return subtraction of elements of a vector pair
+func PairSub(p *mat.Dense) *mat.VecDense {
+	m, _ := p.Dims()
+	sub := mat.NewVecDense(m, nil)
+	sub.SubVec(PairLeft(p), PairRight(p))
+	return sub
+}
+
+// return true if a matrix is a vector pair
+func PairCheck(p *mat.Dense) bool {
+	_, n := p.Dims()
+	return n == 2
 }
 
 // Returns true if two pairs equal each other
-func (p Pair) Eqs(p1 *Pair, tol float64) bool {
+func PairEqs(p, p1 *mat.Dense, tol float64) bool {
+	m, _ := p.Dims()
+	m1, _ := p1.Dims()
 	// if the dimensions do not match, pairs are not equal
-	if p.Dim != p1.Dim {
+	if m != m1 || !PairCheck(p) || !PairCheck(p1) {
 		return false
 	}
[…]

### 2020-10-05 automata/pairstack.go

@@ -1,47 +1,54 @@
+// stack data structure for real valued vector pairs, used in the HKC algorithm
+
 package automata
 
 import (
 	"errors"
-	"fmt"
+
+	"gonum.org/v1/gonum/mat"
 )
 
-// stack data structure for real valued vector pairs, used in the HKC algorithm
-type PairStack struct {
-	stack []Pair
-}
-
-// create a new empty pair stack
-func NewPairStack() PairStack {
-	return PairStack{stack: []Pair{}}
-}
-
-// returns true if the stack is empty, false otherwise
-func (ps PairStack) Empty() bool {
-	return len(ps.stack) == 0
+func NewPairStack() *mat.Dense {
+	m := mat.NewDense(1, 1, nil)
+	m.Reset()
+	return m
 }
 
 // returns the size
-func (ps PairStack) Size() int {
-	return len(ps.stack)
+func PairStackSize(s *mat.Dense) int {
+	if s.IsEmpty() {
+		return 0
+	}
+	_, n := s.Dims()
+	return n
 }
 
 // push a pair into the stack
-func (ps *PairStack) Push(p Pair) {
-	ps.stack = append(ps.stack, p)
-}
-
-func (ps PairStack) String() string {
-	return fmt.Sprintf("%v", ps.stack)
+func PairStackPush(s *mat.Dense, p *mat.Dense) *mat.Dense {
+	if s.IsEmpty() {
+		return mat.DenseCopyOf(p)
+	}
+	s.Augment(s, p)
+	return s
 }
 
 // pop a pair from the stack
-func (ps *PairStack) Pop() (*Pair, error) {
-	siz := ps.Size()
-	if siz == 0 {
+func PairStackPop(s *mat.Dense) (*mat.Dense, error) {
+	if s.IsEmpty() {
 		return nil, errors.New("stack is empty")
 	}
-	el := ps.stack[siz-1]
-	ps.stack = ps.stack[:siz-1]
+	m, n := s.Dims()
+	if n%2 != 0 {
+		return nil, errors.New("inconsisten stack: odd number of elements")
+	}
 
-	return &el, nil
+	pair := s.Slice(0, m, n-2, n).(*mat.Dense)
+
+	if n-2 > 0 {
+		s = s.Slice(0, m, 0, n-2).(*mat.Dense)
+	} else {
+		s.Reset()
+	}
+
+	return pair, nil
 }

### 2020-10-05 automata/random.go

@@ -1,0 +1,72 @@
+package automata
+
+import (
+	"math/rand"
+
+	"github.com/0x0f0f0f/lwa-techniques/lin"
+	"gonum.org/v1/gonum/mat"
+)
+
+// generate a random automaton on natural numbers
+func RandNatAutomaton(syms, states, maxweight int, tol float64) Automaton {
+	// create the alphabet
+	A := make([]string, syms)
+	if syms <= 57 {
+		for i := 0; i < syms; i++ {
+			A[i] = string(rune(i + 65))
+		}
+	}
+
+	T := map[string]*mat.Dense{}
+	for _, sym := range A {
+		T[sym] = lin.RandIntDense(states, maxweight)
+		lin.PokeHoles(T[sym], rand.Intn((states*states)/2))
+	}
+
+	O := lin.RandNatVec(states, maxweight)
+	for lin.IsZero(O) {
+		O = lin.RandNatVec(states, maxweight)
+	}
+
+	aut := Automaton{
+		A:   A,
+		T:   T,
+		O:   O,
+		Dim: states,
+		Tol: tol,
+	}
+
+	return aut
+}
+
+// generate a random automaton on real numbers
+func RandAutomaton(syms, states int, maxweight float64, tol float64) Automaton {
+	// create the alphabet
+	A := make([]string, syms)
+	if syms <= 57 {
+		for i := 0; i < syms; i++ {
+			A[i] = string(rune(i + 65))
+		}
+	}
+
+	T := map[string]*mat.Dense{}
+	for _, sym := range A {
+		T[sym] = lin.RandDense(states, maxweight)
+		lin.PokeHoles(T[sym], rand.Intn((states*states)/2))
+	}
+
+	O := lin.RandVec(states, maxweight)
+	for lin.IsZero(O) {
+		O = lin.RandVec(states, maxweight)
+	}
+
+	aut := Automaton{
+		A:   A,
+		T:   T,
+		O:   O,
+		Dim: states,
+		Tol: tol,
+	}
+
+	return aut
+}

### 2020-10-05 automata/relation.go

@@ -1,8 +1,7 @@
 package automata
 
 import (
-	"bytes"
-	"fmt"
+	"errors"
 
 	"github.com/0x0f0f0f/lwa-techniques/lin"
 	"gonum.org/v1/gonum/mat"
@@ -12,20 +11,41 @@ import (
 // the relation is a congruence if it is an equivalence
 // and is closed under linear combinations.
 type Relation struct {
-	s   []*Pair         // the set of pairs in the relations
-	u   []*mat.VecDense // generating set for the congruence closure
-	tol float64
+	s    *mat.Dense // the set of pairs in the relations
+	u    *mat.Dense // generating set for the congruence closure
+	size int        // how many pairs
+	dim  int        // number of rows
+	tol  float64
 }
 
 // creates a new empty relation
-func NewRelation(tol float64) Relation {
-	return Relation{tol: tol}
+func NewRelation(tol float64, dim int) Relation {
+	s := mat.NewDense(dim, 1, nil)
+	u := mat.NewDense(dim, 1, nil)
+	s.Reset()
+	u.Reset()
+	return Relation{tol: tol, dim: dim, s: s, u: u}
+}
+
+func (r Relation) GetPair(i int) (*mat.Dense, error) {
+	if i < 0 || i >= r.size {
+		return nil, errors.New("index out of bounds")
+	}
+	return r.s.Slice(0, r.dim, i*2, (i*2)+2).(*mat.Dense), nil
 }
 
 // returns true if the relation contains the pair of vectors. O(n)
-func (r Relation) Has(p *Pair) bool {
-	for _, v := range r.s {
-		if v.Eqs(p, r.tol) {
+func (r Relation) Has(p *mat.Dense) bool {
+	m, _ := r.s.Dims()
+	if m != r.dim && !PairCheck(p) {
+		panic(errors.New("dimension mismatch"))
+	}
+	for i := 0; i < r.size; i++ {
+		p1, err := r.GetPair(i)
+		if err != nil {
+			panic(err)
+		}
+		if PairEqs(p1, p, r.tol) {
 			return true
 		}
 	}
@@ -33,37 +53,54 @@ func (r Relation) Has(p *Pair) bool {
 }
 
 // adds a pair of vectors to the relation
-func (r *Relation) Add(p *Pair) {
[…]

### 2020-10-05 lin/subspace.go

@@ -71,6 +71,5 @@ func Complement(spanset mat.Matrix, tol float64) mat.Matrix {
 	}
 
 	basis, _ := Nullspace(spanset.T(), tol)
-
 	return basis
 }

### 2020-10-05 lin/util.go

@@ -1,3 +1,6 @@
+// this file contains miscellaneous utility functions for various
+// linear algebra applications using Gonum
+
 package lin
 
 import (
@@ -32,7 +35,7 @@ func IsZeroTol(vec *mat.VecDense, tol float64) bool {
 }
 
 // returns true if two vectors are equal with a tolerance of tol
-func EqVecTol(a, b *mat.VecDense, tol float64) bool {
+func EqVecTol(a, b mat.Vector, tol float64) bool {
 	sub := mat.VecDenseCopyOf(a)
 	sub.SubVec(a, b)
 
@@ -48,11 +51,20 @@ func EyeDense(n int) *mat.Dense {
 	return a
 }
 
+// generate a random float64 between -w and w
+func randFloat64(w float64) float64 {
+	sign := 1.0
+	if rand.Intn(2) == 0 {
+		sign = -1.0
+	}
+	return sign * rand.Float64() * w
+}
+
 // generate a normally distributed random n*n matrix
-func RandDense(n int) *mat.Dense {
+func RandDense(n int, maxweight float64) *mat.Dense {
 	data := make([]float64, n*n)
 	for i := range data {
-		data[i] = rand.NormFloat64()
+		data[i] = randFloat64(maxweight)
 	}
 	return mat.NewDense(n, n, data)
 }
@@ -97,10 +109,19 @@ func LinearCombination(a *mat.Dense, b *mat.VecDense) *mat.VecDense {
 }
 
 // generate a normally distributed random n*1 vector
-func RandVec(n int) *mat.VecDense {
+func RandVec(n int, maxweight float64) *mat.VecDense {
 	data := make([]float64, n)
 	for i := range data {
-		data[i] = rand.NormFloat64()
+		data[i] = randFloat64(maxweight)
+	}
+	return mat.NewVecDense(n, data)
+}
+
+// generate a normally distributed random n*1 vector on natural numebrs
+func RandNatVec(n, max int) *mat.VecDense {
+	data := make([]float64, n)
+	for i := range data {
+		data[i] = float64(rand.Intn(max))
 	}
 	return mat.NewVecDense(n, data)
 }
@@ -120,11 +141,12 @@ func StringMat(a mat.Matrix) string {
 	return fmt.Sprintf("%.5g", mat.Formatted(a, mat.Squeeze()))
 }
 
+// print a matrix to stdout
 func PrintMat(a mat.Matrix) {
[…]

### 2020-10-05 main.go

@@ -5,9 +5,7 @@ import (
 	"math/rand"
 	"time"
 
-	"github.com/0x0f0f0f/lwa-techniques/automata"
-	"github.com/0x0f0f0f/lwa-techniques/lin"
-	"gonum.org/v1/gonum/mat"
+	"github.com/0x0f0f0f/lwa-techniques/randtest"
 )
 
 func check(err error) {
@@ -16,66 +14,21 @@ func check(err error) {
 	}
 }
 
-const tol = 10e-12
-
-func randTest(dim, numSyms, numSamples, maxweight int) (verified, bprt, hkct, null float64) {
-	az := automata.RandAutomaton(numSyms, dim, maxweight, tol)
-
-	az.BackwardsPartitionRefinement()
-
-	//fmt.Println(az)
-	lin.CleanTolDense(az.LLWBperp, tol)
-
-	samples := make([]*mat.VecDense, numSamples)
-	llwb := lin.Complement(az.LLWBperp, tol).(*mat.Dense)
-	_, dimLLWB := llwb.Dims()
-
-	lin.CleanTolDense(llwb, tol)
-
-	if mat.Equal(llwb, mat.NewDense(dim, 1, nil)) {
-		return 0, 0, 0, 1
-	}
-
-	//lin.PrintMat(llwb)
-
-	for i := range samples {
-		samples[i] = lin.LinearCombination(llwb, lin.RandVec(dimLLWB))
-	}
-
-	for i := range samples {
-		j := rand.Intn(numSamples)
-		for j == i {
-			j = rand.Intn(numSamples)
-		}
-		resBPR := az.BPREquivalence(samples[i], samples[j])
-		resHKC, _ := az.HKC(samples[i], samples[j], 30)
-
-		if resBPR {
-			bprt++
-		}
-
-		if resHKC {
-			hkct++
-		}
-
-		if resBPR == resHKC {
-			verified++
-			//fmt.Println(resBPR, resHKC)
-			//lin.PrintMat(samples[i])
-			//lin.PrintMat(samples[j])
-		}
-	}
-
-	return
-}
-
 func main() {
-	//defer profile.Start(profile.MemProfile).Stop()
+	// defer profile.Start(profile.MemProfile).Stop()
+
+	opts := randtest.TestOptions{
+		Dim:        4,
+		NumSymbols: 2,
+		NumSamples: 1000,
+		MaxWeight:  3,
+		Mode:       "nat",
+	}
 
[…]

### 2020-10-05 paper/paper.tex
first public release

@@
 \usepackage[T1]{fontenc}    % use 8-bit T1 fonts
-\usepackage{hyperref}       % hyperlinks
+\usepackage[hidelinks]{hyperref}       % hyperlinks
 \usepackage{url}            % simple URL typesetting
@@
 
-\title{DRAFT: Go Implementation of Up-To Techniques for Equivalence of Weighted Languages}
+\title{DRAFT: Go Implementation of Up-To Techniques for Weighted Language Equivalence}
 
@@
 and performance on randomly generated samples. 
-We finally provide comparison of runtime statistics and suggest 
-which of the two algorithm is the best choice for different usage cases.
+We finally provide a comparison and runtime statistics.
 \end{abstract}
@@
 
-\section{Implementation}
-\label{sec:impl}
+\input{sections/implementation.tex}
 
-The algorithms and data structures for this paper are implemented in the Go programming 
-language. This implementation makes use of the Gonum library for numerical 
-computations. We only import the Gonum libraries for matrices and linear algebra 
-and visual plotting of samples and functions.
-Real numbers are implemented with double precision floating point numbers,  
-known as the \texttt{float64} type in the Go programming language.
+\input{sections/runtime.tex}
 
-\begin{defn}
-  \textbf{Applications of SVD} \\
-
-  Let's consider the singular value decomposition of a matrix $A \in \R^{m \times n}$:
-
-  \begin{equation*}
-    \begin{aligned}
-      A = U \Sigma V^T & \quad & \Sigma = \diag{\sigma_1, \sigma_2, \hdots, \sigma_r  } 
-       & \quad &  U \in \R^{m \times m} & \quad & V \in \R^{n \times n}
-    \end{aligned}
-  \end{equation*}
-
-  Where $V$ and $U$ are orthogonal and the singular values are ordered: $\sigma_1 \geq \sigma_2 \geq \hdots \geq \sigma_r \geq 0$.
-  It follows that $\mrank{A}$ is equal to the number of nonzero singular values, and
-  as explained in \cite{svd}:
-
-  \begin{enumerate}
-    \item  $\mrank{A} = \mrank{\Sigma} = r$
-    \item The column space of $A$ is spanned by the first $r$ columns of $U$.
-    \item The null space of $A$ is spanned by the last $n − r$ columns of $V$.
-    \item The row space of $A$ is spanned by the first $r$ columns of $V$.
-    \item The null space of $A^T$ is spanned by the last $m − r$ columns of $U$.
-  \end{enumerate}
-
-  Of our interest, are only the computation of the null space and columns space.
-  The implementation, applying SVD, 
-  can be found in files \texttt{lin/colspace.go} and \texttt{lin/nullspace.go}.
-\end{defn}
-
-\subsection{Implementing the backwards partition refinement algorithm}
-
-To compute $\llwb$ at the last step of the algorithm,
-we need to compute $R_j^0$.
-If $V$ is a vector space and $W$ is a
-subspace of $W$, the annihilator of $W$, respectively $W^0$ is 
-a subspace of the space $V^*$ of linear functionals on $V$.
-$W^0$ are the functionals that annihilate on $W$. Since 
-we are working on subspaces of $\R^n$, we can directly compute 
-the orthogonal complement in our implementation instead of the
-annihilator.
-
-\begin{prop}
-  If $V$ is a finite dimensional vector space defined with an inner product
-  $\langle \cdot , \cdot \rangle$ and $W$ is a subspace of $V$
-  then the image of the annhilitaor $W^0$ through the linear 
-  isomorphism $\varphi: V^* \to V$ induced by the inner product, 
-  is the orthogonal of $W$ with respect to the said inner product.
-\end{prop}
-
-\begin{proof}
-  Let $V$ be an inner product space over the field $\K$ with an inner product defined as
-  $\langle \cdot , \cdot \rangle : V \times V \to \K$. 
-  Every linear functional can be 
-  represented with a vector. Let $\xi : V \to \K$ be a functional, 
-  $\xi \in  W^0$. Because $\xi(w)=0 \quad \forall  w \in W$, 
-  if $v$ represents $\xi$ we have that $(v, w)=\xi(w)=0$ for all $w \in W$. 
-  We obtain that $\varphi(W^0) \subseteq W^{\perp}$.
-  If $v \in W^\perp$  
-  then the functional $x \mapsto (v, x)$ cancels over $W$ 
-  (by the definition of orthogonality).
-\end{proof}
-
-To compute the orthogonal complement of a vector subspace $W$, we
-compute $W^\perp = \mker{A^T}$, where $A$ is the matrix with 
-column vectors in the spanning set of $W$ as its columns. Precisely, $W$ is 
-represented as the 
-column space of $A$. Proof is available in \cite{ila}.
+\input{sections/conclusion.tex}
 
@@
 \bibliographystyle{unsrt}
-\bibliography{references}  %%% Remove comment to use the external .bib file (using bibtex).
-%%% and comment out the ``thebibliography'' section.
+\bibliography{references}
 
-% 
-% %%% Comment out this section when you \bibliography{references} is enabled.
-% \begin{thebibliography}{1}
-% 
-% \bibitem{kour2014real}
-% George Kour and Raid Saabne.
-% \newblock Real-time segmentation of on-line handwritten arabic script.
-% \newblock In {\em Frontiers in Handwriting Recognition (ICFHR), 2014 14th
-%   International Conference on}, pages 417--422. IEEE, 2014.
-% 
-% \bibitem{kour2014fast}
-% George Kour and Raid Saabne.
-% \newblock Fast classification of handwritten on-line arabic characters.
-% \newblock In {\em Soft Computing and Pattern Recognition (SoCPaR), 2014 6th
-%   International Conference of}, pages 312--318. IEEE, 2014.
-% 
-% \bibitem{hadash2018estimate}
-% Guy Hadash, Einat Kermany, Boaz Carmeli, Ofer Lavi, George Kour, and Alon
-%   Jacovi.
-% \newblock Estimate and replace: A novel approach to integrating deep neural
-%   networks with existing applications.
-% \newblock {\em arXiv preprint arXiv:1804.09028}, 2018.
-% 
-% \end{thebibliography}
-% 
-% 
 \appendix

### 2020-10-05 paper/sections/algorithms.tex
first public release

@@
 The first algorithm we implement to compute language equivalence, called \texttt{HKC},
-is adapted from \cite{DBLP:journals/corr/Bonchi0K17}. It was first introduced by 
+is adapted from \cite{DBLP:journals/corr/Bonchi0K17}. 
+The algorithm returns \texttt{true} $\iff \sqmap{v_1} - \sqmap{v_2}$.
+It was first introduced by 
 Bonchi and Pous in \cite{bonchi2013checking}.
@@
 proven to be sound and complete \cite{DBLP:journals/corr/Bonchi0K17}.
-Originally, the algorithm was defined for 
-WAs over semirings, but we recall that in this work we are only 
+It is defined for 
+WAs over semirings, but in this implementation we are only 
 considering fields, in particular 
@@
 $L = (V, t, o)$,
-computes a basis of the subspace of $V$
+it computes a basis of the subspace of $V$
 representing the complementary relation of $\llwb$ (we later show it to be the 
@@
 $R_0$ in a \textit{backwards} direction. Intuitively, "going backwards" means 
-working with the transpose transitions functions $t_a^t$.
+working with the transpose transitions functions $t_a^T$.
 
-In the next section we compare execution results of our implementation of the algorithms
-\texttt{BPR} and \texttt{HKC} to verify the correctness of the implementation,
-and to analyze when one of the two algorithms is the convenient choice.
+In the next sections we will compare execution of our implementation of the algorithms
+\texttt{BPR} and \texttt{HKC} to verify correctness,
+and to provide insight on runtime results.
 Lemma \ref{lem:coincide}, introduced above, is key to our work. By stating that 
@@
 TODO costo computazionale HKC.
+
 \texttt{BPR} has a cost of $O(n^4)$ operations 
 to initially compute the largest linear weighted bisimulation,
-which can be eventually reduced to $O(n^3)$.
-In our implementation, by computing a basis of the orthogonal complement of $\llwb$,
-the cost of checking if two vectors are language equivalent is reduced to the
-cost of matrix multiplication ($O(n^2)$). \texttt{BPR} is efficient when we
+which can be eventually reduced to $O(n^3)$ \cite{BONCHI201277}.
+In our implementation, by initially computing a basis of the orthogonal complement of $\llwb$,
+the cost of checking if two vectors are language equivalent is then reduced to the
+cost of matrix multiplication ($O(n^2)$). \texttt{BPR} is a great choice when we
 have to decide if a large number of vectors in a WA are language equivalent.
@@
 \begin{eqnarray}
-  R_0 = \mker{o}^0, & \quad & R_{i+1} = R_i + \sum_{a \in A} t_a^t(R_i) \label{back} 
+  R_0 = \mker{o}^0, & \quad & R_{i+1} = R_i + \sum_{a \in A} t_a^T(R_i) \label{back} 
 \end{eqnarray}

### 2020-10-05 paper/sections/conclusion.tex
first public release

\section{Conclusion and Future Work}

%TODO 
TODO da fare

### 2020-10-05 paper/sections/implementation.tex
first public release

\section{Implementation}
\label{sec:impl}

The algorithms and data structures are implemented in the Go programming 
language. Real numbers are implemented with double precision floating point numbers,  
precisely of \texttt{float64} type.

This implementation makes use of the Gonum library, 
an excellent toolkit for high-performance numerical computations.
We only import the Gonum libraries for matrices, linear algebra 
and visual plotting.
Although not GPU accelerated, Gonum matrix operations are run on 
the CPU and accelerated with BLAS and LAPACK.

\subsection{Data Structures}
In this implementation, the data structure for representing weighted automata is a \texttt{struct} containing
the following attributes:
\begin{enumerate}
    \item An integer \texttt{Dim} representing the number of states $n$. 
    (note: states are finite and indexed on natural numbers $0, \hdots, n-1$)
    \item A slice of strings representing the \textit{alphabet} $A$.
    \item A map of strings as keys (the alphabet symbols) and dense $n \times n$ \texttt{float64} matrices as values,
    representing the set of transition functions.
    \item A dense \texttt{float64} vector, representing the linearization of the output function. 
    \item A \texttt{float64} value representing the tolerance to be used in numerical computations.
    \item An optionally \texttt{nil}, dense floating point matrix providing a basis for the orthogonal 
    complement of $\llwb$
\end{enumerate}
\begin{note}
    \textit{Slices} in Go are a convenient and efficient extension of the concept of arrays: 
    they provide an abstraction for indexed, variable length sequences of typed data, and 
    provide useful helper functions for creating, appending and selecting elements. 
\end{note}
The definition can be found in file \texttt{automata/automata.go}.
We then provide methods for reading an automaton from a text stream, applying transition
and output functions to vectors, and generating random automata with real and natural valued weights.

\subsection{Implementation of HKC}
Instead of creating dedicated structs,
we exploit the \texttt{mat.Dense} data type in Gonum to efficiently represent:
\begin{itemize}
    \item Sets of vectors with $n \times k$ dense matrices ($k$ is the number of vectors in the set)
    \item Pairs of vectors with $n \times 2$ dense matrices
    \item Sets and the "\texttt{todo}" stack of pairs with $n \times 2k$ dense matrices ($k$ is the number of pairs in the set or stack)
\end{itemize}

To increase efficiency of the methods for inclusion checking and insertion in sets of vectors,
one could keep the columns ordered (by vector norm) in the corresponding matrix.

To represent the congruence relation $R$, we introduce a \texttt{struct} containing:
\begin{itemize}
    \item A dense matrix \texttt{s} of size $n \times 2k$ containing the set of pairs in the relation
    \item A dense matrix \texttt{u} to represent the generating set $U_R$ for the congruence closure (see definition \ref{def:congclos}).
    \item Two integers representing the number of pairs in the set, and the size of vectors.
    \item A tolerance value to be used in equivalence checks. Best results were obtained with $10^{-14}$.
\end{itemize}

When adding a pair of vectors to the congruence relation, we extend the columns
of the matrix \texttt{s} with the pair $(v_1, v_2)$, if and only if $(v_1 - v_2)$ 
is not already in $U$. To check if the pair is in the congruence closure $c(R)$, 
we check if $(v_1 - v_2)$ is contained in U.

\captionof{figure}{Implementation of the \texttt{HKC}$(v_1,v_2)$ procedure}
\lstinputlisting[language=go]{../automata/hkc.go}

%TODO sistema

\subsection{Implementation of BPR}
Given a WA $L$, at the first step of \texttt{BPR}, we set $R_0 = o$, 
with $o$ being the dense vector representing the output function 
of $L$, as seen in \cite{BONCHI201277}.
For each step $i = 0, \hdots, j+1$, with $j \leq \dim(L)$, the implemented algorithm:
\begin{enumerate}
    \item For each $a \in A$ computes $t_a^T(R_i)$ through matrix multiplication
    \item Concatenates those matrices $t_a^T(R_i)$ to $R_i$ in a resulting matrix $G$ 
    \item Instead of checking linear independence of the columns through Gaussian elimination,
    it computes $R_{i+1}$ as the orthonormal basis of the column space of $G$, through singular 
    value decomposition.
\end{enumerate}

At the end of \texttt{BPR}, we store the basis for the orthogonal complement
of $\llwb$ as an attribute of the automaton. To check if two vectors $v_1 \llwb v_2$,
we check that $R_j^T(v_1 - v_2) = \vec{0}$, with a prefixed tolerance.

To compute a basis for $\llwb$, at the last step of the algorithm,
we would need to compute $R_j^0$.
If $V$ is a vector space and $W$ is a
subspace of $W$, the annihilator of $W$, respectively $W^0$ is 
a subspace of the space $V^*$ of linear functionals on $V$.
$W^0$ are the functionals that annihilate on $W$. Since 
we are working on subspaces of $\R^n$, we can directly compute 
the orthogonal complement in our implementation instead of the
annihilator.

\begin{prop}
  If $V$ is a finite dimensional vector space defined with an inner product
  $\langle \cdot , \cdot \rangle$ and $W$ is a subspace of $V$
  then the image of the annhilator $W^0$ through the linear 
  isomorphism $\varphi: V^* \to V$ induced by the inner product, 
  is the orthogonal of $W$ with respect to the said inner product.
\end{prop}

\begin{proof}
  Let $V$ be an inner product space over the field $\K$ with an inner product defined as
  $\langle \cdot , \cdot \rangle : V \times V \to \K$. 
  Every linear functional can be 
  represented with a vector. Let $\xi : V \to \K$ be a functional, 
  $\xi \in  W^0$. Because $\xi(w)=0 \quad \forall  w \in W$, 
  if $v$ represents $\xi$ we have that $(v, w)=\xi(w)=0$ for all $w \in W$. 
  We obtain that $\varphi(W^0) \subseteq W^{\perp}$.
  If $v \in W^\perp$  
  then the functional $x \mapsto (v, x)$ cancels over $W$ 
  (by the definition of orthogonality).
\end{proof}

To compute the orthogonal complement of a vector subspace $W$, we
compute $W^\perp = \mker{A^T}$, where $A$ is the matrix whose column space 
is $W$. Proof is available in \cite{ila}.

\captionof{figure}{Implementation of Backwards Partition Refinement}
\lstinputlisting[language=go]{../automata/backwards.go}

\begin{note}
    \textbf{Applications of SVD} \\

    Let's consider the singular value decomposition of a matrix $A \in \R^{m \times n}$:

    \begin{equation*}
      \begin{aligned}
        A = U \Sigma V^T & \quad & \Sigma = \diag{\sigma_1, \sigma_2, \hdots, \sigma_r  } 
         & \quad &  U \in \R^{m \times m} & \quad & V \in \R^{n \times n}
      \end{aligned}
    \end{equation*}

    Where $V$ and $U$ are orthogonal and the singular values are ordered: $\sigma_1 \geq \sigma_2 \geq \hdots \geq \sigma_r \geq 0$.
    It follows that $\mrank{A}$ is equal to the number of nonzero singular values, and
    as explained in \cite{svd}:

    \begin{enumerate}
      \item  $\mrank{A} = \mrank{\Sigma} = r$
      \item The column space of $A$ is spanned by the first $r$ columns of $U$.
      \item The null space of $A$ is spanned by the last $n − r$ columns of $V$.
      \item The row space of $A$ is spanned by the first $r$ columns of $V$.
      \item The null space of $A^T$ is spanned by the last $m − r$ columns of $U$.
    \end{enumerate}

    Of our interest, are only the computation of the null space and column space.
    The implementation 
    can be found in files \texttt{lin/colspace.go} and \texttt{lin/nullspace.go}.
  \end{note}

### 2020-10-05 paper/sections/runtime.tex
first public release

\section{Runtime Results}

TODO da fare e sistemare.

The program was executed on an x86\_64 AMD Ryzen 5 2600X Six Core CPU, running at 3.6GHz with 32GB RAM, 
running Void Linux. 
For 3 test runs on 100000 automata composed of 4 states, 2 alphabet symbols and transition
matrices with weights on natural numbers $\leq 3$, with a tolerance of 10e-14,
we found a non empty kernel of $\llwb$ for an average of 1718 automata.

%TODO
(TODO spiega: più aumentano i simboli/stati più si riduce la probabilità di 
trovare $\llwb$ non vuoto?? investiga, trova un modo di generare automi con LLWB 
non vuoto)

By generating 1000 random couples of language equivalent vectors $(v_1, v_2)$ such that
$(v_1 - v_2) \in \mker{\llwb}$ and 1000 couples of completely random vectors
$(w_1, w_2)$, for each automata, we verified with 100\% confidence that:

\begin{center}
    $v_1 \; R_j \; v_2 \iff$ \texttt{HKC}$(v_1, v_2) =$ \texttt{false}
\end{center}

Where $R_j$ is the largest linear weighted bisimulation $\llwb$ computed by \texttt{BPR}.

%TODO

Sample output (19992 KBs in memory):
\begin{verbatim}
    $ time go run main.go 
    tests completed            100000...

    1000 samples per automaton
    100000 automata
    LLWP is not empty for 1726 automata
    language equivalence HKC = BPR for 3452000 samples
    BPR equivalence verified for 1726000 samples
    HKC equivalence verified for 1726000 samples
    confidence is 100%

    real    0m21.322s
    user    0m23.614s
    sys     0m0.961s
\end{verbatim}

### 2020-10-05 automata/backwards.go

@@ -20,12 +20,12 @@ func (a *Automaton) BackwardsPartitionRefinement() {
 	// lastIndex := 0
 
 	for i := 1; i <= a.Dim; i++ {
-		// Σ_{a \in A} T_a^t(R_i)
+		// \sum_{a \in A} T_a^T(R_i)
 		for _, sym := range a.A {
 			newBasis := a.ApplyTransposeTransitionBasis(sym, lastBasis)
 			currBasis = lin.Union(currBasis, newBasis)
-			currBasis = lin.OrthonormalColumnSpaceBasis(currBasis, a.Tol).(*mat.Dense)
 		}
+		currBasis = lin.OrthonormalColumnSpaceBasis(currBasis, a.Tol).(*mat.Dense)
 
 		lastBasis = currBasis
 	}
@@ -35,6 +35,8 @@ func (a *Automaton) BackwardsPartitionRefinement() {
 	// a.LLWB = lin.Complement(currBasis).(*mat.Dense)
 }
 
+// method that, after an LLWB is computed through BPR,
+// checks the equivalence of 2 vectors
 func (a Automaton) BPREquivalence(v1, v2 *mat.VecDense) bool {
 	if a.LLWBperp == nil {
 		log.Fatalln("largest linear weighted bisimulation not computed for automaton")

### 2020-10-05 automata/hkc.go

@@ -1,3 +1,5 @@
+// this file contains the implementation of the HKC procedure
+
 package automata
 
 import (
@@ -5,7 +7,7 @@ import (
 )
 
 // checks the language equivalence of two state vectors for a given weighted automaton
-func (a Automaton) HKC(v1, v2 *mat.VecDense, maxiter int) (bool, error) {
+func (a Automaton) HKC(v1, v2 *mat.VecDense) (bool, error) {
 	rel := NewRelation(0, a.Dim)
 	todo := NewPairStack()
 
@@ -17,7 +19,6 @@ func (a Automaton) HKC(v1, v2 *mat.VecDense, maxiter int) (bool, error) {
 	// insert (v1, v2) into the todo list
 	todo = PairStackPush(todo, p)
 
-	i := 0
 	for !todo.IsEmpty() {
 		// extract (v1', v2') from todo
 		q, err := PairStackPop(todo)
@@ -49,9 +50,6 @@ func (a Automaton) HKC(v1, v2 *mat.VecDense, maxiter int) (bool, error) {
 
 		// insert (v1', v2') into R
 		rel.Add(q)
-
-		// fmt.Println("R = ", rel)
-		i++
 	}
 
 	return true, nil

### 2020-10-05 main.go

@@ -52,14 +52,14 @@ func main() {
 	notnull := float64(numTests) - totnull
 
 	fmt.Println()
-	fmt.Println(numSamples, "samples per test")
-	fmt.Println(numTests, "tests")
-	fmt.Println("LLWP is not empty in", notnull, "tests")
-	fmt.Println("language equivalence HKC = BPR for", totalVerified, "samples")
-	percent := (totalVerified * 100) / (notnull * float64(numSamples))
+	fmt.Println(numSamples, "samples per automaton")
+	fmt.Println(numTests, "automata")
+	fmt.Println("LLWP is not empty for", notnull, "automata")
+	fmt.Printf("language equivalence HKC = BPR for %d samples\n", int(totalVerified))
+	percent := (totalVerified * 100) / (notnull * 2 * float64(numSamples))
 
-	fmt.Println("BPR equivalence verified for ", totbpr, "samples")
-	fmt.Println("HKC equivalence verified for ", tothkc, "samples")
-	fmt.Printf("total success rate is %.20g%%\n", percent)
+	fmt.Printf("BPR equivalence verified for %d samples\n", int(totbpr))
+	fmt.Printf("HKC equivalence verified for %d samples\n", int(tothkc))
+	fmt.Printf("confidence is %.20g%%\n", percent)
 
 }

### 2020-10-05 randtest/randtest.go

@@ -74,12 +74,32 @@ func RandTest(o TestOptions) TestResults {
 	}
 
 	for i := range samples {
+		// test for vectors in span of LLWB
 		j := rand.Intn(o.NumSamples)
 		for j == i {
 			j = rand.Intn(o.NumSamples)
 		}
 		resBPR := az.BPREquivalence(samples[i], samples[j])
-		resHKC, _ := az.HKC(samples[i], samples[j], 30)
+		resHKC, _ := az.HKC(samples[i], samples[j])
+
+		if resBPR {
+			results.Bprt++
+		}
+
+		if resHKC {
+			results.Hkct++
+		}
+
+		if resBPR == resHKC {
+			results.Verified++
+		}
+
+		// test for totally random vectors
+		w1 := lin.RandVec(az.Dim, 100)
+		w2 := lin.RandVec(az.Dim, 100)
+
+		resBPR = az.BPREquivalence(w1, w2)
+		resHKC, _ = az.HKC(w1, w2)
 
 		if resBPR {
 			results.Bprt++

### 2020-10-07 paper/paper.tex
updated tests with useful parameters

@@
 
-\title{DRAFT: Go Implementation of Up-To Techniques for Weighted Language Equivalence}
+\title{DRAFT: Go Implementation of Techniques for Weighted Language Equivalence}
 
@@
 a quantity expressing the weight (or probability) of the execution of each transition.
-In this work we propose an implementation of two algorithms for computing language 
-equivalence in finite state weighted automata (WAs). The first, a
-linear partition refinement algorithm, computes the largest linear weighted bisimulation
-for any given LWA (Linear Weighted Automaton) through an iterative method, 
-the second algorithm checks the language equivalence 
-of two vectors (states) for a given weighted automata by using an additional
+In this work we propose an implementation of two algorithms for computing
+the language equivalence relation in finite state weighted automata (WAs). 
+The first algorithm checks the language equivalence 
+of two vectors (states) in a weighted automaton, 
+with an up-to technique by using an additional
 data structure representing a congruence relation.
+The second algorithm, a
+linear partition refinement algorithm, computes the 
+whole equivalence relation, precisely the
+largest linear weighted bisimulation,
+by linearizing the state space and iteratively refining
+the relation. 
 We then compare results of the two algorithms to verify their correctness

### 2020-10-07 paper/sections/algorithms.tex
updated tests with useful parameters

@@
 is adapted from \cite{DBLP:journals/corr/Bonchi0K17}. 
-The algorithm returns \texttt{true} $\iff \sqmap{v_1} - \sqmap{v_2}$.
+The algorithm returns \texttt{true} $\iff \sqmap{v_1} = \sqmap{v_2}$.
 It was first introduced by 
@@
 
-% TODO
-(TODO chiedi a filippo: nelle conclusioni di \cite{BONCHI201277} dice che è diverso 
-dall'algoritmo visto nel paper di boreale \cite{boreale2009weighted}. è il caso 
-anche per i field e i numeri reali???? o sono lo equivalenti?).
+Adapted from \cite{BONCHI201277}, 
+\texttt{BPR}, or \textit{Backwards Partition Refinement}
+is a fixed-point iterative method that, given an LWA
+$L = (V, t, o)$, it computes the whole
 
-Adapted from \cite{BONCHI201277}, 
-\texttt{BPR} is a fixed-point iterative method that, given an LWA
-$L = (V, t, o)$,
-it computes a basis of the subspace of $V$
+computes a basis of the subspace of $V$
 representing the complementary relation of $\llwb$ (we later show it to be the 
@@
 called \textit{Forward Partition Refinement}, which directly computes
-a basis for $\llwb$
-but is shown to be way less efficient than the backwards version.
+a basis for $\llwb$ but is shown to be way less efficient than the backwards version.
+The advantage of the backwards
+It has been first introduced in \cite{boreale2009weighted}.
+Our implementation is directly modeled on the algorithm shown in 
+\cite{boreale2009weighted}, since we are fixing the field 
+to $\R$ and computing the orthogonal complements instead of 
+dual spaces and annihilators. In \cite{BONCHI201277}, the authors 
+extended.
 

### 2020-10-07 automata/automata.go

@@ -24,13 +24,14 @@ type Automaton struct {
 	// with LLWBperp, we denote a basis of the orthogonal
 	// complement of the basis LLWB
 	LLWBperp *mat.Dense
-	Tol      float64 // tolerance value
+	BPRTol   float64 // tolerance value for BPR
+	HKCTol   float64 // tolerance value for HKC
 }
 
 // applies the output function to a given state vector (o * v)
-func (a Automaton) GetOutput(v *mat.VecDense) float64 {
+func (a Automaton) GetOutput(v *mat.VecDense, tol float64) float64 {
 	res := mat.Dot(a.O, v)
-	if math.Abs(res) < a.Tol {
+	if math.Abs(res) < tol {
 		res = 0.0
 	}
 	return res

### 2020-10-07 automata/backwards.go

@@ -25,7 +25,7 @@ func (a *Automaton) BackwardsPartitionRefinement() {
 			newBasis := a.ApplyTransposeTransitionBasis(sym, lastBasis)
 			currBasis = lin.Union(currBasis, newBasis)
 		}
-		currBasis = lin.OrthonormalColumnSpaceBasis(currBasis, a.Tol).(*mat.Dense)
+		currBasis = lin.OrthonormalColumnSpaceBasis(currBasis, a.BPRTol).(*mat.Dense)
 
 		lastBasis = currBasis
 	}
@@ -50,5 +50,5 @@ func (a Automaton) BPREquivalence(v1, v2 *mat.VecDense) bool {
 	mul.Reset()
 	mul.MulVec(a.LLWBperp.T(), sub)
 
-	return lin.IsZeroTol(mul, a.Tol)
+	return lin.IsZeroTol(mul, a.BPRTol)
 }

### 2020-10-07 automata/hkc.go

@@ -30,8 +30,8 @@ func (a Automaton) HKC(v1, v2 *mat.VecDense) (bool, error) {
 			continue
 		}
 
-		o1 := a.GetOutput(PairLeft(q))
-		o2 := a.GetOutput(PairRight(q))
+		o1 := a.GetOutput(PairLeft(q), a.HKCTol)
+		o2 := a.GetOutput(PairRight(q), a.HKCTol)
 		if o1 != o2 {
 			return false, nil
 		}

### 2020-10-07 automata/random.go

@@ -8,7 +8,7 @@ import (
 )
 
 // generate a random automaton on natural numbers
-func RandNatAutomaton(syms, states, maxweight int, tol float64) Automaton {
+func RandNatAutomaton(syms, states, maxweight int) Automaton {
 	// create the alphabet
 	A := make([]string, syms)
 	if syms <= 57 {
@@ -33,14 +33,13 @@ func RandNatAutomaton(syms, states, maxweight int, tol float64) Automaton {
 		T:   T,
 		O:   O,
 		Dim: states,
-		Tol: tol,
 	}
 
 	return aut
 }
 
 // generate a random automaton on real numbers
-func RandAutomaton(syms, states int, maxweight float64, tol float64) Automaton {
+func RandAutomaton(syms, states int, maxweight float64) Automaton {
 	// create the alphabet
 	A := make([]string, syms)
 	if syms <= 57 {
@@ -65,7 +64,6 @@ func RandAutomaton(syms, states int, maxweight float64, tol float64) Automaton {
 		T:   T,
 		O:   O,
 		Dim: states,
-		Tol: tol,
 	}
 
 	return aut

### 2020-10-07 automata/relation.go

@@ -78,7 +78,7 @@ func (r *Relation) Add(p *mat.Dense) {
 	_, un := r.u.Dims()
 	for j := 0; j < un; j++ {
 		v := r.u.ColView(j).(*mat.VecDense)
-		if lin.EqVecTol(v, sub, r.tol) {
+		if mat.EqualApprox(v, sub, r.tol) {
 			subInU = true
 		}
 	}
@@ -101,7 +101,7 @@ func (r Relation) PairIsInCongruenceClosure(p *mat.Dense) bool {
 	_, un := r.u.Dims()
 	for j := 0; j < un; j++ {
 		v := r.u.ColView(j).(*mat.VecDense)
-		if lin.EqVecTol(v, sub, r.tol) {
+		if mat.EqualApprox(v, sub, r.tol) {
 			return true
 		}
 	}

### 2020-10-07 lin/colspace.go

@@ -15,7 +15,10 @@ func OrthonormalColumnSpaceBasis(a mat.Matrix, tol float64) mat.Matrix {
 	}
 	u := mat.NewDense(1, 1, nil)
 	u.Reset()
+	//fmt.Println(svd.Cond())
 	svd.UTo(u)
+	//fmt.Println(mat.Cond(a, 2))
+	//fmt.Println(mat.Cond(u, 2))
 
 	// The column space of A is spanned by the first r columns of U.
 	j := 0

### 2020-10-07 lin/gauss_test.go

@@ -1,33 +1,0 @@
-package lin
-
-import (
-	"log"
-	"math"
-	"testing"
-)
-
-var tc = testCase{
-	a: [][]float64{
-		{1.00, 0.00, 0.00, 0.00, 0.00, 0.00},
-		{1.00, 0.63, 0.39, 0.25, 0.16, 0.10},
-		{1.00, 1.26, 1.58, 1.98, 2.49, 3.13},
-		{1.00, 1.88, 3.55, 6.70, 12.62, 23.80},
-		{1.00, 2.51, 6.32, 15.88, 39.90, 100.28},
-		{1.00, 3.14, 9.87, 31.01, 97.41, 306.02}},
-	b: []float64{-0.01, 0.61, 0.91, 0.99, 0.60, 0.02},
-	x: []float64{-0.01, 1.602790394502114, -1.6132030599055613,
-		1.2454941213714368, -0.4909897195846576, 0.065760696175232},
-}
-
-func TestGaussPartial(t *testing.T) {
-	x, err := GaussPartialo(tc.a, tc.b)
-	if err != nil {
-		log.Fatal(err)
-	}
-	for i, xi := range x {
-		if math.Abs(tc.x[i]-xi) > ε {
-			log.Println("out of tolerance")
-			t.Errorf("expected %.20g", tc.x)
-		}
-	}
-}

### 2020-10-07 lin/nullspace_test.go

@@ -15,7 +15,7 @@ func TestNullspace(t *testing.T) {
 		4.0, 5.0, 6.0,
 		7.0, 8.0, 9.0,
 	})
-	ker, res := Nullspace(a)
+	ker, res := Nullspace(a, test_tol)
 
 	//fmt.Printf("%.60e\n", res)
 
@@ -38,7 +38,7 @@ func TestNullspace2(t *testing.T) {
 			1, 7, 2,
 			-2, 3, 1,
 		})
-	ker, _ := Nullspace(a)
+	ker, _ := Nullspace(a, test_tol)
 
 	expker := mat.NewDense(3, 1,
 		[]float64{

### 2020-10-07 lin/subspace_test.go

@@ -7,6 +7,8 @@ import (
 	"gonum.org/v1/gonum/mat"
 )
 
+const test_tol = 10e-14
+
 func TestInSubspace(t *testing.T) {
 	m := mat.NewDense(3, 1, []float64{1, 1, 0})
 
@@ -16,8 +18,8 @@ func TestInSubspace(t *testing.T) {
 
 	nv := mat.NewVecDense(3, []float64{3, 3, 1})
 
-	assert.True(t, InSubspace(m, v))
-	assert.False(t, InSubspace(m, nv))
+	assert.True(t, InSubspace(m, v, test_tol))
+	assert.False(t, InSubspace(m, nv, test_tol))
 
 	// matrix is not modified
 	assert.True(t, mat.Equal(m, om))
@@ -40,7 +42,7 @@ func TestIntersect(t *testing.T) {
 		-1, -1, 1,
 	})
 
-	base := Intersect(u, w)
+	base := Intersect(test_tol, u, w)
 
 	expected := mat.NewDense(5, 1, []float64{
 		-0.447213595499957927703604809721582569181919097900390625,
@@ -67,7 +69,7 @@ func TestComplement(t *testing.T) {
 		-0.4082482904638629062077370690531097352504730224609375,
 	})
 
-	wperp := Complement(wspan)
+	wperp := Complement(wspan, test_tol)
 
 	assert.True(t, mat.Equal(expected, wperp))
 

### 2020-10-07 lin/util.go

@@ -34,14 +34,6 @@ func IsZeroTol(vec *mat.VecDense, tol float64) bool {
 	return true
 }
 
-// returns true if two vectors are equal with a tolerance of tol
-func EqVecTol(a, b mat.Vector, tol float64) bool {
-	sub := mat.VecDenseCopyOf(a)
-	sub.SubVec(a, b)
-
-	return IsZeroTol(sub, tol)
-}
-
 // create an n*n identity matrix
 func EyeDense(n int) *mat.Dense {
 	a := mat.NewDense(n, n, nil)

### 2020-10-07 randtest/data.go

@@ -1,0 +1,117 @@
+// this file contains data structures relevant to tests
+
+package randtest
+
+import "fmt"
+
+// Represent a sample result
+const (
+	TP int = iota // true positive
+	TN            // true negative
+	FP            // false positive
+	FN            // false negative
+)
+
+type BatchTestOptions struct {
+	AutOptions  *AutomatonTestOptions // initial automaton test options
+	NumAutomata int                   // number of automata to generate and test
+}
+
+func (opt BatchTestOptions) Print() {
+	fmt.Println("========= BATCH OPTIONS ===========")
+	fmt.Println(opt.NumAutomata, "automata")
+	opt.AutOptions.Print()
+}
+
+// BatchResult represents the result of many weighted
+// language equivalence sample tests on many automata
+type BatchResult struct {
+	TP        float64
+	TN        float64
+	FP        float64
+	FN        float64
+	Null      float64 // number of automata where ker(LLWB) is null
+	Total     float64 // Total number of tested automata
+	Accuracy  float64 // (TP+TN)/(TP+TN+FP+FN). percent of correctness
+	Recall    float64
+	Precision float64
+	F1        float64
+	opt       *BatchTestOptions
+}
+
+// Accumulate adds results of an automaton test to the results of a batch test
+func (r *BatchResult) Accumulate(ar AutomatonResult) {
+	r.TP += ar.TP
+	r.TN += ar.TN
+	r.FP += ar.FP
+	r.FN += ar.FN
+	if ar.Null {
+		r.Null++
+	}
+	r.Total++
+}
+
+// ComputeStats computes relevant statistics on a batch test
+func (r *BatchResult) ComputeStats() {
+	T := r.TP + r.TN
+	r.Accuracy = T / (T + r.FP + r.FN)
+	r.Recall = r.TP / (r.TP + r.FN)
+	r.Precision = r.TP / (r.TP + r.FP)
+	r.F1 = 2 * ((r.Precision * r.Recall) / (r.Precision + r.Recall))
+}
+
+func (r BatchResult) Print() {
+	r.opt.Print()
+	fmt.Println("========= RESULTS ===========")
+	fmt.Println("LLWP is not empty for", float64(r.opt.NumAutomata)-r.Null, "automata")
+	fmt.Printf("TP: %10d TN: %10d\n", int(r.TP), int(r.TN))
[…]

### 2020-10-07 randtest/randtest.go

@@ -2,6 +2,7 @@ package randtest
 
 import (
 	"errors"
+	"fmt"
 	"math/rand"
 
 	"github.com/0x0f0f0f/lwa-techniques/automata"
@@ -9,110 +10,88 @@ import (
 	"gonum.org/v1/gonum/mat"
 )
 
-const tol = 10e-12
+func BatchTest(opt *BatchTestOptions) BatchResult {
+	batchResults := BatchResult{opt: opt}
 
-type TestOptions struct {
-	Dim        int
-	NumSymbols int
-	NumSamples int
-	MaxWeight  int
-	Mode       string
+	for i := 0; i < opt.NumAutomata; i++ {
+		fmt.Printf("testing automata %20d...\r", i)
+
+		batchResults.Accumulate(TestRandAutomaton(opt.AutOptions))
+	}
+
+	fmt.Println()
+	batchResults.ComputeStats()
+	return batchResults
+
 }
 
-type TestResults struct {
-	// number of samples on which language equivalence
-	// is verified by both HKC and BPR
-	Verified float64
-	// number of samples on which bpr equivalence is verified
-	Bprt float64
-	// number of samples on which hkc equivalence is verified
-	Hkct float64
-	// true if the LLWB computed by BPR is empty
-	Null bool
-}
-
-func RandTest(o TestOptions) TestResults {
+func TestRandAutomaton(o *AutomatonTestOptions) AutomatonResult {
 	var az automata.Automaton
 
 	switch o.Mode {
 	case "real":
-		az = automata.RandAutomaton(o.NumSymbols, o.Dim, float64(o.MaxWeight), tol)
+		az = automata.RandAutomaton(o.NumSymbols, o.NumStates, float64(o.MaxWeight))
 
 	case "nat":
-		az = automata.RandNatAutomaton(o.NumSymbols, o.Dim, o.MaxWeight, tol)
+		az = automata.RandNatAutomaton(o.NumSymbols, o.NumStates, o.MaxWeight)
 
 	default:
 		panic(errors.New("unknown mode"))
 	}
+	az.BPRTol = o.BPRTol
+	az.HKCTol = o.HKCTol
 
 	// fmt.Println(az)
 	az.BackwardsPartitionRefinement()
 
-	lin.CleanTolDense(az.LLWBperp, tol)
-
 	samples := make([]*mat.VecDense, o.NumSamples)
[…]

### 2020-10-08 paper/automata.tex
refined f1-tolerance test, working on paper

\tikzset{every picture/.style={line width=0.75pt}} %set default line width to 0.75pt

\begin{tikzpicture}[x=0.75pt,y=0.75pt,yscale=-1,xscale=1]
%uncomment if require: \path (0,514); %set diagram left start at 0, and has height of 514

%Curve Lines [id:da7520674866212254] 
\draw    (207.69,98.63) .. controls (223.51,60.79) and (169.45,55.3) .. (187,101.61) ;
\draw [shift={(206.33,101.61)}, rotate = 296.28] [fill={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.08]  [draw opacity=0] (8.93,-4.29) -- (0,0) -- (8.93,4.29) -- cycle    ;
%Straight Lines [id:da36516510379788225] 
\draw    (139.5,236.5) -- (139.5,261.86) ;
\draw [shift={(139.5,264.86)}, rotate = 270] [fill={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.08]  [draw opacity=0] (8.93,-4.29) -- (0,0) -- (8.93,4.29) -- cycle    ;
%Straight Lines [id:da06517112562563687] 
\draw    (249.5,236) -- (249.5,261.36) ;
\draw [shift={(249.5,264.36)}, rotate = 270] [fill={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.08]  [draw opacity=0] (8.93,-4.29) -- (0,0) -- (8.93,4.29) -- cycle    ;
%Straight Lines [id:da05488487732501923] 
\draw    (211,114) -- (231.4,114) ;
\draw [shift={(234.4,114)}, rotate = 180] [fill={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.08]  [draw opacity=0] (8.93,-4.29) -- (0,0) -- (8.93,4.29) -- cycle    ;

% Text Node
\draw    (196.5, 113.15) circle [x radius= 14.14, y radius= 15.56]   ;
\draw (196.5,113.15) node  [font=\normalsize]  {$x_{1}$};
% Text Node
\draw    (140.18, 220.54) circle [x radius= 14.14, y radius= 15.56]   ;
\draw (140.18,220.54) node  [font=\normalsize]  {$x_{2}$};
% Text Node
\draw    (249.18, 220.54) circle [x radius= 14.14, y radius= 15.56]   ;
\draw (249.18,220.54) node  [font=\normalsize]  {$x_{3}$};
% Text Node
\draw (178,249.9) node [anchor=north west][inner sep=0.75pt]  [font=\footnotesize]  {$b,1/3$};
% Text Node
\draw (186,178.9) node [anchor=north west][inner sep=0.75pt]  [font=\footnotesize]  {$b,3$};
% Text Node
\draw (236.5,152.9) node [anchor=north west][inner sep=0.75pt]  [font=\footnotesize]  {$a,1$};
% Text Node
\draw (172,44.4) node [anchor=north west][inner sep=0.75pt]  [font=\footnotesize]  {$a,1/b,1$};
% Text Node
\draw (120,152.9) node [anchor=north west][inner sep=0.75pt]  [font=\footnotesize]  {$a,1/3$};
% Text Node
\draw (122,242.65) node [anchor=north west][inner sep=0.75pt]  [font=\footnotesize]  {$1$};
% Text Node
\draw (262,242.65) node [anchor=north west][inner sep=0.75pt]  [font=\footnotesize]  {$1$};
% Text Node
\draw (236.4,117.4) node [anchor=north west][inner sep=0.75pt]  [font=\footnotesize]  {$2$};
% Connection
\draw    (154.02,223.76) .. controls (181.71,243.14) and (208.08,243.86) .. (233.12,225.93) ;
\draw [shift={(235.43,224.21)}, rotate = 502.59] [fill={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.08]  [draw opacity=0] (8.93,-4.29) -- (0,0) -- (8.93,4.29) -- cycle    ;
% Connection
\draw    (235.45,216.79) .. controls (208,196.14) and (181.57,195.41) .. (156.17,214.59) ;
\draw [shift={(153.82,216.42)}, rotate = 321.14] [fill={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.08]  [draw opacity=0] (8.93,-4.29) -- (0,0) -- (8.93,4.29) -- cycle    ;
% Connection
\draw    (147.25,207.06) -- (188.04,129.28) ;
\draw [shift={(189.43,126.63)}, rotate = 477.67] [fill={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.08]  [draw opacity=0] (8.93,-4.29) -- (0,0) -- (8.93,4.29) -- cycle    ;
% Connection
\draw    (242.46,206.85) -- (204.54,129.53) ;
\draw [shift={(203.22,126.84)}, rotate = 423.87] [fill={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.08]  [draw opacity=0] (8.93,-4.29) -- (0,0) -- (8.93,4.29) -- cycle    ;

\end{tikzpicture}

### 2020-10-08 paper/sections/algorithms.tex
refined f1-tolerance test, working on paper

@@
 \cite{hopcroft1971linear} with \textit{congruence closure}, is 
-proven to be sound and complete \cite{DBLP:journals/corr/Bonchi0K17}.
+proven to be sound and complete.
 It is defined for 
@@
 \textit{Backwards Partition Refinement}, that we will call BPR for short. 
-
 Adapted from \cite{BONCHI201277}, 
-\texttt{BPR}, or \textit{Backwards Partition Refinement}
+\texttt{BPR}
 is a fixed-point iterative method that, given an LWA
-$L = (V, t, o)$, it computes the whole
-
-computes a basis of the subspace of $V$
+$L = (V, t, o)$, computes a basis of the subspace of $V$
 representing the complementary relation of $\llwb$ (we later show it to be the 
@@
 a basis for $\llwb$ but is shown to be way less efficient than the backwards version.
-The advantage of the backwards
-It has been first introduced in \cite{boreale2009weighted}.
+
 Our implementation is directly modeled on the algorithm shown in 
-\cite{boreale2009weighted}, since we are fixing the field 
-to $\R$ and computing the orthogonal complements instead of 
-dual spaces and annihilators. In \cite{BONCHI201277}, the authors 
-extended.
+\cite{boreale2009weighted}, since we are fixing weights on $\R$ and computing the orthogonal complements instead of 
+dual spaces and annihilators.
 
@@
 
-The \texttt{BPR} algorithm starts from a relation $R_0$, that is the complement 
+From Lemma \ref{lem:coincide}, it follows that given an LWA $L = (V,o,t)$ and a corresponding
+basis of $\llwb$ computed by BPR, one can check language equivalence 
+of two vectors in the state space, $v_1 \sim_l v_2 $,  by checking if $(v_1 - v_2) \in \mker{\llwb}$.
+Therefore, we can say that BPR "\textit{minimizes}", or it computes the whole
+binary linear relation $\llwb$, coinciding with $\sim_l$.
+
+The \texttt{BPR} algorithm starts from the basis of a relation $R_0$, that is the complement 
 of the relation identifying vectors with equal weights.
@@
 
-In the next sections we will compare execution of our implementation of the algorithms
+\texttt{BPR} has a cost of $O(n^4)$ operations 
+to initially compute the largest linear weighted bisimulation,
+which can be eventually reduced to $O(n^3)$ \cite{BONCHI201277}.
+In our implementation, by initially computing a basis of the orthogonal complement of $\llwb$,
+the cost of checking if two vectors are language equivalent is then reduced to the
+cost of matrix multiplication ($O(n^2)$). It follows that 
+\texttt{BPR} is a great choice when we
+have to decide if a large number of vectors in a WA are language equivalent.
+
+In the next sections we will compare execution results of our implementation of the algorithms
 \texttt{BPR} and \texttt{HKC} to verify correctness,
@@
 %TODO
-TODO costo computazionale HKC.
-
-\texttt{BPR} has a cost of $O(n^4)$ operations 
-to initially compute the largest linear weighted bisimulation,
-which can be eventually reduced to $O(n^3)$ \cite{BONCHI201277}.
-In our implementation, by initially computing a basis of the orthogonal complement of $\llwb$,
-the cost of checking if two vectors are language equivalent is then reduced to the
-cost of matrix multiplication ($O(n^2)$). \texttt{BPR} is a great choice when we
-have to decide if a large number of vectors in a WA are language equivalent.
+%TODO costo computazionale HKC.
 
@@
 \subsection{HKC Algorithm}
-We give a pseudocode definition of the \texttt{HKC} procedure:
+We give the pseudocode definition of the \texttt{HKC} procedure from \cite{DBLP:journals/corr/Bonchi0K17}:
 
@@
 Where $\mker{o}^0$ is an annihilator.
+
+\begin{note}
+  Given two vector spaces $V_1, V_2$ we write $V_1 + V_2$ to 
+  denote $\mspan{V_1 \cup V_2}$
+\end{note}
+
 The algorithm stops when $R_{j+1} = R_j$. An index $j \leq \dim(V)$ is 
-guaranteed to exist, such that the algorithms terminates at step $j$.
+guaranteed to exist, such that the algorithms stops at step $j$.
 It follows that $\llwb = R_j^0$.

### 2020-10-08 paper/sections/conclusion.tex
refined f1-tolerance test, working on paper

@@
 
-%TODO 
-TODO da fare
+During the testing phase, we have found that on randomly generated automata 
+over the $\R$ field, the probability of BPR returning an empty basis of $\llwb$
+increases together with the number of states in the automaton, and the number of
+ symbols in the alphabet.
+In a given linear weighted automaton $L = (V, o,t )$
+over the $\R$ field and an alphabet $A$, the chance of BPR returning an empty basis also
+increases together with the maximum weight in modulo of the transition matrices,
+$\forall a \in A, \; \max |(t_a)_{ij}|$, for $i,j=1\hdots,n$, with 
+$n = \dim(V)$. Authors of \cite{BONCHI201277} have confirmed
+that it is normal for this to happen and that this fact is not due
+to an implementation error.
+
+Since the tests in this work rely heavily on a non-empty basis of $\llwb$ to generate
+language equivalent vectors, the automata with an empty $\mker{\llwb}$ had to be ignored during the 
+tests. This introduced substantial overhead as the chance of BPR returning an empty basis 
+grew closer to 1, together with the various parameter of the automata: 
+most of the computing time was spent on generating and running BPR on automata which did
+not contain any language equivalent vectors, making tests on large automata not possible.
+
+To provide better runtime results for the language equivalence problem, 
+a topic worth further attention is the development of a semi-randomized method to 
+generate structured large sized automata that have a low chance of BPR
+returning an empty basis of $\llwb$.
+
+Another idea for the future of this implementation, is the addition
+of a \texttt{Semiring} interface type that would permit much more insightful analysis
+on weighted automata, not only on the field of real numbers.
+
+To further verify of the algorithms implemented in this work, 
+and any additional algorithm that may be implemented, 
+one could compare the results of this package with the results 
+of the PAWS tool for the analysis of weighted systems \cite{konig2017paws}, developed 
+at the University of Duisburg-Essen.

### 2020-10-08 paper/sections/implementation.tex
refined f1-tolerance test, working on paper

@@
 \subsection{Data Structures}
-In this implementation, the data structure for representing weighted automata is a \texttt{struct} containing
-the following attributes:
-\begin{enumerate}
-    \item An integer \texttt{Dim} representing the number of states $n$. 
-    (note: states are finite and indexed on natural numbers $0, \hdots, n-1$)
-    \item A slice of strings representing the \textit{alphabet} $A$.
-    \item A map of strings as keys (the alphabet symbols) and dense $n \times n$ \texttt{float64} matrices as values,
-    representing the set of transition functions.
-    \item A dense \texttt{float64} vector, representing the linearization of the output function. 
-    \item A \texttt{float64} value representing the tolerance to be used in numerical computations.
-    \item An optionally \texttt{nil}, dense floating point matrix providing a basis for the orthogonal 
-    complement of $\llwb$
-\end{enumerate}
+In this implementation, the data structure for representing weighted automata is a \texttt{struct}:
+
+\captionof{figure}{Source code for the Weighted Automaton data structure, found in \texttt{automata/automata.go}}
+\lstinputlisting[language=go]{../automata/automata.go}
+
+%\begin{enumerate}
+%    \item An integer \texttt{Dim} representing the number of states $n$. 
+%    (note: states are finite and indexed on natural numbers $0, \hdots, n-1$)
+%    \item A slice of strings representing the \textit{alphabet} $A$.
+%    \item A map of strings as keys (the alphabet symbols) and dense $n \times n$ \texttt{float64} matrices as values,
+%    representing the set of transition functions.
+%    \item A dense \texttt{float64} vector, representing the linearization of the output function. 
+%    \item A \texttt{float64} value representing the tolerance to be used in numerical computations.
+%    \item An optionally \texttt{nil}, dense floating point matrix providing a basis for the orthogonal 
+%    complement of $\llwb$
+%\end{enumerate}
 \begin{note}
@@
 \end{note}
-The definition can be found in file \texttt{automata/automata.go}.
-We then provide methods for reading an automaton from a text stream, applying transition
+
+We then provide methods for reading an automaton from a text stream, applying transitions
 and output functions to vectors, and generating random automata with real and natural valued weights.
@@
 of $L$, as seen in \cite{BONCHI201277}.
-For each step $i = 0, \hdots, j+1$, with $j \leq \dim(L)$, the implemented algorithm:
+To compute $R_{i+1}$ at each step, the implemented BPR algorithm:
 \begin{enumerate}
-    \item For each $a \in A$ computes $t_a^T(R_i)$ through matrix multiplication
-    \item Concatenates those matrices $t_a^T(R_i)$ to $R_i$ in a resulting matrix $G$ 
-    \item Instead of checking linear independence of the columns through Gaussian elimination,
-    it computes $R_{i+1}$ as the orthonormal basis of the column space of $G$, through singular 
+    \item Computes $t_a^T(R_i)$ through matrix multiplication for each $a \in A$
+    \item Concatenates the resulting matrices to $R_i$ in a resulting matrix $G$ 
+    \item Computes $R_{i+1}$ as the orthonormal basis of the column space of $G$, through singular 
     value decomposition.
@@
 compute $W^\perp = \mker{A^T}$, where $A$ is the matrix whose column space 
-is $W$. Proof is available in \cite{ila}.
+is $W$.
 

### 2020-10-08 paper/sections/introduction.tex
refined f1-tolerance test, working on paper

@@
 \label{sec:intro}
-TODO da fare
 
-Checking whether two nondeterministic finite automata (NFA) accept the same
-language is important in many application domains such as compiler construction and model checking. 
-Unfortunately, solving this problem is costly: it is
-PSPACE-complete (TODO cita) 
-% TODO cita
+Weighted automata (WAs) are a generalization of non-deterministic automata.
+When reading a symbol, a non-deterministic automaton 
+can transition in different states simultaneously. Weighted automata
+introduce \textit{weights} over transitions, which can for example, 
+represent the cost of a transition, 
+or in probabilistic systems, the chance of such transition to happen.
+WAs can be represented with a set of states, an output function and
+a set of transition matrices, indexed over the symbols in the alphabet
+the automaton can read. While those automata are typically defined over semirings,
+our implementation will focus for simplicity on automata with 
+transitions defined over the real number field. 
+The current configuration of a finite weighted automaton $W$,
+defined on $n$ states, will be represented with a column vector of length
+$n$, with values over the semiring or field on which transition weights of the
+automaton $W$ are also defined.
 
-In \cite{DBLP:journals/corr/Bonchi0K17}, up-to techniques are defined for
-weighted systems over arbitrary semirings, while in \cite{BONCHI201277}, up-to techniques
-are defined for Linear Weighted Automata (LWAs), under a more abstract coalgebraic perspective.
+The goal of this work is to provide an high-performance implementation 
+in the Go programming language of two different techniques 
+to compute \textit{weighted language equivalence}.
+Such equivalence relation is a bisimulation:
+a relation $R$ is a bisimulation whenever two states $v_1, v_2$ in $R$ 
+can simulate each other, resulting in a pair that is still in $R$.
+Two state vectors $v_1, v_2$ in a weighted automaton are said to be 
+weighted language equivalent, written as $v_1 \sim_l v_2$, when 
+they simulate each other by accepting the same words with the same resulting output weights.
 
-A well detailed example of the comparison of algorithms to compute 
-language equivalence, precisely between \texttt{HKC} and an alternative
-algorithm called the \textit{antichain algorithm} (\cite{de2006antichains}),
+The first technique we implement, 
+defined in \cite{DBLP:journals/corr/Bonchi0K17}, is an up-to technique
+for weighted language equivalence called HKC. It is defined for
+weighted systems over arbitrary semirings and can be implemented with 
+set theoretic constructs. The second technique is defined in \cite{BONCHI201277}:
+a coalgebraic perspective is adopted to define a technique for language equivalence 
+which exploits the linear representation of an automaton. 
+This latter technique "\textit{minimizes}": by linearizing the state space of a
+weighted automaton, 
+it computes a basis for an entire linear relation (see definition \ref{def:linrel})
+which coincides
+with weighted language equivalence.
+This technique for finite weighted automata over fields was first 
+introduced by Michele Boreale in \cite{boreale2009weighted}.
+
+Another example of the comparison between algorithms to compute 
+language equivalence, precisely between HKC and an alternative
+algorithm called the antichain algorithm (\cite{de2006antichains}),
 was published in 2017 \cite{fu2017equivalence}.

### 2020-10-08 paper/sections/notation.tex
first public release
refined f1-tolerance test, working on paper

@@
 \label{sec:notation}
-
-\begin{note}
-  Given two vector spaces $V_1, V_2$ we write $V_1 + V_2$ to 
-  denote $\mspan{V_1 \cup V_2}$
-\end{note}
 
@@
 $ \sqmap {v_1} = \sqmap{v_2}$. One can extend the notion of language 
-equivalence to states rather than for vectors by assigning 
+equivalence to states rather than vectors by assigning 
 to each state $x \in X$ the corresponding  unit vector 
@@
   transitive. 
+  An equivalence relation which is compatible with all the operations of
+the algebraic structure on which it is defined on, is called a 
+\textit{congruence relation}. Compatibility with the algebraic structure operations
+means that algebraic operations applied on equivalent elements will still
+yield equivalent elements.
+
 \end{defn}
@@
 
-An equivalence relation which is compatible with all the operations of
-the algebraic structure on which it is defined on, is called a 
-\textit{congruence relation}. Compatibility with the algebraic structure operations
-means that algebraic operations applied on equivalent elements will still
-yield equivalent elements.
+\begin{defn}
+  \label{def:congclos}
+  \textit{Generating set for the congruence closure:} \\
+  Let $\K$ be a field, $X$ a finite set and $R \subseteq \K^X \times \K^X$
+  a relation. Let $(v, v') \in \K^X \times \K^X$ be a pair of vectors.
+  The generating set is defined as  $U_R = \left\lbrace u - u' \mid (u, u') \in R \right\rbrace$.
+  Then $(v, v') \in c(R) \iff (v - v') \in U_R$
+\end{defn}
 
@@
   $o: V \to \K$ is a linear map associating to each state its output weight,
-  and $t = \left\lbrace t_a = V \times V \right\rbrace_{a \in A}$ is
+  and $t = \left\lbrace t_a \in \K^{V \times V} \right\rbrace_{a \in A}$ is
   the set of transition functions, represented with liner maps 
@@
   in $V$.
-  As in \cite{boreale2009weighted}, we have that $\dim{(L)} = \dim{(V)}$.
+  As seen in \cite{boreale2009weighted}, we have that $\dim{(L)} = \dim{(V)}$.
 \end{defn}
@@
 by considering the free vector space generated by the set of states $X$ in the WA,
-and by linearizing $o$ and $t$. If $X$ is finite, as in our implementations
-of the algorithms, we can use the same matrices for 
+and by linearizing $o$ and $t$. If $X$ is finite we can use the same matrices for 
 $t$ and $o$ in both the WA and the corresponding LWA.
@@
 \begin{defn}
-  The language recognized by a vector $v \in V$ of an LWA $(V,o,t)$ is defined
+  The language recognized by a vector $v \in V$ in an LWA $(V,o,t)$ is defined
   for all words $w \in A*$ as $\langlwa{v}{V}(w) = o(v_n)$ where $v_n$ is the 
   vector reached from $v$ through the composition of the transition functions
-  corresponding to the words in $w$.
+  corresponding to each symbol in $w$.
 
@@
 
-We define $\llwb$ as the behavioral equivalence for a given LWA $(V, o, t)$ as
+We define the equivalence relation  $\llwb$ for a given LWA $L = (V, o, t)$ as
 
@@
 
-Proof is available in section 3.3 of \cite{BONCHI201277}
+Proofs are available in section 3.3 of \cite{BONCHI201277}
 
@@
 \begin{defn}
+  \label{def:linrel}
   \textit{Linear Relations:}\\
@@
   \end{equation*}
-  The relation $R$ is linear if there exists a subspace $U$ such that $R \equiv R_U$.
+  The relation $R$ is linear if there exists a subspace $U$ such that $R = R_U$.
   A linear relation is a total equivalence relation on $V$.
@@
 
-  Let $(X, o, t)$ be a WA and $(\K^X, o^\sharp, t^\sharp)$ the corresponding linear 
-  weighted automaton. Then $\forall x \in X, \;\; \sqmap{x} = \langlwa{x}{\K^X}$
+  %Let $(X, o, t)$ be a WA and $(\K^X, o^\sharp, t^\sharp)$ the corresponding linear 
+  %weighted automaton. Then $\forall x \in X, \;\; \sqmap{x} = \langlwa{x}{\K^X}$
 \end{lem}
+
+\begin{figure}[htbp!]
+  \centering
+  \input{automata.tex}
+  \begin{equation*}
+    \begin{aligned}
+      o = \begin{pmatrix}
+        2 & 1 & 1
+      \end{pmatrix} & \quad &  
+      T_a = \begin{pmatrix}
+        1 & \frac{1}{3} & 1 \\ 
+        0 & 0 & 0 \\
+        0 & 0 & 0 \\ 
+      \end{pmatrix} & \quad &
+      T_b = \begin{pmatrix}
+        1 & 0 & 0 \\ 
+        0 & 0 & 3 \\
+        0 & \frac{1}{3} & 0 \\ 
+      \end{pmatrix} & \quad &
+    \end{aligned}
+  \end{equation*}
+  \caption{Example of a Weighted Automata}
+  \label{fig:autexmp}
+\end{figure}

### 2020-10-08 paper/sections/runtime.tex
fix
refined f1-tolerance test, working on paper

@@
 
-TODO da fare e sistemare.
+%TODO da fare e sistemare.
 
+%\subsection{Tolerance and F1 Score}
+The first runtime result we introduce is the study of the influence 
+of the tolerance value choice over the correctness of the algorithms results.
+Tests took place on randomly generated automata. A basis of  $\llwb$ was 
+computed for each automaton with BPR. An arbitrary number of vector pairs were generated 
+as a random linear combination of the vectors in the spanning set of $\llwb$. 
+An identical quantity of completely random vector pairs were generated.
+BPREquivalence and HKC procedures were then executed with those pairs in input.
+
+By defining pairs of vectors as
+
+\begin{itemize}
+    \item \textit{true positives} (TP) if reported as language equivalent by both 
+    procedures
+    \item \textit{true negatives} (TN) if reported as not language equivalent by both 
+    procedures
+    \item \textit{false positives} (FP) if reported as language equivalent by HKC but not by BPR
+    \item \textit{false negatives} (FN) if reported as  language equivalent by BPR but not by HKC
+\end{itemize}
+
+We have introduced \textit{accuracy, precision, recall} and \textit{F1 score} over 
+the tests results:
+
+\begin{equation*}
+    \begin{aligned}
+        \text{Accuracy} = \dfrac{TP + TN}{TP + TN + FP + FN} & \quad & 
+        \text{Recall} = \dfrac{TP}{TP + FN} \\ && \\
+        \text{Precision} = \dfrac{TP}{TP + FP} & \quad &
+        \text{F1} = 2 * \dfrac{\text{Precision} \cdot \text{Recall}}{\text{Precision} + \text{Recall}}  \\ && \\
+    \end{aligned}
+\end{equation*}
+
+We have then collected 
+the F1 score in relation to varying tolerance values.
+The plot is shown in Figure \ref{fig:f1}.  
 The program was executed on an x86\_64 AMD Ryzen 5 2600X Six Core CPU, running at 3.6GHz with 32GB RAM, 
-running Void Linux. 
-For 3 test runs on 100000 automata composed of 4 states, 2 alphabet symbols and transition
-matrices with weights on natural numbers $\leq 3$, with a tolerance of 10e-14,
-we found a non empty kernel of $\llwb$ for an average of 1718 automata.
+running Void Linux. Executed sequentially, F1 score tests took 2m26.19s
+Parallelized with a worker pool, the tests took 1m13.12s.
 
-%TODO
-(TODO spiega: più aumentano i simboli/stati più si riduce la probabilità di 
-trovare $\llwb$ non vuoto?? investiga, trova un modo di generare automi con LLWB 
-non vuoto)
+% //TODO: condizionamento BPR
+TODO: BPR conditioning
 
-By generating 1000 random couples of language equivalent vectors $(v_1, v_2)$ such that
-$(v_1 - v_2) \in \mker{\llwb}$ and 1000 couples of completely random vectors
-$(w_1, w_2)$, for each automata, we verified with 100\% confidence that:
+\begin{figure}[htbp!]
+    \centering
+    \includegraphics[width=.8\textwidth]{./plots/f1-tol.pdf}
 
-\begin{center}
-    $v_1 \; R_j \; v_2 \iff$ \texttt{HKC}$(v_1, v_2) =$ \texttt{false}
-\end{center}
+    \caption{F1 score over tolerance tests, 1000 vector pairs tested per automaton.}
+    \label{fig:f1}
+\end{figure}
 
-Where $R_j$ is the largest linear weighted bisimulation $\llwb$ computed by \texttt{BPR}.
-
-%TODO
-
-Sample output (19992 KBs in memory):
-\begin{verbatim}
-    $ time go run main.go 
-    tests completed            100000...
-
-    1000 samples per automaton
-    100000 automata
-    LLWP is not empty for 1726 automata
-    language equivalence HKC = BPR for 3452000 samples
-    BPR equivalence verified for 1726000 samples
-    HKC equivalence verified for 1726000 samples
-    confidence is 100%
-
-    real    0m21.322s
-    user    0m23.614s
-    sys     0m0.961s
-\end{verbatim}
+TODO empty LLWB chance, memory and time plots, varying automata size parameters

### 2020-10-08 automata/automata.go

@@ -1,16 +1,13 @@
 // This file contains weighted automata data structure definition
-// and basic transition functions.
 
 package automata
 
 import (
-	"math"
-
 	"gonum.org/v1/gonum/mat"
 )
 
 type Automaton struct {
-	// The input alphabet
+	// The input alphabet slice
 	A []string
 	// Transition matrices are maps from input strings
 	// to dense real valued matrices
@@ -26,35 +23,4 @@ type Automaton struct {
 	LLWBperp *mat.Dense
 	BPRTol   float64 // tolerance value for BPR
 	HKCTol   float64 // tolerance value for HKC
-}
-
-// applies the output function to a given state vector (o * v)
-func (a Automaton) GetOutput(v *mat.VecDense, tol float64) float64 {
-	res := mat.Dot(a.O, v)
-	if math.Abs(res) < tol {
-		res = 0.0
-	}
-	return res
-}
-
-// apply (multiply) a transition function for a given symbol s, to a vector v
-func (a Automaton) ApplyTransition(s string, v *mat.VecDense) *mat.VecDense {
-	res := mat.VecDenseCopyOf(v)
-	res.MulVec(a.T[s], v)
-	return res
-}
-
-// apply (multiply) a transpose transition function for a given symbol s, to a vector v
-func (a Automaton) ApplyTransposeTransition(s string, v *mat.VecDense) *mat.VecDense {
-	res := mat.VecDenseCopyOf(v)
-	res.MulVec(a.T[s].T(), v)
-	return res
-}
-
-// apply (multiply) a transpose transition function for a given symbol s, to a
-// matrix b, which column space spans a subspace
-func (a Automaton) ApplyTransposeTransitionBasis(s string, b *mat.Dense) *mat.Dense {
-	res := mat.DenseCopyOf(b)
-	res.Mul(a.T[s].T(), b)
-	return res
 }

### 2020-10-08 automata/methods.go

@@ -1,0 +1,40 @@
+// this file contains transition functions methods.
+package automata
+
+import (
+	"math"
+
+	"gonum.org/v1/gonum/mat"
+)
+
+// GetOutput applies the output function to a given state vector (o * v)
+func (a Automaton) GetOutput(v *mat.VecDense, tol float64) float64 {
+	res := mat.Dot(a.O, v)
+	if math.Abs(res) < tol {
+		res = 0.0
+	}
+	return res
+}
+
+// ApplyTransition applies (multiplies) a transition function for a given symbol s
+// to a vector v
+func (a Automaton) ApplyTransition(s string, v *mat.VecDense) *mat.VecDense {
+	res := mat.VecDenseCopyOf(v)
+	res.MulVec(a.T[s], v)
+	return res
+}
+
+// apply (multiply) a transpose transition function for a given symbol s, to a vector v
+func (a Automaton) ApplyTransposeTransition(s string, v *mat.VecDense) *mat.VecDense {
+	res := mat.VecDenseCopyOf(v)
+	res.MulVec(a.T[s].T(), v)
+	return res
+}
+
+// apply (multiply) a transpose transition function for a given symbol s, to a
+// matrix b, which column space spans a subspace
+func (a Automaton) ApplyTransposeTransitionBasis(s string, b *mat.Dense) *mat.Dense {
+	res := mat.DenseCopyOf(b)
+	res.Mul(a.T[s].T(), b)
+	return res
+}

### 2020-10-08 lin/colspace.go

@@ -6,16 +6,15 @@ import (
 	"gonum.org/v1/gonum/mat"
 )
 
-// compute an orthonormal basis of the column space of a through
-// svd decomposition. Cost is O(n^3)
-func OrthonormalColumnSpaceBasis(a mat.Matrix, tol float64) mat.Matrix {
+// OrthonormalColumnSpaceBasis computes an orthonormal basis of the column space of a through
+// svd decomposition. Cost is O(n^3). Also returns the condition number
+func OrthonormalColumnSpaceBasis(a mat.Matrix, tol float64) (mat.Matrix, float64) {
 	var svd mat.SVD
 	if ok := svd.Factorize(a, mat.SVDFullU); !ok {
 		log.Fatal("failed to factorize A")
 	}
 	u := mat.NewDense(1, 1, nil)
 	u.Reset()
-	//fmt.Println(svd.Cond())
 	svd.UTo(u)
 	//fmt.Println(mat.Cond(a, 2))
 	//fmt.Println(mat.Cond(u, 2))
@@ -32,5 +31,5 @@ func OrthonormalColumnSpaceBasis(a mat.Matrix, tol float64) mat.Matrix {
 	m, _ := u.Dims()
 	//fmt.Println(m, n, j)
 	basis := u.Slice(0, m, 0, j)
-	return basis
+	return basis, svd.Cond()
 }

### 2020-10-08 randtest/data.go

@@ -15,12 +15,15 @@ const (
 type BatchTestOptions struct {
 	AutOptions  *AutomatonTestOptions // initial automaton test options
 	NumAutomata int                   // number of automata to generate and test
+	Verbose     bool
 }
 
 func (opt BatchTestOptions) Print() {
-	fmt.Println("========= BATCH OPTIONS ===========")
-	fmt.Println(opt.NumAutomata, "automata")
-	opt.AutOptions.Print()
+	if opt.Verbose {
+		fmt.Println("========= BATCH OPTIONS ===========")
+		fmt.Println(opt.NumAutomata, "automata")
+		opt.AutOptions.Print()
+	}
 }
 
 // BatchResult represents the result of many weighted
@@ -61,15 +64,17 @@ func (r *BatchResult) ComputeStats() {
 }
 
 func (r BatchResult) Print() {
-	r.opt.Print()
-	fmt.Println("========= RESULTS ===========")
-	fmt.Println("LLWP is not empty for", float64(r.opt.NumAutomata)-r.Null, "automata")
-	fmt.Printf("TP: %10d TN: %10d\n", int(r.TP), int(r.TN))
-	fmt.Printf("FP: %10d FN: %10d\n", int(r.FP), int(r.FN))
-	fmt.Printf("accuracy:  %.20g\n", r.Accuracy)
-	fmt.Printf("precision: %.20g\n", r.Precision)
-	fmt.Printf("recall:    %.20g\n", r.Recall)
-	fmt.Printf("F1:        %.20g\n", r.F1)
+	if r.opt.Verbose {
+		r.opt.Print()
+		fmt.Println("========= RESULTS ===========")
+		fmt.Println("LLWP is not empty for", float64(r.opt.NumAutomata)-r.Null, "automata")
+		fmt.Printf("TP: %10d TN: %10d\n", int(r.TP), int(r.TN))
+		fmt.Printf("FP: %10d FN: %10d\n", int(r.FP), int(r.FN))
+		fmt.Printf("accuracy:  %.20g\n", r.Accuracy)
+		fmt.Printf("precision: %.20g\n", r.Precision)
+		fmt.Printf("recall:    %.20g\n", r.Recall)
+		fmt.Printf("F1:        %.20g\n", r.F1)
+	}
 }
 
 type AutomatonTestOptions struct {

### 2020-10-08 randtest/f1-tol.go

@@ -1,0 +1,139 @@
+package randtest
+
+import (
+	"fmt"
+	"time"
+
+	"github.com/alitto/pond"
+	"github.com/jinzhu/copier"
+	"gonum.org/v1/plot"
+	"gonum.org/v1/plot/plotter"
+	"gonum.org/v1/plot/plotutil"
+	"gonum.org/v1/plot/vg"
+)
+
+func F1TolTask() {
+
+	start := time.Now()
+
+	aopts := &AutomatonTestOptions{
+		NumStates:  4,
+		NumSymbols: 2,
+		NumSamples: 1000,
+		MaxWeight:  2,
+		Mode:       "nat",
+	}
+
+	bopts := &BatchTestOptions{
+		AutOptions:  aopts,
+		NumAutomata: 3000,
+		Verbose:     false,
+	}
+
+	tols := []float64{1e-17, 1e-16, 1e-15, 1e-14, 1e-13, 1e-12, 1e-11, 1e-10, 1e-9, 1e-8, 1e-7, 1e-6}
+	tolstr := []string{"1e-17", "1e-16", "1e-15", "1e-14", "1e-13", "1e-12", "1e-11", "1e-10", "1e-9", "1e-8", "1e-7", "1e-6"}
+
+	// points on the graph
+	ptsBoth := make(plotter.XYs, len(tols))
+	ptsBPR := make(plotter.XYs, len(tols))
+	ptsHKC := make(plotter.XYs, len(tols))
+
+	// ================================================================================================
+
+	pool := pond.New(10, 100)
+
+	// tests varying both tolerances
+	pool.Submit(func() {
+		fmt.Println("Running test varying both tolerances")
+		for i, tol := range tols {
+			aopts.BPRTol = tol
+			aopts.HKCTol = tol
+
+			res := BatchTest(bopts)
+			res.Print()
+			ptsBoth[i].X = float64(i)
+			ptsBoth[i].Y = res.F1
+		}
+	})
+
+	// tests varying HKC tolerance
+	pool.Submit(func() {
+		fmt.Println("Running test varying HKC tolerance")
+		HKCAutomataOpts := &AutomatonTestOptions{}
+		copier.Copy(HKCAutomataOpts, aopts)
+		HKCAutomataOpts.BPRTol = 1e-13
+
+		var HKCBatchOpts BatchTestOptions
[…]

### 2020-10-08 randtest/randtest.go

@@ -90,8 +90,8 @@ func TestSamplePair(az automata.Automaton, v1, v2 *mat.VecDense) int {
 	} else if !BPReq && !HKCeq {
 		return TN // true negative
 	} else if !BPReq && HKCeq {
-		return FP // false negative
+		return FP // false positive
 	} else {
-		return FN // false positive
+		return FN // false negative
 	}
 }

### 2020-10-08 randtest/results.go

@@ -1,0 +1,6 @@
+package randtest
+
+type TestingTask interface {
+	Run() error
+	SaveData() error
+}

### 2020-10-11 paper/sections/runtime.tex
added tolerance tests, added comments to code. 0 problems in code

@@
 %\subsection{Tolerance and F1 Score}
-The first runtime result we introduce is the study of the influence 
-of the tolerance value choice over the correctness of the algorithms results.
+We the studied of the influence 
+of the tolerance value over the correctness of the algorithms results.
 Tests took place on randomly generated automata. A basis of  $\llwb$ was 
 computed for each automaton with BPR. An arbitrary number of vector pairs were generated 
-as a random linear combination of the vectors in the spanning set of $\llwb$. 
-An identical quantity of completely random vector pairs were generated.
+as an uniformly distributed random linear combination of the vectors in the spanning set of $\llwb$. 
+The same quantity of vector pairs was generated randomly with an uniform distribution.
 BPREquivalence and HKC procedures were then executed with those pairs in input.
@@
     \centering
-    \includegraphics[width=.8\textwidth]{./plots/f1-tol.pdf}
-
-    \caption{F1 score over tolerance tests, 1000 vector pairs tested per automaton.}
+    \includegraphics[width=.75\textwidth]{./plots/f1-tol-1e-06.pdf}
+    \includegraphics[width=.75\textwidth]{./plots/f1-tol-1e-13.pdf}
+    \caption{F1 score over tolerance tests, 2000 vector pairs tested per automaton.}
     \label{fig:f1}
@@
 
+\begin{figure}[htbp!]
+    \centering
+    \includegraphics[width=.75\textwidth]{./plots/f1-tol-1e-16.pdf}
+    \includegraphics[width=.75\textwidth]{./plots/f1-tol-1e-20.pdf}
+    \caption{F1 score over tolerance tests, 2000 vector pairs tested per automaton.}
+    \label{fig:f1-2}
+\end{figure}
+
 TODO empty LLWB chance, memory and time plots, varying automata size parameters

### 2020-10-11 automata/backwards_test.go

@@ -25,9 +25,10 @@ func TestBackwardsPartitionRefinement(t *testing.T) {
 				0, 0.33333333333333333333333333333333333333, 0,
 			}),
 		},
-		Dim: 3,
-		O:   mat.NewVecDense(3, []float64{2, 1, 1}),
-		Tol: 10e-14,
+		Dim:    3,
+		O:      mat.NewVecDense(3, []float64{2, 1, 1}),
+		BPRTol: 10e-14,
+		HKCTol: 10e-14,
 	}
 
 	a.BackwardsPartitionRefinement()
@@ -38,7 +39,7 @@ func TestBackwardsPartitionRefinement(t *testing.T) {
 
 	// confront BPR and HKC results
 	resBPR := a.BPREquivalence(v1, v2)
-	resHKC, _ := a.HKC(v1, v2, 3000)
+	resHKC, _ := a.HKC(v1, v2)
 	fmt.Println(resBPR, resHKC)
 	assert.Equal(t, resHKC, resBPR)
 
@@ -46,7 +47,7 @@ func TestBackwardsPartitionRefinement(t *testing.T) {
 	v2 = mat.NewVecDense(3, []float64{2, 4, 0.5})
 
 	resBPR = a.BPREquivalence(v1, v2)
-	resHKC, _ = a.HKC(v1, v2, 3000)
+	resHKC, _ = a.HKC(v1, v2)
 	fmt.Println(resBPR, resHKC)
 	assert.Equal(t, resHKC, resBPR)
 
@@ -63,9 +64,10 @@ func TestBackwardsPartitionRefinement2(t *testing.T) {
 				0, 0, 1,
 			}),
 		},
-		Dim: 3,
-		O:   mat.NewVecDense(3, []float64{1, 1, 1}),
-		Tol: 10e-14,
+		Dim:    3,
+		O:      mat.NewVecDense(3, []float64{1, 1, 1}),
+		HKCTol: 10e-14,
+		BPRTol: 10e-14,
 	}
 
 	a.BackwardsPartitionRefinement()
@@ -76,7 +78,7 @@ func TestBackwardsPartitionRefinement2(t *testing.T) {
 
 	// confront BPR and HKC results
 	resBPR := a.BPREquivalence(v1, v2)
-	resHKC, _ := a.HKC(v1, v2, 3000)
+	resHKC, _ := a.HKC(v1, v2)
 	fmt.Println(resBPR, resHKC)
[…]

### 2020-10-11 automata/io.go

@@ -84,7 +84,8 @@ func readDense(scanner *bufio.Scanner, rows, cols int) (*mat.Dense, error) {
 	return mat.NewDense(rows, cols, data), nil
 }
 
-// Read an linear weighted automaton from a reader. If prompt is true
+// ReadAutomaton reads an linear weighted automaton from
+//  a reader. If prompt is true
 // then questions are asked (printed to stderr) before scanning
 func ReadAutomaton(r io.Reader, prompt bool) (*Automaton, error) {
 	a := &Automaton{}
@@ -146,7 +147,7 @@ func ReadAutomaton(r io.Reader, prompt bool) (*Automaton, error) {
 	return a, nil
 }
 
-// Decorated representation of an automaton
+// FancyString returns a decorated representation of an automaton
 func (a Automaton) FancyString() string {
 	var buf bytes.Buffer
 

### 2020-10-11 automata/methods.go

@@ -1,18 +1,14 @@
 // this file contains transition functions methods.
+
 package automata
 
 import (
-	"math"
-
 	"gonum.org/v1/gonum/mat"
 )
 
 // GetOutput applies the output function to a given state vector (o * v)
-func (a Automaton) GetOutput(v *mat.VecDense, tol float64) float64 {
+func (a Automaton) GetOutput(v *mat.VecDense) float64 {
 	res := mat.Dot(a.O, v)
-	if math.Abs(res) < tol {
-		res = 0.0
-	}
 	return res
 }
 
@@ -24,14 +20,16 @@ func (a Automaton) ApplyTransition(s string, v *mat.VecDense) *mat.VecDense {
 	return res
 }
 
-// apply (multiply) a transpose transition function for a given symbol s, to a vector v
+// ApplyTransposeTransition applies
+// (multiplies) a transpose transition function for a given symbol s, to a vector v
 func (a Automaton) ApplyTransposeTransition(s string, v *mat.VecDense) *mat.VecDense {
 	res := mat.VecDenseCopyOf(v)
 	res.MulVec(a.T[s].T(), v)
 	return res
 }
 
-// apply (multiply) a transpose transition function for a given symbol s, to a
+// ApplyTransposeTransitionBasis applies
+// (multiplies) a transpose transition function for a given symbol s, to a
 // matrix b, which column space spans a subspace
 func (a Automaton) ApplyTransposeTransitionBasis(s string, b *mat.Dense) *mat.Dense {
 	res := mat.DenseCopyOf(b)

### 2020-10-11 automata/pair.go

@@ -1,3 +1,6 @@
+// this file contains methods for handling vector pairs, exploiting the mat.Dense
+// matrix type
+
 package automata
 
 import (
@@ -6,7 +9,7 @@ import (
 	"gonum.org/v1/gonum/mat"
 )
 
-// creates a new pair of vectors
+// NewPair creates a new pair of vectors
 func NewPair(l, r *mat.VecDense) (*mat.Dense, error) {
 	ld, _ := l.Dims()
 	rd, _ := r.Dims()
@@ -22,17 +25,17 @@ func NewPair(l, r *mat.VecDense) (*mat.Dense, error) {
 	return m, nil
 }
 
-// return left element of a vector pair
+// PairLeft returns left element of a vector pair
 func PairLeft(p *mat.Dense) *mat.VecDense {
 	return p.ColView(0).(*mat.VecDense)
 }
 
-// return right element of a vector pair
+// PairRight returns right element of a vector pair
 func PairRight(p *mat.Dense) *mat.VecDense {
 	return p.ColView(1).(*mat.VecDense)
 }
 
-// return subtraction of elements of a vector pair
+// PairSub returns the subtraction of the elements of a vector pair
 func PairSub(p *mat.Dense) *mat.VecDense {
 	m, _ := p.Dims()
 	sub := mat.NewVecDense(m, nil)
@@ -40,13 +43,13 @@ func PairSub(p *mat.Dense) *mat.VecDense {
 	return sub
 }
 
-// return true if a matrix is a vector pair
+// PairCheck returns true if a matrix is a vector pair
 func PairCheck(p *mat.Dense) bool {
 	_, n := p.Dims()
 	return n == 2
 }
 
-// Returns true if two pairs equal each other
+// PairEqs returns true if two pairs equal each other
 func PairEqs(p, p1 *mat.Dense, tol float64) bool {
 	m, _ := p.Dims()
 	m1, _ := p1.Dims()

### 2020-10-11 automata/pairstack.go

@@ -8,13 +8,15 @@ import (
 	"gonum.org/v1/gonum/mat"
 )
 
+// NewPairStack creates a new pair stack, exploiting the mat.Dense matrix
+// data type
 func NewPairStack() *mat.Dense {
 	m := mat.NewDense(1, 1, nil)
 	m.Reset()
 	return m
 }
 
-// returns the size
+// PairStackSize returns the size of a pair stack
 func PairStackSize(s *mat.Dense) int {
 	if s.IsEmpty() {
 		return 0
@@ -23,7 +25,7 @@ func PairStackSize(s *mat.Dense) int {
 	return n
 }
 
-// push a pair into the stack
+// PairStackPush pushes a pair into the stack
 func PairStackPush(s *mat.Dense, p *mat.Dense) *mat.Dense {
 	if s.IsEmpty() {
 		return mat.DenseCopyOf(p)
@@ -32,7 +34,7 @@ func PairStackPush(s *mat.Dense, p *mat.Dense) *mat.Dense {
 	return s
 }
 
-// pop a pair from the stack
+// PairStackPop pops a pair from the stack
 func PairStackPop(s *mat.Dense) (*mat.Dense, error) {
 	if s.IsEmpty() {
 		return nil, errors.New("stack is empty")

### 2020-10-11 lin/gauss.go

@@ -1,86 +1,0 @@
-package lin
-
-import (
-	"math"
-
-	"gonum.org/v1/gonum/mat"
-)
-
-type testCase struct {
-	a [][]float64
-	b []float64
-	x []float64
-}
-
-func GaussPartial(a0 mat.Matrix) (*mat.Dense, *mat.VecDense, error) {
-	a := mat.DenseCopyOf(a0)
-	a.Copy(a0)
-
-	m, n := a.Dims()
-
-	// permutation matrix
-	p := mat.NewDense(m, m, nil)
-
-	for k := 0; k < m; k++ {
-		//fmt.Printf("before step %d:\n%.20g\n", k, mat.Formatted(a, mat.Squeeze()))
-
-		iMax := 0
-		max := -1.0
-		for i := k; i < m; i++ {
-			// compute scale factor s = max abs in row
-			s := -1.0
-			for j := k; j < n; j++ {
-				x := math.Abs(a.At(i, j))
-				if x > s {
-					s = x
-				}
-			}
-			// scale the abs used to pick the pivot
-			if abs := math.Abs(a.At(i, k)) / s; abs > max {
-				iMax = i
-				max = abs
-			}
-		}
-		//pivot := a.At(iMax, k)
-		//if pivot == 0 {
-		//	return a, nil, errors.New("singular matrix")
-		//}
-
-		// make permutation matrix for row swap
-		swaps := make([]int, m)
-		for i := 0; i < m; i++ {
-			swaps[i] = i
-		}
-		swaps[k] = iMax
-		p.Permutation(m, swaps)
-
-		//fmt.Printf("swap %d with %d\n", iMax, k)
-		//fmt.Printf("perm %d:\n%.20g\n", k, mat.Formatted(p, mat.Squeeze()))
-
-		a.Mul(p, a)
-
-		for i := k + 1; i < m; i++ {
-			for j := k + 1; j < n; j++ {
-				new := a.At(i, j) - a.At(k, j)*(a.At(i, k)/a.At(k, k))
-				a.Set(i, j, new)
-			}
-			a.Set(i, k, 0)
-		}
-
-		//fmt.Printf("step %d:\n%.20g\n", k, mat.Formatted(a, mat.Squeeze()))
-	}
-	// calculate the solution vector x
-	//x := make([]float64, m)
-	//for i := m - 1; i >= 0; i-- {
[…]

### 2020-10-11 lin/independent.go

@@ -1,76 +1,0 @@
-package lin
-
-import (
-	"math"
-
-	"gonum.org/v1/gonum/mat"
-)
-
-// find which column vectors in m are linearly independent and
-// return them
-func Independent(a *mat.Dense, tol float64) *mat.Dense {
-	var qr mat.QR
-	qr.Factorize(a)
-	m, n := a.Dims()
-	r := mat.NewDense(m, m, nil)
-	r.Reset()
-	qr.RTo(r)
-
-	rank := 0
-	cols := []int{}
-	for k := 0; k < n; k++ {
-		if math.Abs(r.At(k, k)) > tol {
-			rank++
-			cols = append(cols, k)
-		}
-	}
-
-	res := mat.NewDense(m, rank, nil)
-
-	j := 0
-	for _, k := range cols {
-		for i := 0; i < m; i++ {
-			res.Set(i, j, a.At(i, k))
-		}
-		j++
-	}
-
-	return res
-}
-
-// find which column vectors in m are linearly independent and
-// return them
-func IndependentGauss(a *mat.Dense, tol float64) *mat.Dense {
-	m, n := a.Dims()
-	ab := a.Grow(0, 1)
-
-	red, _, err := GaussPartial(ab)
-	if err != nil {
-		panic(err)
-	}
-
-	red = red.Slice(0, m, 0, n).(*mat.Dense)
-
-	rank := 0
-	cols := []int{}
-	for k := 0; k < n; k++ {
-		if k < m {
-			if math.Abs(red.At(k, k)) > tol {
-				rank++
-				cols = append(cols, k)
-			}
-		}
-	}
-
-	res := mat.NewDense(m, rank, nil)
-
-	j := 0
-	for _, k := range cols {
-		for i := 0; i < m; i++ {
-			res.Set(i, j, a.At(i, k))
-		}
-		j++
-	}
-
-	return res
-}

### 2020-10-11 lin/nullspace.go

@@ -7,6 +7,8 @@ import (
 	"gonum.org/v1/gonum/mat"
 )
 
+// Nullspace computes a null space basis of a given matrix,
+// with respect to tolerance.
 // columns of the returned matrix form an orthonormal basis
 // for the nullspace of matrix a, computed
 // through svd decomposition. also returns the maximum residual

### 2020-10-11 lin/subspace.go

@@ -4,7 +4,7 @@ import (
 	"gonum.org/v1/gonum/mat"
 )
 
-// returns true iff the vector b is included in U's column space
+// InSubspace returns true iff the vector b is included in U's column space
 // to do so we check if the matrix U and [U|b] have the same rank
 func InSubspace(u *mat.Dense, b *mat.VecDense, tol float64) bool {
 	var svd mat.SVD
@@ -24,7 +24,7 @@ func InSubspace(u *mat.Dense, b *mat.VecDense, tol float64) bool {
 	return ranku == rankub
 }
 
-// compute a basis for the intersection of many vector spaces
+// Intersect computes a basis for the intersection of many vector spaces
 func Intersect(tol float64, spansets ...mat.Matrix) mat.Matrix {
 	a := mat.DenseCopyOf(spansets[0])
 	for i, span := range spansets {
@@ -43,7 +43,7 @@ func Intersect(tol float64, spansets ...mat.Matrix) mat.Matrix {
 	return ker
 }
 
-// compute a basis for the union of many vector spaces
+// Union computes a basis for the union of many vector spaces
 func Union(spansets ...*mat.Dense) *mat.Dense {
 	n, j := spansets[0].Dims()
 
@@ -62,7 +62,7 @@ func Union(spansets ...*mat.Dense) *mat.Dense {
 	return basis
 }
 
-// compute the orthogonal complement of a vector subspace of R^n
+// Complement computes the orthogonal complement of a vector subspace of R^n
 func Complement(spanset mat.Matrix, tol float64) mat.Matrix {
 	n, m := spanset.Dims()
 	// the orthogonal complement of {} is R^n

### 2020-10-11 lin/util.go

@@ -12,7 +12,7 @@ import (
 	"gonum.org/v1/gonum/mat"
 )
 
-// returns true if a vector is composed only of zero values
+// IsZero returns true if a vector is composed only of zero values
 func IsZero(vec *mat.VecDense) bool {
 	for _, el := range vec.RawVector().Data {
 		if el != 0 {
@@ -23,7 +23,7 @@ func IsZero(vec *mat.VecDense) bool {
 	return true
 }
 
-// returns true if a vector is composed only of zero values, with a tolerance of tol
+// IsZeroTol returns true if a vector is composed only of zero values, with a tolerance of tol
 func IsZeroTol(vec *mat.VecDense, tol float64) bool {
 	for _, el := range vec.RawVector().Data {
 		if math.Abs(el) > tol {
@@ -34,7 +34,7 @@ func IsZeroTol(vec *mat.VecDense, tol float64) bool {
 	return true
 }
 
-// create an n*n identity matrix
+// EyeDense creates an n*n identity matrix
 func EyeDense(n int) *mat.Dense {
 	a := mat.NewDense(n, n, nil)
 	for i := 0; i < n; i++ {
@@ -52,7 +52,7 @@ func randFloat64(w float64) float64 {
 	return sign * rand.Float64() * w
 }
 
-// generate a normally distributed random n*n matrix
+// RandDense generates a normally distributed random n*n matrix
 func RandDense(n int, maxweight float64) *mat.Dense {
 	data := make([]float64, n*n)
 	for i := range data {
@@ -61,7 +61,7 @@ func RandDense(n int, maxweight float64) *mat.Dense {
 	return mat.NewDense(n, n, data)
 }
 
-// generate a normally distributed random integer n*n matrix
+// RandIntDense generates a normally distributed random integer n*n matrix
 func RandIntDense(n, max int) *mat.Dense {
 	data := make([]float64, n*n)
 	for i := range data {
@@ -70,7 +70,7 @@ func RandIntDense(n, max int) *mat.Dense {
 	return mat.NewDense(n, n, data)
 }
 
-// set z randomly chosen values in the matrix to 0
+// PokeHoles sets z randomly chosen values in the matrix to 0
 func PokeHoles(a *mat.Dense, z int) {
 	m, n := a.Dims()
 	for k := 0; k < z; k++ {
@@ -81,7 +81,7 @@ func PokeHoles(a *mat.Dense, z int) {
 	}
 }
 
[…]

### 2020-10-11 main.go

@@ -17,5 +17,9 @@ func main() {
 	// defer profile.Start(profile.MemProfile).Stop()
 	rand.Seed(time.Now().UnixNano())
 
-	randtest.F1TolTask()
+	//randtest.F1TolTask(1e-6)
+	//randtest.F1TolTask(1e-13)
+	randtest.F1TolTask(1e-15 / 2)
+	//randtest.F1TolTask(1e-16)
+	//randtest.F1TolTask(1e-20)
 }

### 2020-10-11 randtest/f1-tol.go

@@ -1,3 +1,5 @@
+// contains test for F1 score related to tolerance values
+
 package randtest
 
 import (
@@ -12,7 +14,7 @@ import (
 	"gonum.org/v1/plot/vg"
 )
 
-func F1TolTask() {
+func F1TolTask(fixedtol float64) {
 
 	start := time.Now()
 
@@ -30,8 +32,8 @@ func F1TolTask() {
 		Verbose:     false,
 	}
 
-	tols := []float64{1e-17, 1e-16, 1e-15, 1e-14, 1e-13, 1e-12, 1e-11, 1e-10, 1e-9, 1e-8, 1e-7, 1e-6}
-	tolstr := []string{"1e-17", "1e-16", "1e-15", "1e-14", "1e-13", "1e-12", "1e-11", "1e-10", "1e-9", "1e-8", "1e-7", "1e-6"}
+	tols := []float64{1e-22, 1e-21, 1e-20, 1e-19, 1e-18, 1e-17, 1e-16, 1e-15, 1e-14, 1e-13, 1e-12, 1e-11, 1e-10, 1e-9, 1e-8, 1e-7, 1e-6}
+	tolstr := []string{"1e-22", "1e-21", "1e-20", "1e-19", "1e-18", "1e-17", "1e-16", "1e-15", "1e-14", "1e-13", "1e-12", "1e-11", "1e-10", "1e-9", "1e-8", "1e-7", "1e-6"}
 
 	// points on the graph
 	ptsBoth := make(plotter.XYs, len(tols))
@@ -61,7 +63,7 @@ func F1TolTask() {
 		fmt.Println("Running test varying HKC tolerance")
 		HKCAutomataOpts := &AutomatonTestOptions{}
 		copier.Copy(HKCAutomataOpts, aopts)
-		HKCAutomataOpts.BPRTol = 1e-13
+		HKCAutomataOpts.BPRTol = fixedtol
 
 		var HKCBatchOpts BatchTestOptions
 		copier.Copy(&HKCBatchOpts, bopts)
@@ -83,7 +85,7 @@ func F1TolTask() {
 
 		BPRAutomataOpts := &AutomatonTestOptions{}
 		copier.Copy(BPRAutomataOpts, aopts)
-		BPRAutomataOpts.HKCTol = 1e-13
+		BPRAutomataOpts.HKCTol = fixedtol
 
 		var BPRBatchOpts BatchTestOptions
 		copier.Copy(&BPRBatchOpts, bopts)
[…]

### 2020-10-11 randtest/randtest.go

@@ -25,9 +25,11 @@ func BatchTest(opt *BatchTestOptions) BatchResult {
 
 }
 
+// Test a random automaton
 func TestRandAutomaton(o *AutomatonTestOptions) AutomatonResult {
 	var az automata.Automaton
 
+	// choose between random real valued weights or natural
 	switch o.Mode {
 	case "real":
 		az = automata.RandAutomaton(o.NumSymbols, o.NumStates, float64(o.MaxWeight))
@@ -41,14 +43,14 @@ func TestRandAutomaton(o *AutomatonTestOptions) AutomatonResult {
 	az.BPRTol = o.BPRTol
 	az.HKCTol = o.HKCTol
 
-	// fmt.Println(az)
 	az.BackwardsPartitionRefinement()
 
 	samples := make([]*mat.VecDense, o.NumSamples)
 	randoms := make([]*mat.VecDense, o.NumSamples)
+
+	// compute a basis of LLWB
 	llwb := lin.Complement(az.LLWBperp, o.BPRTol).(*mat.Dense)
 	_, dimLLWB := llwb.Dims()
-
 	lin.CleanTolDense(llwb, o.BPRTol)
 
 	if mat.Equal(llwb, mat.NewDense(o.NumStates, 1, nil)) {
@@ -57,10 +59,9 @@ func TestRandAutomaton(o *AutomatonTestOptions) AutomatonResult {
 		}
 	}
 
-	//lin.PrintMat(llwb)
-
 	autResult := AutomatonResult{Null: false}
 
+	// generate language equivalent (in LLWB) and random pairs of vectors
 	for i := range samples {
 		samples[i] = lin.LinearCombination(llwb, lin.RandVec(dimLLWB, 100))
 		randoms[i] = lin.RandVec(az.Dim, 100)

### 2020-10-11 paper/sections/conclusion.tex
finishing paper, added docs

@@
 
-Another idea for the future of this implementation, is the addition
+Another topic worth further investigation is the cause of the 
+sudden drop at 1e-15 in the F1 score plots.
+
+An idea for the future of this implementation, is the addition
 of a \texttt{Semiring} interface type that would permit much more insightful analysis

### 2020-10-11 paper/sections/implementation.tex
finishing paper, added docs

@@
     \item Two integers representing the number of pairs in the set, and the size of vectors.
-    \item A tolerance value to be used in equivalence checks. Best results were obtained with $10^{-14}$.
+    \item A tolerance value to be used in equivalence checks.
 \end{itemize}
@@
 
-At the end of \texttt{BPR}, we store the basis for the orthogonal complement
+At the end of \texttt{BPR}, we store $R_j$, the basis for the orthogonal complement
 of $\llwb$ as an attribute of the automaton. To check if two vectors $v_1 \llwb v_2$,
-we check that $R_j^T(v_1 - v_2) = \vec{0}$, with a prefixed tolerance.
+we check that $R_j^T(v_1 - v_2) = \vec{0}$, with a given tolerance.
 
@@
   $\xi \in  W^0$. Because $\xi(w)=0 \quad \forall  w \in W$, 
-  if $v$ represents $\xi$ we have that $(v, w)=\xi(w)=0$ for all $w \in W$. 
+  if $v$ represents $\xi$ we have that $\langle v, w \rangle =\xi(w)=0$ for all $w \in W$. 
   We obtain that $\varphi(W^0) \subseteq W^{\perp}$.
   If $v \in W^\perp$  
-  then the functional $x \mapsto (v, x)$ cancels over $W$ 
+  then the functional $\xi \mapsto \langle v, x \rangle$ cancels over $W$ 
   (by the definition of orthogonality).

### 2020-10-11 paper/sections/introduction.tex
finishing paper, added docs

@@
 the automaton can read. While those automata are typically defined over semirings,
-our implementation will focus for simplicity on automata with 
-transitions defined over the real number field. 
+for simplicity, our implementation will focus only on automata with 
+transitions defined over the field of real numbers. 
 The current configuration of a finite weighted automaton $W$,

### 2020-10-11 paper/sections/notation.tex
finishing paper, added docs

@@
   is a triple  $L = (V, o, \left\lbrace t_a \right\rbrace_{a \in A})$ 
-  where $V$ is a vector space representing the state space, 
+  where $V$ is a vector space representing the state space, and $\dim{(V)} = n$;
+  We have that
   $o: V \to \K$ is a linear map associating to each state its output weight,
-  and $t = \left\lbrace t_a \in \K^{V \times V} \right\rbrace_{a \in A}$ is
+  and $t = \left\lbrace t_a \in \K^{n \times n} \right\rbrace_{a \in A}$ is
   the set of transition functions, represented with liner maps 
@@
   in $V$.
-  As seen in \cite{boreale2009weighted}, we have that $\dim{(L)} = \dim{(V)}$.
+  As seen in \cite{boreale2009weighted}, we have that $\dim{(L)} = \dim{(V)} = n$.
 \end{defn}

### 2020-10-11 paper/sections/runtime.tex
finishing paper, added docs

@@
 computed for each automaton with BPR. An arbitrary number of vector pairs were generated 
-as an uniformly distributed random linear combination of the vectors in the spanning set of $\llwb$. 
+as uniformly distributed random linear combinations of the vectors in the spanning set of $\llwb$. 
 The same quantity of vector pairs was generated randomly with an uniform distribution.
@@
 
-We have introduced \textit{accuracy, precision, recall} and \textit{F1 score} over 
-the tests results:
+We then borrow four concepts from binary classification: \textit{accuracy, precision, recall} 
+and \textit{F1 score}:
 
@@
         \text{Precision} = \dfrac{TP}{TP + FP} & \quad &
-        \text{F1} = 2 * \dfrac{\text{Precision} \cdot \text{Recall}}{\text{Precision} + \text{Recall}}  \\ && \\
+        \text{F1} = 2 \cdot \dfrac{\text{Precision} \cdot \text{Recall}}{\text{Precision} + \text{Recall}}  \\ && \\
     \end{aligned}
@@
 
-We have then collected 
+We have then computed 
 the F1 score in relation to varying tolerance values.
-The plot is shown in Figure \ref{fig:f1}.  
+The plot is shown in Figures \ref{fig:f1} and \ref{fig:f1-2}.  
 The program was executed on an x86\_64 AMD Ryzen 5 2600X Six Core CPU, running at 3.6GHz with 32GB RAM, 
-running Void Linux. Executed sequentially, F1 score tests took 2m26.19s
-Parallelized with a worker pool, the tests took 1m13.12s.
+running Void Linux. Executed sequentially, F1 score tests took an average 2m26.19s,
+Parallelized by using a worker pool, the tests lasted the times reported in 
+Figure \ref{fig:f1} and \ref{fig:f1-2}.
+For example, in the first plot of Figure \ref{fig:f1-2}, the test took 82 seconds.
+Since we computed BPR for 3000 random automata, over 17 tolerance values, for 3 line plots,
+with a chance of finding a non-empty basis for $\llwb$ of approximately $1/3$,
+the BPR algorithm was executed an approximate average of
+$\frac{3000 \cdot 3 \cdot 17}{82} \approx 1865$ times per second. 
+The HKC algorithm, was executed 2000 times on each automaton with a non-empty 
+basis of $\llwb$, therefore, it was executed an approximate average of 
+$\frac{3000 \cdot 17 \cdot 2000 \cdot 3}{3 \cdot 82} \approx 1243902$ times per second.
 
 % //TODO: condizionamento BPR
-TODO: BPR conditioning
+%TODO: BPR conditioning
 
@@
 
-TODO empty LLWB chance, memory and time plots, varying automata size parameters
+%TODO empty LLWB chance, memory and time plots, varying automata size parameters

### 2020-10-11 automata/automata.go

@@ -1,4 +1,4 @@
-// This file contains weighted automata data structure definition
+// This file contains weighted automata data structure definitions
 
 package automata
 
@@ -6,6 +6,8 @@ import (
 	"gonum.org/v1/gonum/mat"
 )
 
+// Automaton represents a general finite weighted automaton on the
+// real number field
 type Automaton struct {
 	// The input alphabet slice
 	A []string

### 2020-10-11 lin/nullspace_test.go

@@ -15,7 +15,7 @@ func TestNullspace(t *testing.T) {
 		4.0, 5.0, 6.0,
 		7.0, 8.0, 9.0,
 	})
-	ker, res := Nullspace(a, test_tol)
+	ker, res := Nullspace(a, testTol)
 
 	//fmt.Printf("%.60e\n", res)
 
@@ -38,7 +38,7 @@ func TestNullspace2(t *testing.T) {
 			1, 7, 2,
 			-2, 3, 1,
 		})
-	ker, _ := Nullspace(a, test_tol)
+	ker, _ := Nullspace(a, testTol)
 
 	expker := mat.NewDense(3, 1,
 		[]float64{

### 2020-10-11 randtest/data.go

@@ -12,12 +12,14 @@ const (
 	FN            // false negative
 )
 
+// BatchTestOptions represents options for running batch tests on automata
 type BatchTestOptions struct {
 	AutOptions  *AutomatonTestOptions // initial automaton test options
 	NumAutomata int                   // number of automata to generate and test
 	Verbose     bool
 }
 
+// Print batch test options
 func (opt BatchTestOptions) Print() {
 	if opt.Verbose {
 		fmt.Println("========= BATCH OPTIONS ===========")
@@ -63,6 +65,7 @@ func (r *BatchResult) ComputeStats() {
 	r.F1 = 2 * ((r.Precision * r.Recall) / (r.Precision + r.Recall))
 }
 
+// Print results of a batch test
 func (r BatchResult) Print() {
 	if r.opt.Verbose {
 		r.opt.Print()
@@ -77,6 +80,7 @@ func (r BatchResult) Print() {
 	}
 }
 
+// AutomatonTestOptions represents settings for generating random automata
 type AutomatonTestOptions struct {
 	NumStates  int     // Number of states in automata
 	NumSymbols int     // Number of symbols in alphabet
@@ -87,6 +91,7 @@ type AutomatonTestOptions struct {
 	HKCTol     float64 // tolerance for HKC
 }
 
+// Print automaton test options
 func (opt AutomatonTestOptions) Print() {
 	fmt.Println("weight kind:", opt.Mode)
 	fmt.Println("max weight in modulo:", opt.MaxWeight)

### 2020-10-11 randtest/f1-tol.go

@@ -14,6 +14,8 @@ import (
 	"gonum.org/v1/plot/vg"
 )
 
+// F1TolTask runs a test suite on random automata, collects F1 in relation
+// to tolerance values and saves a plot in PDF and PNG format
 func F1TolTask(fixedtol float64) {
 
 	start := time.Now()

### 2020-10-11 randtest/randtest.go

@@ -10,6 +10,8 @@ import (
 	"gonum.org/v1/gonum/mat"
 )
 
+// BatchTest runs a batch test with given options
+// and computes statistics at the end
 func BatchTest(opt *BatchTestOptions) BatchResult {
 	batchResults := BatchResult{opt: opt}
 
@@ -25,7 +27,7 @@ func BatchTest(opt *BatchTestOptions) BatchResult {
 
 }
 
-// Test a random automaton
+// TestRandAutomaton tests a single random automaton with the given options
 func TestRandAutomaton(o *AutomatonTestOptions) AutomatonResult {
 	var az automata.Automaton
 
@@ -82,6 +84,9 @@ func TestRandAutomaton(o *AutomatonTestOptions) AutomatonResult {
 	return autResult
 }
 
+// TestSamplePair tests two of vectors for language equivalence
+// compares the results and returns the corresponding number for identifying
+// if the result is true/false positive/negative
 func TestSamplePair(az automata.Automaton, v1, v2 *mat.VecDense) int {
 	BPReq := az.BPREquivalence(v1, v2)
 	HKCeq, _ := az.HKC(v1, v2)

### 2020-10-11 randtest/results.go

@@ -1,6 +1,0 @@
-package randtest
-
-type TestingTask interface {
-	Run() error
-	SaveData() error
-}
